### 2018-01-13 Project log/multiple_outcomes_project_log.Rmd
Add files

---
title: ''
author: ''
header-includes:
- \usepackage[T1]{fontenc}
- \usepackage{microtype}
- \usepackage[margin=1in]{geometry}
- \usepackage{fancyhdr}
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead{}
- \fancyfoot{}
- \fancyhead[C]{Multiple outcomes}
- \fancyfoot[RO,LE]{\thepage}
- \usepackage{booktabs}
- \usepackage{lettrine}
- \usepackage{paralist}
- \usepackage{setspace}\singlespacing
- \usepackage{url}
- \usepackage{parskip}
- \usepackage{color,soul}
- \usepackage{palatino}
- \usepackage{float}
output: pdf_document
bibliography: refs_fake.bib
csl: aje_style.csl
---

\doublespacing

\begin{center}
\textbf{ \LARGE{Project Log: Multiple Outcomes} }
\vspace{10mm}
\end{center}

\doublespacing

\tableofcontents

\section{Stream of consciousness}

Since we are bootstrapping and then fitting parametric model (OLS), I think we should use Fox's or Westfall's residual resampling approach (and indeed, this seemed to work in the last simulation). Seems like the basic approach automatically violates OLS assumptions, meaning our inference will be wrong and the probability of rejecting the test could be non-nominal. Our application of bootstrapping is unusual in that we need to resample in a way that is compatible with parametric assumptions of the test we're going to apply for each outcome.

\textbf{Metric \#1}: 95\% CI for number of hits above expected value that we see in observed data (resample under original distribution, using first half of my theory)

\textbf{Metric \#2}: 95\% CI under the null for number of hits (resample under null, using second half of my theory)

What about other link functions or hypothesis tests not based on regression? How would we do residual resampling? Davison \& Hinkley talk about other GLMs. I think we'd end up with either some form of residual or $Y$ itself for each outcome, and then we'd jointly resample these things as planned.

Also, there might be models for which you can't resample as planned. For example, with multiple regression, we

<!-- NEW SECTION -->
\section{Possible resampling algorithms}

Note that all of these methods assume $Var(Y^{*} | X^{*}=0) = Var(Y^{*} | X^{*}=1)$; in the resampled data, this is the average of the original within-stratum variances. I think this is fine since we are \emph{not} avoiding parametric inference for OLS (unlike in other bootstrapping applications).

\underline{Basic "fix $X$, then resample $Y$"}
  \begin{enumerate}
  \item Fix $X$ and resample only $Y$: $(Y^{*}_i, X^{*}_i) = \left( \text{sample}(Y_i), X_i \right)$. So $Y^{*}_i \amalg X^{*}_i$. 
  \item Regress: $Y^{*}_i = \beta_0 + \beta X^{*}_i + \epsilon$.
  \item Test $H_0: \beta = 0$.
  \end{enumerate}

Has multimodality problem. No guarantee that residuals are normal, so OLS inference will be wrong anytime the original data are generated under $H_A$.

\begin{figure}[H]
\centering
\caption{Original data.}
\includegraphics[width=100mm]{orig_data.png}
\end{figure}

\begin{figure}[H]
\centering
\caption{Resampled data under basic algorithm. Loses association between sex and Y, so regression residuals are bimodal.}
\includegraphics[width=100mm]{tyler.png}
\end{figure}

\underline{Westfall's "center, then resample $Y$"}
  \begin{enumerate}
  \item Fix $X$ and set $Y^{*}_i = \text{sample}\left( Y - \widehat{Y} \right)$. So $Y^{*}_i \amalg X^{*}_i$. 
  \item Regress: $Y^{*}_i = \beta_0 + \beta X^{*}_i + \epsilon$.
  \item Test $H_0: \beta = 0$.
  \end{enumerate}

Avoids multimodality problem, but could still have non-normal residual distribution, as in the above. So OLS inference could be wrong, but only if original data already violate homoskedasticity. Works well in my simulations.

\begin{figure}[H]
\centering
\caption{Resampled data under Westfall's algorithm. Residuals are normal.}
\includegraphics[width=100mm]{westfall.png}
\end{figure}

\underline{Fox/Davison's residual resampling}

See Davison Algorithm 6.1 in "Linear Regression" chapter. 
\begin{enumerate}
  \item Fix $X$ and set $Y^{*}_i = \widehat{Y}_i + \text{sample}\left( Y - \widehat{Y} \right)$. So $Y^{*}_i \not\!\perp\!\!\!\perp X^{*}_i$.
  \item Regress: $Y^{*}_i = \beta_0 + \beta X_i + \epsilon$.
  \item Test $H_0: \beta = \widehat{\beta}$ (i.e., the fitted coefficient from original sample).
  \end{enumerate}

Avoids multimodality problem residuals will be normal as long as they are normal within strata of $X$, so OLS inference should be correct. $Var(Y^{*} | X^{*}=0) = Var(Y^{*} | X^{*}=0)$ and is an average of the original conditional variances.

\begin{figure}[H]
\centering
\caption{Resampled data under Fox's algorithm. Residuals are normal.}
\includegraphics[width=100mm]{fox.png}
\end{figure}

<!-- NEW SECTION -->
\newpage
\section{Other online resources}

\subsection{$\star$ CV: Permutation tests vs. bootstrap tests}

What we are doing is similar to a permutation test because we're not centering the observations. In this answer, Snow says "permutation tests test a specific null hypothesis of exchangeability, i.e. that only the random sampling/randomization explains the difference seen". Also, permutation tests are more powerful.

"Resample $Y$ while fixing $X$" strategy is "similar in spirit" to permutation tests and assumes that the distribution of $Y$ in each group of $X$ is identical. In contrast, "mean-center, then resample within each group of $X$" does not make this assumption.

\url{https://stats.stackexchange.com/questions/20217/bootstrap-vs-permutation-hypotheis-testing}

\subsection{$\star$ CV: Center first or resample first?}

\url{https://stats.stackexchange.com/questions/136661/using-bootstrap-under-h0-to-perform-a-test-for-the-difference-of-two-means-repl/187630}

Goes over difference between "resample, then center" (OP's first bootstrap) and "center, then resample" (OP's second bootstrap).

Accepted answer says that first approach is actually testing whether the distribution of x and y are identical. I DON'T REALLY GET THIS.

\subsection{CV: Cases when naive bootstrap fails}

\url{https://stats.stackexchange.com/questions/9664/what-are-examples-where-a-naive-bootstrap-fails/9722#9722}

\subsection{CV: Cases when naive bootstrap fails}
\url{https://stats.stackexchange.com/questions/11210/assumptions-regarding-bootstrap-estimates-of-uncertainty?noredirect=1&lq=1}

\subsection{CV: Applying bootstrap to arbitrary smooth function of the data}
\url{https://stats.stackexchange.com/questions/246632/smoothness-of-a-statistic-for-bootstrapping}

<!-- NEW SECTION -->
\newpage
\section{Notes on past literature}

\subsection{Davison \& Hinkley text}

\begin{itemize}
\item Linear regression chapter has great explanation of case resampling (Tyler's approach) vs. residual resampling (Fox, sort of Westfall). 
\item Pg 12 in "Linear Regression" chapter: Says that with case resampling, "the null hypothesis being tested is stronger than just zero slope"
\end{itemize}

\subsection{Good (2005) book chapter}

\begin{itemize}
\item Clear explanation of difference in assumptions between nonparametric bootstrap (equal parameters) and permutation test (exchangeability). 
\item I think the "resample, then center" strategy is closer to a permutation test since it forces equal distributions, while the "center, then resample" or "just resample" strategies are more like bootstrapping. 
\end{itemize}

\subsection{Fox chapter on bootstrapping regressions}

\begin{itemize}
\item {In section on bootstrap hypothesis testing, uses this residual-resampling algorithm: 
  \begin{enumerate}
  \item Fix $X$ and set $Y^{*}_i = \widehat{Y}_i + \text{sample}\left( Y - \widehat{Y} \right)$. So $Y^{*}_i \not\!\perp\!\!\!\perp X^{*}_i$.
  \item Regress: $Y^{*}_i = \beta_0 + \beta X_i + \epsilon$.
  \item Test $H_0: \beta = \widehat{\beta}$ (i.e., the fitted coefficient from original sample).
  \end{enumerate}
}
\end{itemize}

\subsection{Hall \& Wilson "Two guidelines"}

\begin{itemize}
\item {"Sample, then center" algorithm (here, I am extrapolating what they present to the regression setting):
  \begin{enumerate}
  \item Resample entire vector $(Y^{*}_i, X^{*}_i) = \text{sample}\left( (Y_i, X_i) \right)$. So $Y^{*}_i \not\!\perp\!\!\!\perp X^{*}_i$. 
  \item Regress: $Y^{*}_i = \beta_0 + \beta X^{*}_i + \epsilon$.
  \item Test $H_0: \beta = \widehat{\beta}$ (i.e., the fitted coefficient from original sample).
  \end{enumerate}
}

\item The point of using asymptotically pivotal statistics (centered and scaled) is that then the asymptotic distribution of the statistic does not depend on any unknowns.

\item When you can’t estimate variance of estimator in order to create pivot, suggests using accelerated bias correction bootstrap.
\end{itemize}

\subsection{$\star$ Troendle 2004 "Slow convergence"}

\begin{itemize}

\item Talks about multiple 2-sample $t$-tests

\item{ Two approaches (see page 3 for very clear overview):
  \begin{enumerate}
  \item Permute a subject’s outcome variables while assigning group labels (with \emph{uncentered} variables)  
  \item Bootstrap by resampling a subject’s entire row (with \emph{centered} outcome variables)
  \end{enumerate}
}

\item {"Center, then sample" bootstrapping algorithm (here, I am extrapolating what they present to the regression setting):
  \begin{enumerate}
  \item Center outcome by setting $\widetilde{Y}_i = Y_i - \widehat{Y}$
  \item Resample entire vector: $(Y^{*}_i, X^{*}_i) = \text{sample}\left( (\widetilde{Y}_i, X_i) \right)$. So $Y^{*}_i \amalg X^{*}_i$. 
  \item Regress: $Y^{*}_i = \beta_0 + \beta X^{*}_i + \epsilon$.
  \item Test $H_0: \beta = 0$.
  \end{enumerate}
}

\end{itemize}

\subsection{Westfall \& Young textbook}

\begin{itemize}

\item I think their algorithm is closest to the one Tyler initially suggested (fix $X$ while resampling $Y$), except that they first center the outcomes (i.e., resample the residuals).

\item{"Center, then resample" strategy
  \begin{enumerate}
  \item Center outcome by setting $\widetilde{Y}_i = Y_i - \widehat{Y}$
  \item Fix $X$ and resample only $Y$: $(Y^{*}_i, X^{*}_i) = \left( \text{sample}(\widetilde{Y}_i), X_i \right)$. So $Y^{*}_i \amalg X^{*}_i$. 
  \item Regress: $Y^{*}_i = \beta_0 + \beta X^{*}_i + \epsilon$.
  \item Test $H_0: \beta = 0$.
  \end{enumerate}
}

\item 106-109: More on OLS

\item Page 79: "Resampling from the residuals, rather than the actual data values, is important. If the actual data are pooled and resampled, then the underlying residual distribution will be estimated using a multi-modal empirical distribution function when the means differ. Such resampling violates the first principle of Hall \& Wilson (1991): '...even if the data might be drawn in a way that fails to satisfy $H_0$, resampling should be done in a way that reflects $H_0$'." I DON'T GET WHY UNCENTERED APPROACH VIOLATES H0!

\item Page 39-41: Analyzes $P( P^* < \alpha )$ in bootstrap iterates

\item Page 92-93: Re-randomization

\item Bootstraps of pivotal statistics converge faster (Section 2.2.2)

\end{itemize}

\subsection{Westfall \& Troendle (2008)}

\begin{itemize}

\item Setting: You have multiple outcomes measured on multiple categorical groups and are interested in full exchangeability on the outcome across the groups (not just different means, for example).

\item Shows that "permutation methods are distribution-free under an appropriate exchangeability assumption and FWER follows mathematically, regardless of sample size"

\item Implemented in SAS’ PROC MULTTEST

\item According to SAS' documentation, this(?) paper shows that: "when subset pivotality holds, the joint distribution of p-values under the subset is identical to that under the complete null"

\end{itemize}

\subsection{Bretz, Hothorn, \& Westfall (2008)'s R paper}

\begin{itemize}

\item They are fixing the Y and permuting the Xs (in their application, these are just the group labels), so are keeping the correlation structure of the Ys as we are (pg. 130)

\item Bootstrapping vs. permutation pros and cons (pg. 140)"

\end{itemize}

\subsection{Bickel}

\begin{itemize}
\item Classic paper that goes over basic asymptotics of bootstrap
\item Uses SLLN
\end{itemize}

\subsection{Chernick textbook}

\begin{itemize}
\item Great table comparing different CI methods with their assumptions (Ch 3, page 8)
\end{itemize}

\subsection{New reference}

\begin{itemize}
\item asdf
\end{itemize}

\newpage
\section{Q \& A}
\subsection{Open questions}

\subsection{Resolved questions}
Q: Why does Blakesley/Westfall’s minP bootstrapping work since it is an extreme order statistic?
A: Probably because minP is not an order statistic of the data themselves.

Q: If we go the confidence interval route, why not just treat the p-values as our data and resample directly from them to produce CI? 
A: Because there are only a fixed number of p-values (k), and this is not necessarily asymptotically large.

Q: Why does bootstrapping work with small samples since we need the ECDF in sample to go to the true CDF? Or does it not work in small samples?
A: Correct. It does not necessarily work in small samples.

<!-- NEW SECTION -->
\newpage
\section{Simulation notes}

\subsection{Summary of previous two}

\begin{itemize}
\item{ Plot1 (CIs)
  \begin{itemize}
    \item Shows that the residual resampling is working correctly (since results are same regardless of whether we generate under H0 or under HA)
    \item As expected, more variable when Y’s are more correlated
  \end{itemize}
}
\end{itemize}

\subsection{2017-8-1 (generate under alternative)}

\begin{itemize}
\item Same as 2017-7-25, but now generating under weak alternative (rho.XY = 0.03) 
\item Expect same exact results for CIs (since they only use results from resampling under H0), but higher rejection rates
\end{itemize}

\subsection{2017-7-25 (generate under null)}

\begin{itemize}
\item CIs are 95\% two-sided ones for both values of alpha 
\item Hypothesis tests are one-sided and have sample alpha as the individual tests
\item N=1,000 and B=2,000 (j=10 per simulation)
\item Figure out why file 125 has extra rows (try running it again)

\item{ \textbf{The apparently too-conservative behavior for the uncorrelated case is a benign artifact} of $\widehat{\theta}$'s discreteness.
  \begin{itemize}
    \item It occurs because when the Y’s are uncorrelated, the distribution of the number of rejections is less right-skewed, so it drops off more quickly in the tails. Since $\widehat{\theta}$ is discrete, we are forced to pick a quantile that doesn’t have exactly 5\% of the mass above it. By default, R inverts the ECDF, so uses the quantile with LESS than 5\% above it. Because the distribution of $\widehat{\theta}$ drops off quickly around the chosen quantile, we end up with conservative performance (see artifact\_plot1 and artifact\_plot2, where I shade in red the proportion that are above the variable’s own critical value).
    \item Artifact might improve a little with more simulation reps, but only to a point. 
    \item For this case, we could benchmark against the truth since $\widehat{\theta}$ is binomial. Indeed, the empirical quantiles of both \texttt{n.rej} and the bootstrap estimates are exactly correct. :) 
  \end{itemize}
}
\end{itemize}

\subsection{Summary so far}

\begin{itemize}
\item Seems like N=1,000 and B=2,000 is enough for good asymptotics 
\item When data are generated under null, resampling the Y’s or the residuals seems to works. But former theoretically should work only in more limited cases. 
\end{itemize}

\subsection{2017-7-24 “Sherlock sim”}

\begin{itemize}
\item Huge: N=10,000, B=10,000
\item Resampling under joint null (Y’s rather than residuals)
\item This seemed to work based on interim results (though I had trouble with existing stitching script, so stitched results in overall\_stitched folder might be wrong)
\end{itemize}

\subsection{2017-7-24 “Test smaller sim”}

\begin{itemize}
\item N=1,000, B=2,000 
\item Switched to resampling residuals instead of Ys themselves for the first time 
\item Seems to work! Going to try on cluster to have larger simulation
\end{itemize}

\subsection{2017-7-22}

\begin{itemize}
\item N=100, B=1,000: seems too small based on the below
\item Looked good at alpha=0.05 (4.4\% rejections) but maybe too conservative for alpha=0.01? (3.6\% rejections)
\item Took about 1.5 hours locally
\end{itemize}

\subsection{2017-7-21}

\begin{itemize}
\item With N=5,000, B=2,000, n.sims = 250, looks good
\item 5.5\% rejection rate on joint null for alpha = 0.05
\item 6\% rejection rate for alpha = 0.01
\item	Took about 30 hours locally
\item Seems good
\end{itemize}

\subsection{Summary so far}

\begin{itemize}
\item Resample under H0 by drawing Y’s separately; reject by inverted CI using its percentiles $\Rightarrow$ \textbf{WORKS} (see Sherlock sim)

\item Resample from original with single Y1; look at variance of mean $\Rightarrow$ \textbf{WORKS}

\item{ Bootstrap from original sample with 100 independent std. normal; reject using z-test that’s only a function of mean $\Rightarrow$ \textbf{DOESN'T WORK} 
  \begin{itemize}
    \item	The bootstrapped distribution of the absolute differences is too variable compared to true distribution
    \item	Hence rejects only 1.8\% of time
  \end{itemize}
}

\item Conjecture: Hall \& Wilson approach seems to rely on the idea that the standardized estimator is pivotal, i.e., that it comes from a location-scale family. This is why we can get away with only sampling from the original distribution. Since our estimator definitely isn’t from a location-scale distribution, maybe that’s why it does not work. Also, based on the different results between currently running one and very first one, I think B=500 is not enough, and B = 10,000 is enough. Not sure about intermediate ones.

\end{itemize}

\subsection{Summary so far}

\begin{itemize}
\item Seems like N=1,000 and B=2,000 is enough for good asymptotics 
\item When data are generated under null, resampling the Y’s or the residuals seems to works. But former theoretically should work only in more limited cases. 
\end{itemize}

For earlier simulation results, see old project log in Word.

<!--
\newpage
Here is a test reference[@evalue]. Use semicolons to separate multiple references. To change reference style, Google for a .csl style sheet for journal of interest of one with similar reference style. 
-->

\newpage
\section*{References}

### 2018-01-13 R code/Auxiliary code/Resample residuals vs. outcomes.R

@@ -1,0 +1,101 @@
+
+########################### REGRESSION ON ORIGINAL DATA ########################### 
+
+# simulate from multiple regression model, comparing residual resampling to full-case resampling
+# super bimodal on sex
+n = 1000
+male = rbinom( n=n, size=1, prob=0.5 )
+X = rnorm( n, mean = 4 * male, sd = 1 )
+bmale = 15
+bX = -1
+Y = rnorm( n, mean = bmale * male + bX * X, sd = 1 )
+
+d = data.frame( X, Y, male )
+
+plot(X,Y)
+
+# original data
+library(ggplot2)
+
+colors = c("pink", "blue")
+ggplot( data = d, aes( x = X, y = Y, color = as.factor(male) ) ) + geom_point(size=3) +
+  scale_color_manual(values = colors) + theme_bw()
+
+# super bimodal marginally on sex
+ggplot( data = d, aes( x = Y, color = as.factor(male) ) ) + geom_density() +
+  scale_color_manual(values = colors) + theme_bw()
+
+# regression
+( mo = lm( Y ~ X + male, data = d ) )
+lm(Y~X, data=d)
+
+# are the residuals normal?
+# yes, as expected
+hist(mo$residuals)
+
+
+########################### REGRESSION ON DATA RESAMPLING JUST Y ###########################
+
+# THIS DOES NOT WORK BECAUSE RESIDUALS ARE NO LONGER NORMAL. 
+# SO WE DEFINITELY CANNOT DO THIS APPROACH. 
+
+d$Y2 = sample( d$Y, replace = TRUE )
+
+# this plot shows that covariates are correlated, but association between X and Y
+#  key point: association between sex and Y also lost
+ggplot( data = d, aes( x = X, y = Y2, color = as.factor(male) ) ) + geom_point(size=3) +
+  scale_color_manual(values = colors) + theme_bw()
+
+ggplot( data = d, aes( x = Y2, color = as.factor(male) ) ) + geom_density() +
+  scale_color_manual(values = colors) + theme_bw()
+
+# regression
+( m2 = lm( Y2 ~ X + male, data = d ) )
+
+# are the residuals normal?
+# NOPE. THIS IS A PROBLEM. 
+hist(m2$residuals)
+
+
+########################### REGRESSION ON DATA RESAMPLING RESIDUALS (WESTFALL) ###########################
+
+# WORKS (NORMAL RESIDUALS)
+
+d$Yr = sample( mo$residuals, replace = TRUE )
+
[…]

### 2018-01-13 R code/For Sherlock/analysis.R

@@ -1,0 +1,229 @@
+
+########################### READ IN DATA ###########################
+
+# check overall progress
+#setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2/Simulation results/2017-7-10")
+setwd("~/Desktop")
+scen.params = read.csv("scen_params.csv")
+
+setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2/Simulation results/2017-7-10/sim_results/overall_stitched")
+s = read.csv("stitched_results.csv")
+
+# # which scenarios are done?
+# table(s$scen)
+
+
+# does it achieve nominal coverage?
+# should be 2.5% true
+#prop.table( table( s$n.rej > s$hi.bt ) )
+# should be 0.05
+mean(s$bt.rej)
+
+# THESE SHOULD BE EQUAL (since generated under H0)
+# real quantiles of n.rej
+quantile( s$n.rej, c(0.025, 0.975) )
+# vs. quantiles estimated by bootstraps
+c( mean(s$lo.bt), mean(s$hi.bt) )
+
+# these should also be equal
+quantile( s$n.rej, c(0.025, 0.975) )
+
+# compare variances
+var(s$n.rej); mean(s$var.n.rej.bt)
+
+# did bootstrap hypo test work?
+# invert-CI method
+prop.table( table( ( s$n.rej > mean(s$hi.bt) ) ) )
+
+# sanity check: should be 5%
+prop.table( table( ( s$n.rej > quantile(s$n.rej, 0.95) ) ) )
+
+
+
+
+
+
+
+
+
+# later get this from parameters
+alpha = scen.params$alpha
+nY = scen.params$nY
+N0 = alpha * nY
+
+
+# simulation rep
+n.boot = max(s$bt.iterate)
+n.sims.run = dim(s)[1] / n.boot
+s$sim.rep = rep( seq(1, n.sims.run), each = n.boot )
+
+# compute CI coverage
+library(data.table)
+st = data.table(s)
+st[ , lo.bt := quantile( n.rej.bt, alpha / 2 ), by=sim.rep ]
+st[ , hi.bt := quantile( n.rej.bt, 1 - (alpha / 2) ), by=sim.rep ]
+
+# coverage
+st$covers = (st$lo.bt <= N0 ) & (st$hi.bt >= N0)
+
+# hypothesis test
[…]

### 2018-01-13 R code/For Sherlock/doParallel.R

@@ -1,0 +1,181 @@
+
+########################### LOAD COMMAND-LINE ARGUMENTS  ###########################
+
+# FOR CLUSTER USE
+# load command line arguments
+args = commandArgs(trailingOnly = TRUE)
+jobname = args[1]
+scen = args[2]  # this will be a letter
+boot.reps = as.numeric( args[3] )
+
+# get scen parameters
+setwd("/share/PI/manishad/multTest")
+scen.params = read.csv( "scen_params.csv" )
+p = scen.params[ scen.params$scen.name == scen, ]
+
+# no longer included in parameters because it's a vector
+alpha = c(0.01, 0.05)
+
+# simulation reps to run within this job
+sim.reps = 10
+
+# FOR LOCAL USE
+# setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/R code/Sandbox/2017-7-21 Test smaller simulation")
+# p = read.csv("scen_params.csv")
+# p$bt.type = "resid"
+# boot.reps = 2000  
+# setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/R code/For Sherlock") # for sourcing functions later
+
+
+
+########################### THIS SCRIPT COMPLETELY RUNS 1 SIMULATION  ###########################
+
+# draws a new dataset
+# runs all 500 bootstrapping iterations
+# writes 1-row csv with number rejected and bootstrap CI
+
+library(doParallel)
+library(foreach)
+library(mvtnorm)
+
+source("functions.R")
+
+# set the number of cores
+registerDoParallel(cores=16)
+
+
+
+# j is the number of simulation iterations to run sequentially
+# so for j=10, we are generating 10 observed datasets, along with 500 bootstrap iterates for each
+
+for ( j in 1:sim.reps ) {
+  
+#   # ONLY FOR LOCAL USE
+#   # monitor progress
+#   cat( "\n\nStarting sim rep", j )
+#   
+#   # occasionally write intermediate results to desktop
+#    if ( j %% 50 == 0 ) {
+#      setwd("~/Desktop")
+#      write.csv(res, "temp_results.csv")
+#    }
+
+  rep.time = system.time( {
+    
+  # make initial dataset from which to bootstrap
[…]

### 2018-01-13 R code/For Sherlock/functions.R

@@ -1,0 +1,416 @@
+
+
+########################### FN: RETURN CORRELATION BETWEEN TWO ARBITRARY CELLS ###########################
+
+# eventually could make this an argument in the scen_params matrix somehow so that it can be varied
+# assume that X1 is the covariate of interest
+# assume that none of the other covariates are associated with any outcomes
+
+cell_corr = function( vname.1, vname.2, .rho.XX, .rho.YY, .rho.XY ) {
+  
+  # use grep to figure out if variables are covariates or outcomes
+  if ( length( grep("X", vname.1 ) ) == 1 ) {
+    vtype.1 = "covariate"
+  } else if ( length( grep("Y", vname.1 ) ) == 1 ) {
+    vtype.1 = "outcome"
+  }
+  
+  if ( length( grep("X", vname.2 ) ) == 1 ) {
+    vtype.2 = "covariate"
+  } else if ( length( grep("Y", vname.2 ) ) == 1 ) {
+    vtype.2 = "outcome"
+  }
+  
+  # case 1: diagonal entry
+  if ( vname.1 == vname.2 ) return(1)
+  
+  # case 2: both are covariates
+  # fixed correlation between covariates
+  if ( vtype.1 == "covariate" & vtype.2 == "covariate" ) return( .rho.XX )
+  
+  # case 3: both are outcomes
+  if ( vtype.1 == "outcome" & vtype.2 == "outcome" ) return( .rho.YY )
+  
+  # case 4: one is a covariate and one is an outcome
+  if ( vtype.1 != vtype.2 ) {
+    # check if this is the covariate of interest
+    # equal correlation between X1 and all outcomes
+    # all other covariates have correlation 0 with outcome
+    if( "X1" %in% c(vname.1, vname.2) ) return( .rho.XY )  # use 0 for null hypothesis
+    else return(0)
+  }
+  
+}
+
+# test drive
+# cell_corr( vname.1 = "X1", vname.2 = "Y3", .rho.XX = 0,
+#            .rho.YY = 0.04, .rho.X1.Y = 0.8, .rho.Xj.Y = .1 )
+
+
+########################### FN: CREATE CORRELATION MATRIX ###########################
+
+# .nX: number of covariates including the one of interest
+# .nY: number of outcomes
+
+make_corr_mat = function( .nX, .nY, .rho.XX, .rho.YY, .rho.XY ) {
+  
+  nVar = .nX + .nY
+  
[…]

### 2018-01-13 R code/For Sherlock/gen_sbatch.R

@@ -1,0 +1,127 @@
+
+
+########################### SET SIMULATION PARAMETERS MATRIX ###########################
+
+path = "/share/PI/manishad/multTest"
+setwd(path)
+
+n = 1000
+nX = 1
+nY = 100
+rho.XX = 0
+rho.YY = c(0, 0.25, 0.5, 0.75)
+rho.XY = 0.03  # null hypothesis: 0
+
+# bootstrap iterates and type
+boot.reps = 2000
+bt.type = "resid"
+
+
+# matrix of scenario parameters
+scen.params = expand.grid( bt.type, n, nX, nY, rho.XX, rho.YY, rho.XY )
+names(scen.params) = c( "bt.type", "n", "nX", "nY", "rho.XX", "rho.YY", "rho.XY" )
+
+# name the scenarios
+# remove letters that are privileged variables in R
+letter.names = c(letters, LETTERS)[ ! c(letters, LETTERS) %in% c("i","T","F") ]
+scen.params$scen.name = letter.names[ 1:dim(scen.params)[1] ]
+n.scen = length(scen.params[,1])
+
+# write the csv file of params (to Sherlock)
+write.csv( scen.params, "scen_params.csv" )
+
+
+########################### FN: SIMULATE 1 DATASET ###########################
+
+# load functions for generating sbatch files
+source("functions.R")
+
+
+# number of sbatches to generate (i.e., iterations within each scenario)
+n.reps.per.scen = 1000
+n.reps.in.doParallel = 10
+n.files = ( n.reps.per.scen / n.reps.in.doParallel ) * n.scen
+
+
+path = "/share/PI/manishad/multTest"
+
+scen.name = rep( scen.params$scen.name, each = ( n.files / n.scen ) )
+jobname = paste("job", 1:n.files, sep="_")
+outfile = paste("rm_", 1:n.files, ".out", sep="")
+errorfile = paste("rm_", 1:n.files, ".err", sep="")
+write_path = paste(path, "/sbatch_files/", 1:n.files, ".sbatch", sep="")
+runfile_path = paste(path, "/testRunFile.R", sep="")
+
+
+sbatch_params <- data.frame(jobname,
+                            outfile,
+                            errorfile,
+                            jobtime = "5:00:00",
+                            quality = "normal",
+                            node_number = 1,
+                            mem_per_node = 64000,
+                            mailtype =  "NONE",
+                            user_email = "mmathur@stanford.edu",
+                            tasks_per_node = 16,
[…]

### 2018-01-13 R code/For Sherlock/stitch_on_sherlock.R

@@ -1,0 +1,50 @@
+
+path = "/share/PI/manishad/multTest"
+setwd(path)
+
+source("functions.R")
+s = stitch_files(.results.singles.path = "/share/PI/manishad/multTest/sim_results",
+             .name.prefix = "results",
+             .stitch.file.name="stitched_raw.csv")
+
+# merge in scenario parameters
+scen.params = read.csv("scen_params.csv")
+names(s)[ names(s) == "scen" ] = "scen.name"
+s = merge( scen.params, s )
+
+# get parameters (ASSUMES ONLY 1 SCENARIO)
+alpha = scen.params$alpha
+nY = scen.params$nY
+N0 = alpha * nY
+
+# simulation rep
+n.boot = max(s$bt.iterate)
+n.sims.run = dim(s)[1] / n.boot
+s$sim.rep = rep( seq(1, n.sims.run), each = n.boot )
+
+# compute CI coverage
+library(data.table)
+st = data.table(s)
+st[ , lo.bt := quantile( n.rej.bt, alpha / 2 ), by=sim.rep ]
+st[ , hi.bt := quantile( n.rej.bt, 1 - (alpha / 2) ), by=sim.rep ]
+
+# coverage
+st$covers = (st$lo.bt <= N0 ) & (st$hi.bt >= N0)
+
+# hypothesis test
+st[ , crit.val := quantile( abs( n.rej.bt - n.rej ), 1 - alpha ), by=sim.rep ]
+st[ , bt.rej := abs( n.rej - N0 ) > crit.val, by = sim.rep ]
+
+# keep only 1 row per scenario (for first boot iterate)
+s2 = data.frame(st)
+s2 = s2[ s2$bt.iterate==1, ]
+
+# write results
+write.csv( s2, "stitched_prepped.csv" )
+
+# zip the entire results folder
+setwd("/share/PI/manishad/multTest")
+system( paste("zip -r sim_results.zip sim_results") )
+
+
+

### 2018-01-13 R code/Old versions/sherlock.R

@@ -1,0 +1,88 @@
+
+# goal: each time we have 16 workers, 
+
+
+######################### RUN IN PARALLEL #########################
+
+# load command line arguments
+args = commandArgs(trailingOnly = TRUE)
+scen = args[1]
+
+# get simulation parameters from csv file
+scen.params = read.csv("scen_params.csv")
+p = scen.params[ scen.params$scen.name == scen, ]
+
+# load required libraries
+require(plyr)
+require(doSNOW)
+require(car)
+
+#set up parallelization backend
+cl = makeCluster(n_clusters)
+registerDoSNOW(cl)
+
+#these should be equal
+n_clusters
+getDoParWorkers()
+
+#give each worker an ID 
+#this is optional, but occasionally is useful 
+#in situations where you need workers to write data
+#every worker writing to the same file will result in 
+#uncaught errors and partially overwritten data
+#worker ids enable each worker to write to its own set of 
+#files
+clusterApply(cl, seq(along=cl), function(id) WORKER.ID <<- paste0("worker_", id))
+
+
+time = system.time({
+  
+  l_ply( c( 1:getDoParWorkers() ), .parallel=TRUE, function(.item) {  
+    #l_ply iterates through each element of its first argument, here c(1:1000), and passes each element to
+    #function() as .item    
+    #instead of c(1:1000), you could pass l_ply() a list where each element is a set of subject specific means
+    #then each worker/iteration would expand those means into a full set of data for that subject
+    #avoid workers writing to the same file
+    #if each worker/iteration generates data for one subject, you will also need to write code that stitches
+    #all subject data together into one data file.
+    
+    #everything in here will be performed by workers in parallel    
+    #packages and source functions used by workers should be loaded within the this block
+    library(mvtnorm)
+    
+    setwd("/share/PI/manishad/multTest")
+    source("functions.R", local=TRUE)
+    
+    # DELETE THIS - testing only
+    #write.csv(.n.Subj, "n.Subj.csv")
+    #  write.csv(cat.parameters, "cat.parameters.csv")
+    
+    # simulate results
+    setwd("/share/PI/manishad/genCov/datasets")
+    
+    
+    sim_scenario( .scen=s, .reps=reps, .boot.reps = boot.reps )
+    rs$scen.name = s
+    
+    # merge in scenario parameters
+    rs2 = merge(p, rs)
+    
[…]

### 2018-01-13 R code/sim_study_local.R

@@ -1,0 +1,59 @@
+
+# remove all variables
+#rm(list=ls(all=TRUE))
+
+#################### SET SCENARIO PARAMETERS #################### 
+n = 50
+alpha = c( 0.05 )
+nX = 1
+nY = 5000
+rho.XX = 0.1
+rho.YY = c( 0.5 )
+rho.XY = 0  # null
+
+reps = 1
+boot.reps = 1
+
+# matrix of scenario parameters
+scen.params = expand.grid( n, alpha, nX, nY, rho.XX, rho.YY, rho.XY )
+names(scen.params) = c( "n", "alpha", "nX", "nY", "rho.XX", "rho.YY", "rho.XY" )
+
+# name the scenarios
+ scen.params$scen.name = c(letters, LETTERS)[ 1:dim(scen.params)[1] ]
+#scen.params$scen.name = c(letters, LETTERS)[ 7: ( dim(scen.params)[1] + 6) ]
+n.scen = length(scen.params[,1])
+
+
+
+#################### RUN SIMULATION #################### 
+setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2/R code/For Sherlock")
+source("functions.R")
+
+path = "~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2/Simulation results"
+setwd(path)
+
+# loop through the scenarios and run each one
+system.time(
+for (s in scen.params$scen.name[1]) {  # REMOVE THE 1
+    rs = sim_scenario( .scen=s, .reps=reps, .boot.reps = boot.reps)
+    rs$scen.name = s
+    rs2 = merge(scen.params, rs)
+    write.csv( rs2, paste(Sys.Date(), "_scen_", s, ".csv", sep="") )
+}
+)
+
+# stitch results files
+s = stitch_files(.results.singles.path = path,
+                      .name.prefix="_scen", .stitch.file.name="stitched_results.csv")
+
+# remove annoying rows
+
+# organize it and remove annoying rows
+s = s[ order(s$alpha, s$rho.YY), c( names(scen.params), "n.rejected", "lo.bt", "hi.bt" ) ]
+
+# higher correlation between Ys => wider CI
+# lower alpha => narrower CI
+
+#################### SUMMARIZE RESULTS ####################
+
+

### 2018-01-13 R code/For Sherlock/doParallel.R
Updating code to compare p-values from resampling under H0 vs. under original distribution

@@ -3,28 +3,28 @@
 
 # FOR CLUSTER USE
 # load command line arguments
-args = commandArgs(trailingOnly = TRUE)
-jobname = args[1]
-scen = args[2]  # this will be a letter
-boot.reps = as.numeric( args[3] )
-
-# get scen parameters
-setwd("/share/PI/manishad/multTest")
-scen.params = read.csv( "scen_params.csv" )
-p = scen.params[ scen.params$scen.name == scen, ]
-
-# no longer included in parameters because it's a vector
-alpha = c(0.01, 0.05)
-
-# simulation reps to run within this job
-sim.reps = 10
+# args = commandArgs(trailingOnly = TRUE)
+# jobname = args[1]
+# scen = args[2]  # this will be a letter
+# boot.reps = as.numeric( args[3] )
+# 
+# # get scen parameters
+# setwd("/share/PI/manishad/multTest")
+# scen.params = read.csv( "scen_params.csv" )
+# p = scen.params[ scen.params$scen.name == scen, ]
+# 
+# # no longer included in parameters because it's a vector
+# alpha = c(0.01, 0.05)
+# 
+# # simulation reps to run within this job
+# sim.reps = 10
 
 # FOR LOCAL USE
-# setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/R code/Sandbox/2017-7-21 Test smaller simulation")
-# p = read.csv("scen_params.csv")
-# p$bt.type = "resid"
-# boot.reps = 2000  
-# setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/R code/For Sherlock") # for sourcing functions later
+setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/Sandbox/2018-1-13")
+p = read.csv("scen_params.csv")
+p$bt.type = "resid"
+boot.reps = 2000
+setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/R code/For Sherlock") # for sourcing functions later
 
 
 

### 2018-01-13 R code/For Sherlock/gen_sbatch.R

@@ -2,10 +2,23 @@
 
 ########################### SET SIMULATION PARAMETERS MATRIX ###########################
 
+# FOR CLUSTER USE
 path = "/share/PI/manishad/multTest"
+
+# FOR LOCAL USE
+path = "~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/Sandbox/2018-1-13"
+
 setwd(path)
 
-n = 1000
+# from previous simulation
+# n = 1000
+# nX = 1
+# nY = 100
+# rho.XX = 0
+# rho.YY = c(0, 0.25, 0.5, 0.75)
+# rho.XY = 0.03  # null hypothesis: 0
+
+n = 50
 nX = 1
 nY = 100
 rho.XX = 0
@@ -13,8 +26,9 @@ rho.YY = c(0, 0.25, 0.5, 0.75)
 rho.XY = 0.03  # null hypothesis: 0
 
 # bootstrap iterates and type
-boot.reps = 2000
-bt.type = "resid"
+# boot.reps = 2000
+boot.reps = 5
+bt.type = c( "fcr", "resid" )  # fcr: resample under HA; resid: resample under H0
 
 
 # matrix of scenario parameters

### 2018-01-14 R code/For Sherlock/doParallel.R
Versions used for generation on 2018-1-14

@@ -3,29 +3,39 @@
 
 # FOR CLUSTER USE
 # load command line arguments
-# args = commandArgs(trailingOnly = TRUE)
-# jobname = args[1]
-# scen = args[2]  # this will be a letter
-# boot.reps = as.numeric( args[3] )
-# 
-# # get scen parameters
-# setwd("/share/PI/manishad/multTest")
-# scen.params = read.csv( "scen_params.csv" )
-# p = scen.params[ scen.params$scen.name == scen, ]
-# 
-# # no longer included in parameters because it's a vector
-# alpha = c(0.01, 0.05)
-# 
-# # simulation reps to run within this job
-# sim.reps = 10
+args = commandArgs(trailingOnly = TRUE)
+jobname = args[1]
+scen = args[2]  # this will be a letter
+boot.reps = as.numeric( args[3] )
 
-# FOR LOCAL USE
-setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/Sandbox/2018-1-13")
-p = read.csv("scen_params.csv")
-p$bt.type = "resid"
-boot.reps = 2000
-setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/R code/For Sherlock") # for sourcing functions later
+# get scen parameters
+setwd("/share/PI/manishad/multTest")
+scen.params = read.csv( "scen_params.csv" )
+p = scen.params[ scen.params$scen.name == scen, ]
 
+# no longer included in parameters because it's a vector
+crit.bonf = 0.05 / p$nY
+alpha = c( crit.bonf, 0.01, 0.05 )
+
+# simulation reps to run within this job
+sim.reps = 10
+
+# ######### FOR LOCAL USE ######### 
+# setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/Sandbox/2018-1-13")
+# p = read.csv("scen_params.csv")  # should be a single row, I think
+# p$bt.type = "resid"
+# #boot.reps = 2000
+# boot.reps = 100
+# sim.reps = 5
+# 
+# # add alpha corresponding to Bonferroni as first one
+# # NOTE THAT CODE ASSUMES BONFERRONI IS THE FIRST ONE
+# crit.bonf = 0.05 / p$nY
+# alpha = c( crit.bonf, 0.01, 0.05 )
+# 
+# # for sourcing functions later
[…]

### 2018-01-14 R code/For Sherlock/functions.R

@@ -145,7 +145,7 @@ fit_model = function( Y.name, .dat, .resid ) {
 
 ########################### FN: GIVEN DATASET, RETURN STATS ###########################
 
-# residualize: should it return dataset with residuals (for resampling under null)?
+# .resid: should it return dataset with residuals (for resampling under null)?
 
 dataset_result = function( .dat, .alpha, .resid = FALSE ) {
 
@@ -154,7 +154,7 @@ dataset_result = function( .dat, .alpha, .resid = FALSE ) {
   Y.names = names( .dat )[ grep( "Y", names(.dat) ) ]
   
   # for each outcome, fit regression model
-  # see if each has p < 0.05 for covariate of interest
+  # see if each has p < alpha for covariate of interest
 
   # this is a list of lists:
   #  length is equal to number of outcomes

### 2018-01-14 R code/For Sherlock/gen_sbatch.R

@@ -4,11 +4,11 @@
 
 # FOR CLUSTER USE
 path = "/share/PI/manishad/multTest"
+setwd(path)
 
 # FOR LOCAL USE
-path = "~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/Sandbox/2018-1-13"
-
-setwd(path)
+# path = "~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/Sandbox/2018-1-13"
+# setwd(path)
 
 # from previous simulation
 # n = 1000
@@ -18,17 +18,16 @@ setwd(path)
 # rho.YY = c(0, 0.25, 0.5, 0.75)
 # rho.XY = 0.03  # null hypothesis: 0
 
-n = 50
+n = 1000
 nX = 1
 nY = 100
 rho.XX = 0
 rho.YY = c(0, 0.25, 0.5, 0.75)
-rho.XY = 0.03  # null hypothesis: 0
+rho.XY = c(0.08, 0.03, 0)  # null hypothesis: 0
 
 # bootstrap iterates and type
-# boot.reps = 2000
-boot.reps = 5
-bt.type = c( "fcr", "resid" )  # fcr: resample under HA; resid: resample under H0
+boot.reps = 2000
+bt.type = c( "resid", "fcr" )  # fcr: resample under HA; resid: resample under H0
 
 
 # matrix of scenario parameters
@@ -88,9 +87,9 @@ sbatch_params <- data.frame(jobname,
 
 
 # run them all
-#works
+# works
 setwd( paste(path, "/sbatch_files", sep="") )
-for (i in 1:n.files) {
+for (i in 2:n.files) {
   system( paste("sbatch -p normal,owners /share/PI/manishad/multTest/sbatch_files/", i, ".sbatch", sep="") )
 }
 

### 2018-01-14 R code/For Sherlock/analysis.R
Work on data wrangling using test data

@@ -1,6 +1,63 @@
 
 ########################### READ IN DATA ###########################
 
+# LOCAL TEST
+setwd("~/Desktop")
+scen.params = read.csv("scen_params.csv")
+
+# just use a single file instead of stitched one
+s = read.csv("short_results_1_.csv", header=TRUE)
+
+# remove columns from individual bootstrap results
+kill = c( names(s)[ grepl( "n.rej.bt", names(s) ) ], "n.rej.5e.04.1" )
+s = s[ , !names(s) %in% kill ]
+
+
+# TEMP: see if reshape problem fixed by random noise
+#s[,11:30] = s[,11:30] + rnorm(n=100, mean=0, sd=0.01)
+temp = s[,c(10,14,15,16)]
+
+
+library(tidyr)
+
+# for use with separate function
+names(temp) = sub( pattern = "t[.]", replacement = "t_", x=names(temp) )
+
+# ~~~~~ HARD-CODES BONFERRONI!!!!
+options(scipen=999)
+bonf = 0.05/100
+names(temp) = sub( pattern = "bonf", replacement = bonf, x=names(temp) )
+
+gathered.messy <- gather(temp, key, value, -scen)
+head(gathered.messy)
+
+# BOOKMARK: STRUGGLING WITH THIS! IT CUTS OFF THE ALPHA LEVEL DIGITS
+tidy = separate( gathered.messy, key, into = c("stat", "alpha"), sep = "_", convert=FALSE )
+
+# http://www.milanor.net/blog/reshape-data-r-tidyr-vs-reshape2/
+
+
+
+# TAKE MEANS BY SCENARIO AND BY ALPHA
+# ALSO EMPIRICAL QUANTILES
+# SO, FOR TEST CASE, WILL BE JUST 3 ROWS
+
+
+# add expected number of hits
+agg$expect = agg$alpha * agg$nY
+
+# for plotting
+agg$group = paste( "nX: ", agg$nX, "; nY: ", agg$nY, sep="" )
+
+
+
+
+
+##### ~~~~~~~~~~~ FROM BEFORE
+
+
+########################### READ IN DATA ###########################
+
 # check overall progress
 #setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2/Simulation results/2017-7-10")
 setwd("~/Desktop")
@@ -87,46 +144,6 @@ mean(s$bt.rej)
 # make simulation rep variable
 s$sim.rep = 1:dim(s)[1]
 
[…]

### 2018-01-14 R code/For Sherlock/analysis.R
Updates to analysis code

@@ -4,37 +4,73 @@
 # LOCAL TEST
 setwd("~/Desktop")
 scen.params = read.csv("scen_params.csv")
+names(scen.params)[ names(scen.params) == "scen.name" ] = "scen"
 
 # just use a single file instead of stitched one
 s = read.csv("short_results_1_.csv", header=TRUE)
 
 # remove columns from individual bootstrap results
-kill = c( names(s)[ grepl( "n.rej.bt", names(s) ) ], "n.rej.5e.04.1" )
+kill = c( names(s)[ grepl( "n.rej.bt", names(s) ) ], "n.rej.5e.04" )
 s = s[ , !names(s) %in% kill ]
+
+# did naive Bonferroni reject?
+s$naive_bonf_rej = s$n.rej.5e.04.1 > 0
 
 
 # TEMP: see if reshape problem fixed by random noise
 #s[,11:30] = s[,11:30] + rnorm(n=100, mean=0, sd=0.01)
-temp = s[,c(10,14,15,16)]
+#temp = s[,c(10,14,15,16)]
 
 
+########################### MAKE DATASETS FOR FIRST PLOT (RESIDUAL RESAMPLING) ###########################
+
+# http://www.milanor.net/blog/reshape-data-r-tidyr-vs-reshape2/
 library(tidyr)
+library(dplyr)
+# replace underscores (for use with separate function)
+#names(temp) = sub( pattern = "t[.]", replacement = "t_", x=names(temp) )
+names(s)[ grepl( "crit", names(s) ) ] = sub( pattern = "t[.]",
+                                             replacement = "t_", x=names(s)[ grepl( "crit", names(s) ) ] )
+names(s)[ grepl( "rej.jt", names(s) ) ] = sub( pattern = "t[.]",
+                                             replacement = "t_", x=names(s)[ grepl( "rej.jt", names(s) ) ] )
 
-# for use with separate function
-names(temp) = sub( pattern = "t[.]", replacement = "t_", x=names(temp) )
+
+######## Aggregated Dataset 1: Our Method's Power ######## 
+# only keep residual resamples and useful columns
+crit.names = names(s)[ grepl( "crit", names(s) ) ]
+rej.jt.names = names(s)[ grepl( "rej.jt", names(s) ) ]
+temp = s[ s$bt.type == "resid", names(s) %in% c( "scen", "rho.YY", "rho.XY", crit.names, rej.jt.names ) ] 
 
[…]

### 2018-01-14 Project log/multiple_outcomes_project_log.Rmd
Simplify stitching

@@
 \section{Stream of consciousness}
+
+2018-1-15
+Mysteries: 
+* Understand why FCR resampling doesn't work (these resamples reject more often than the original data).
+I found that prob of rejection even when there is only 1 covariate is too high.
+In my simple script, try resampling residuals and re-attaching them to fitted values.
+Does that work?
+* Is it theoretically valid to use the weird CI for excess hits using the observed hits and the CI under the null?
 
@@
 
+\subsection{2017-8-1 (generate under $H_0$ and $H_A$)}
+
+\begin{itemize}
+\item FCR resampling doesn't seem to work because even when data generated under the null, the resamples have more than the expected number of rejections
+\item Not due to lower truncation because using alpha = 0.5 doesn't help
+\item Seems like correlations are preserved in individual resamples?
+\item But I think the correlations are more variable, even if their mean is still 0, because sometimes we draw samples that have a lot of repetition
+\end{itemize}
+
 \subsection{Summary of previous two}
@@
 
-\item{ Bootstrap from original sample with 100 independent std. normal; reject using z-test that’s only a function of mean $\Rightarrow$ \textbf{DOESN'T WORK} 
+\item{ Bootstrap from original sample with 100 independent std. normal; reject using z-test that’s only a function of mean because SD is treated as known $\Rightarrow$ \textbf{DOESN'T WORK} 
   \begin{itemize}

### 2018-01-14 R code/For Sherlock/analysis.R

@@ -7,7 +7,13 @@ scen.params = read.csv("scen_params.csv")
 names(scen.params)[ names(scen.params) == "scen.name" ] = "scen"
 
 # just use a single file instead of stitched one
-s = read.csv("short_results_1_.csv", header=TRUE)
+s = read.csv("stitched.csv", header=TRUE)
+
+# what percent done are we?
+nrow(s) / (1000 * nrow(scen.params))
+
+# how many reps per scenario do we have?
+table( s$scen )
 
 # remove columns from individual bootstrap results
 kill = c( names(s)[ grepl( "n.rej.bt", names(s) ) ], "n.rej.5e.04" )
@@ -36,26 +42,35 @@ names(s)[ grepl( "rej.jt", names(s) ) ] = sub( pattern = "t[.]",
 
 
 ######## Aggregated Dataset 1: Our Method's Power ######## 
-# only keep residual resamples and useful columns
+
+# only keep useful columns
 crit.names = names(s)[ grepl( "crit", names(s) ) ]
 rej.jt.names = names(s)[ grepl( "rej.jt", names(s) ) ]
-temp = s[ s$bt.type == "resid", names(s) %in% c( "scen", "rho.YY", "rho.XY", crit.names, rej.jt.names ) ] 
+temp = s[ , names(s) %in% c( "scen", "rho.YY", "rho.XY", crit.names, rej.jt.names ) ] 
 
 # ~~~~~ HARD-CODES BONFERRONI!!!!
 options(scipen=999)
 bonf = 0.05/100
 names(temp) = sub( pattern = "bonf", replacement = bonf, x=names(temp) )
+names(temp) = sub( pattern = "5e.04.1", replacement = bonf, x=names(temp) )
 
 # get in tidy format
 gathered.messy = gather(temp, key, value, -scen, -rho.YY, -rho.XY )
 head(gathered.messy)
 
 tidy = separate( gathered.messy, key, into = c("stat", "alpha"), sep = "_", convert=FALSE )
+tidy = merge( scen.params, tidy )
 
-# make summary datasets
-power = tidy %>% group_by( scen, alpha ) %>% filter( stat == "rej.jt" ) %>% summarise( power = mean(value) )
-power = merge( scen.params, power )
+# make summary dataset
+power = tidy %>% group_by( scen, alpha, rho.XY, rho.YY ) %>% filter( stat == "rej.jt" & bt.type == "resid" ) %>% summarise( power = mean(value) )
[…]

### 2018-01-14 R code/For Sherlock/doParallel.R

@@ -20,22 +20,39 @@ alpha = c( crit.bonf, 0.01, 0.05 )
 # simulation reps to run within this job
 sim.reps = 10
 
-# ######### FOR LOCAL USE ######### 
+######### FOR LOCAL USE #########
 # setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/Sandbox/2018-1-13")
 # p = read.csv("scen_params.csv")  # should be a single row, I think
-# p$bt.type = "resid"
-# #boot.reps = 2000
-# boot.reps = 100
-# sim.reps = 5
-# 
-# # add alpha corresponding to Bonferroni as first one
-# # NOTE THAT CODE ASSUMES BONFERRONI IS THE FIRST ONE
-# crit.bonf = 0.05 / p$nY
-# alpha = c( crit.bonf, 0.01, 0.05 )
-# 
-# # for sourcing functions later
-# setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/git_multiple_outcomes/R code/For Sherlock")
-# ######### END OF LOCAL PART ######### 
+n = 1000
+nX = 1
+nY = 100
+rho.XX = 0
+rho.YY = c(0)
+rho.XY = c(0)  # null hypothesis: 0
+
+# bootstrap iterates and type
+boot.reps = 200
+bt.type = c( "fcr" )  # fcr: resample under HA; resid: resample under H0
+
+
+# matrix of scenario parameters
+scen.params = expand.grid( bt.type, n, nX, nY, rho.XX, rho.YY, rho.XY )
+names(scen.params) = c( "bt.type", "n", "nX", "nY", "rho.XX", "rho.YY", "rho.XY" )
+
+# name the scenarios
+# remove letters that are privileged variables in R
+letter.names = c(letters, LETTERS)[ ! c(letters, LETTERS) %in% c("i","T","F") ]
+scen.params$scen.name = letter.names[ 1:dim(scen.params)[1] ]
+p = scen.params
+
+# add alpha corresponding to Bonferroni as first one
+# NOTE THAT CODE ASSUMES BONFERRONI IS THE FIRST ONE
+crit.bonf = 0.05 / p$nY
+alpha = c( crit.bonf, 0.01, 0.05 )
+
+# for sourcing functions later
[…]

### 2018-01-14 R code/For Sherlock/stitch_on_sherlock.R

@@ -1,50 +1,21 @@
 
+# run this interactively in ml load R
+
 path = "/share/PI/manishad/multTest"
 setwd(path)
 
 source("functions.R")
-s = stitch_files(.results.singles.path = "/share/PI/manishad/multTest/sim_results",
-             .name.prefix = "results",
-             .stitch.file.name="stitched_raw.csv")
+s = stitch_files(.results.singles.path = "/share/PI/manishad/multTest/sim_results/short",
+                 .results.stitched.write.path = "/share/PI/manishad/multTest/sim_results",
+             .name.prefix = "short",
+             .stitch.file.name="stitched.csv")
 
-# merge in scenario parameters
-scen.params = read.csv("scen_params.csv")
-names(s)[ names(s) == "scen" ] = "scen.name"
-s = merge( scen.params, s )
-
-# get parameters (ASSUMES ONLY 1 SCENARIO)
-alpha = scen.params$alpha
-nY = scen.params$nY
-N0 = alpha * nY
-
-# simulation rep
-n.boot = max(s$bt.iterate)
-n.sims.run = dim(s)[1] / n.boot
-s$sim.rep = rep( seq(1, n.sims.run), each = n.boot )
-
-# compute CI coverage
-library(data.table)
-st = data.table(s)
-st[ , lo.bt := quantile( n.rej.bt, alpha / 2 ), by=sim.rep ]
-st[ , hi.bt := quantile( n.rej.bt, 1 - (alpha / 2) ), by=sim.rep ]
-
-# coverage
-st$covers = (st$lo.bt <= N0 ) & (st$hi.bt >= N0)
-
-# hypothesis test
-st[ , crit.val := quantile( abs( n.rej.bt - n.rej ), 1 - alpha ), by=sim.rep ]
-st[ , bt.rej := abs( n.rej - N0 ) > crit.val, by = sim.rep ]
-
-# keep only 1 row per scenario (for first boot iterate)
-s2 = data.frame(st)
-s2 = s2[ s2$bt.iterate==1, ]
-
-# write results
-write.csv( s2, "stitched_prepped.csv" )
-
-# zip the entire results folder
-setwd("/share/PI/manishad/multTest")
-system( paste("zip -r sim_results.zip sim_results") )
-
+# # write results
+# write.csv( s, "stitched_prepped.csv" )
+# 
+# # zip the entire results folder
+# setwd("/share/PI/manishad/multTest")
+# system( paste("zip -r sim_results.zip sim_results") )
+# 
 
 

### 2018-03-01 Manuscript/manuscript.Rmd
Try to fix git syncing

---
title: ''
author: ''
header-includes:
- \usepackage[T1]{fontenc}
- \usepackage{microtype}
- \usepackage[margin=1in]{geometry}
- \usepackage{fancyhdr}
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead{}
- \fancyfoot{}
- \fancyhead[C]{Multiple Outcomes}
- \fancyfoot[RO,LE]{\thepage}
- \usepackage{booktabs}
- \usepackage{lettrine}
- \usepackage{paralist}
- \usepackage{setspace}\singlespacing
- \usepackage{url}
- \usepackage{parskip}
- \usepackage{color,soul}
- \usepackage{palatino}
- \usepackage{amsmath,amsfonts,amsthm, textcomp}
- \usepackage{bm}
- \usepackage{caption}
- \usepackage{subcaption}
- \usepackage{tikz}
- \usepackage{verbatim}
- \usetikzlibrary{matrix}
- \usepackage{float}
output: pdf_document
bibliography: refs_fake.bib
csl: aje_style.csl
---

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\newtheorem{remark}{Remark}
\newtheorem{fact}{Fact}
\newtheorem{example}{Example}
\newtheorem{algorithm}{Algorithm}

<!--
\doublespacing

\begin{center}
\textbf{ \LARGE{Title line one:} } \\ \vspace{5mm}
\textbf{ \LARGE{title line two} }
\vspace{10mm}
\end{center}

\doublespacing

\vspace{10mm}
\begin{center}
\large{ \emph{ Maya B. Mathur$^{1, 2\ast}$, Peng Ding$^{3}$, and Tyler J. VanderWeele$^{1,4}$ } }
\end{center}

\vspace{15mm}

\small{$^{1}$ Department of Biostatistics, Harvard T. H. Chan School of Public Health, Boston, MA, USA}

\small{$^{2}$Quantitative Sciences Unit, Stanford University, Palo Alto, CA, USA}

\small{$^{3}$Department of Statistics, University of California at Berkeley, Berkeley, CA, USA}

\small{$^{4}$Department of Epidemiology, Harvard T. H. Chan School of Public Health, Boston, MA, USA}

\vspace{15mm}
\begin{singlespacing} 
\small{$\ast$: Corresponding author:

mmathur@stanford.edu

Quantitative Sciences Unit (c/o Inna Sayfer)

1070 Arastradero Road

Palo Alto, CA

94305

}
\end{singlespacing}

\vspace{10mm}

\newpage
-->

\section{Notes to self}

<!-------------------------- --------------------------------->
<!-------------- SECTION: SETTING AND NOTATION --------------->
<!-------------------------- --------------------------------->

\section{Setting and notation}

Suppose that $K$ random variables are measured on $N$ subjects, with the resulting matrix denoted $\mathbf{Z} \in \mathbb{R}_{N \times K}$. Let $Z_{nk}$ denote, for the $n^{th}$ subject, the $k^{th}$ random variable. Consider a bootstrapping algorithm (which is defined below) that generates, for iterate $j$, a dataset containing the random variables $\left( Z_{n1}^{(j)}, \dotsb, Z_{nK}^{(j)} \right)$ for each subject $n$. There are a total of $B$ bootstrapped datasets.

We refer to probability density functions as $f$ and cumulative distribution functions as $F$. Specifically, we refer to the joint distribution of the data as follows. With $\mathcal{B} \in \mathbb{R}_{N \times K}$ denoting a matrix of constants with entries $\left( b_{11}, \dotsb, b_{NK} \right)$, let:
\begin{align*}
F_{\mathbf{Z}} \left( \mathbf{B} \right) &= P \left( Z_{11} \le b_{11}, \dotsb, Z_{NK} \le b_{NK} \right)
\end{align*}

<!-- Define the joint CDF of $\left( Z_{n1}^{(j)}, \dotsb, Z_{nK}^{(j)} \right)$ as:

\begin{align*}
F_{\left( Z_{n1}^{(j)}, \dotsb, Z_{nK}^{(j)} \right)} \left( r_1, \dotsb, r_K \right) &= P \left( Z_{n1}^{(j)} < r_1, \dotsb, Z_{nK}^{(j)} < r_K \right)
\end{align*}
-->

We will usually refer simply to $\mathbf{Z}$ and $F_{\mathbf{Z}}$ for brevity and will denote their counterparts in bootstrapped dataset $j$ as $\mathbf{Z}^{(j)}$ and $F_{\mathbf{Z}^{(j)}}$.

Further suppose that we conduct $W$ hypothesis tests, each with level $\alpha$. Let $c_{w,\alpha}$ be the critical value for the test statistic, $T_w$, of the $w^{th}$ test. We refer to the vector of test statistics as $\mathbf{T}$. Assume that each test concerns a point null hypothesis (\textbf{Assumption 1}), that each test statistic is a continuous function of the data with continuous first derivatives, $\mathbf{Z}$ (\textbf{Assumption 2}), and that each test is parametric with consistent estimates of all parameters in the likelihood identifiable from the observed data (\textbf{Assumption 3}). We define the "strong null" as the case in which all $W$ point null hypotheses hold and use the superscript $^0$ to denote variables generated under the strong null.

Define the statistic corresponding to the total number of rejections in $W$ tests in a sample generated under the strong null as $\widehat{\theta}^0 = \sum_{w=1}^W 1 \big\{ T_w^0 > c_{w,\alpha} \big\}$. Similarly, the total number of rejections in bootstrap sample $j$ is $\widehat{\theta}^{(j)} = \sum_{w=1}^W 1 \big\{ T_w^{(j)} > c_{w,\alpha} \big\}$.

Respectively define the true distribution of the number of rejections under the strong null, its counterpart in the bootstrap samples, and its empirical estimator in the bootstrap samples as:
\begin{align*}
F_{\widehat{\theta}^0}(r) &= P \left( \widehat{\theta}^0 \le r \right) \\
F_{\widehat{\theta}^{(j)}}(r) &= P \left( \widehat{\theta}^{(j)} \le r \right) \\
\widehat{F}_{\widehat{\theta}^{(j)}} \left( r \right) &= \frac{1}{B} \sum_{j^{*}=1}^B 1 \Big\{ \widehat{\theta}^{(j^{*})} \le r \Big\}
\end{align*}

<!------------- --------------------------------->
<!-------------- SECTION: RESULTS --------------->
<!------------- --------------------------------->

\section{Results}

We now show that under a parametric bootstrapping algorithm defined below, the empirical distribution of the number of rejections in the bootstrap samples converges in distribution to the true distribution of the number of rejections in sample generated under the strong null. Figure \ref{roadmap} depicts the structure of the proofs.

\begin{figure}[H]
\caption{Overview of theory to follow. The dashed arrows denote convergence in probability in either $N$ or both $N$ and $B$. The solid arrow denotes almost sure convergence in $B$. Parenthetical acronyms refer to remarks, a lemma, and a theorem developed below.}
\label{roadmap}
\begin{center}
\begin{tikzpicture}
  \matrix (m) [matrix of math nodes,row sep=3em,column sep=4em,minimum width=2em]
  {
     F_{ \mathbf{T}^{(j)} } & F_{ \mathbf{T}^0 } \\
      %& F_{ \mathbf{Z} }
      \\};
  \path[-stealth]
  % the - argument removes the arrowhead
    (m-1-1) edge[dashed] node [below] {(R1)} (m-1-2)

    % diagonal line
	%(m-1-1) edge[dashed] node [below] {(C1)} (m-2-2)

    %(m-1-2) edge[dotted] node [right] {(L2)} (m-2-2)
            ;
\end{tikzpicture} % pic 1
\qquad
\begin{tikzpicture}
  \matrix (m) [matrix of math nodes,row sep=3em,column sep=4em,minimum width=2em]
  {
     F_{ \widehat{\theta}^{(j)} } & F_{ \widehat{\theta}^0 }  \\
      \widehat{F}_{ \widehat{\theta}^{(j)} } & \\};
  \path[-stealth]
  % the - argument removes the arrowhead
    (m-1-1) edge[dashed] node [below] {(L1)} (m-1-2)

    % diagonal line
	(m-2-1) edge[] node [left] {(R2)} (m-1-1)

    (m-2-1) edge [dashed] node [below] {(T1)} (m-1-2)
            ;
\end{tikzpicture}
\end{center}
\end{figure}

<!-------------- Definition 1 --------------->
\bigskip \begin{algorithm}[Parametric bootstrap under the strong null] 
\label{cbt_def}
We will refer throughout to bootstrap samples generated as follows:
\begin{enumerate}
\item Use the observed data to obtain consistent point estimates of any unknown parameters in the likelihood. 
\item For any parameters in the likelihood that are constrained under the null hypothesis, set them equal to their null values (which are uniquely defined under Assumption 1 regarding point null hypotheses). For nuisance parameters, set them equal to their consistent estimates.  
\item For any covariates treated as fixed in the likelihood, set them equal to their observed values.
\item Generate outcome data from the estimated likelihood under the strong null as defined above. 
\end{enumerate}
\end{algorithm}

(\hl{CITE ME - e.g., Fox, Romano}.)
Note that unlike some resampling schemes that can induce \hl{XXX} problems, this algorithm guarantees that the bootstrapped data satisfy all parametric assumptions of the hypothesis tests to be conducted, preserving the nominal $\alpha$ of each test.

<!-------------- Example 1 --------------->
\bigskip
\begin{example}[Parametric bootstrap for multiple linear regression] 
Suppose that one of the hypothesis tests is conducted using an OLS model regressing the $w^{th}$ variable on the first variable, adjusting for the second variable:
$$Z_{nw} = \beta_0 + \beta_1 Z_{n1} + \beta_2 Z_{n2} + \epsilon_{nw}$$
with $H_0: \beta_1 = 0$ and $\epsilon \sim N(0, \sigma^2)$. Then we can apply Algorithm \ref{cbt_def} as follows:
\begin{enumerate}
\item Fit the model to the observed data to estimate the three nuisance parameters as $\widehat{\beta}_0$, $\widehat{\beta}_2$, and $\widehat{\sigma}^2$. 
\item For each bootstrap iterate, set the design matrix equal to that of the observed data.
\item Set $\beta_1 = 0$ (invoking the strong null), and set nuisance parameters $\beta_0 = \widehat{\beta}_0$, $\beta_2 = \widehat{\beta}_2$, and $\sigma^2 = \widehat{\sigma}^2$. 
\item Thus, for each bootstrap iterate, generate the bootstrapped outcomes as $Z_{nw}^{(j)} = \widehat{\beta}_0 + \widehat{\beta}_2 Z_{n2} + \epsilon^{(j)}_{nw}$, where $\epsilon^{(j)}_{nw} \sim N(0,\widehat{\sigma}^2)$. 
\end{enumerate}
\end{example}

<!-------------- Example 2 --------------->
\bigskip
\begin{example}[Parametric bootstrap for linear mixed model] 
The approach in Definition \ref{cbt_def} can accommodate parametric models with correlated observations, such as a linear mixed model with random intercepts by subject. One could obtain consistent estimates of the random intercepts using, for example, empirical Bayesian estimation, and these estimates can be plugged into the estimated likelihood under the strong null. 
\end{example}

<!-------------- Remark 1 --------------->
\bigskip
\begin{remark}
\label{CMT_results}
The distribution of the bootstrapped test statistics is consistent in $N$ for the true distribution of the test statistics under the strong null. By repeatedly invoking the Continuous Mapping Theorem (CMT):
\begin{align}
\label{consistent}
f_{\mathbf{Z}^{(j)}} &\xrightarrow[N \to \infty]{P} f_{\mathbf{Z}^0} \tag{CMT; Assumption 3} \\
f_{\mathbf{T}^{(j)}} &\xrightarrow[N \to \infty]{P} f_{\mathbf{T}^0} \tag{CMT; Assumption 2} \\
F_{\mathbf{T}^{(j)}} &\xrightarrow[N \to \infty]{P} F_{\mathbf{T}^0} \tag{CMT; integration preserves continuity} \\
\end{align}
\end{remark}

<!-------------- Remark 2 --------------->
\bigskip \begin{remark}
\label{glivenko}
By the i.i.d construction of the bootstrap samples and the Glivenko-Cantelli theorem, the empirical distribution of the number of rejections in the bootstrap samples converges almost surely in $B$ to its true distribution under the strong null:
$$\widehat{F}_{\widehat{\theta}^{(j)}} \xrightarrow[B \to \infty]{A.S.} F_{\widehat{\theta}^{(j)}} $$
\end{remark}

<!-------------- Fact 2 --------------->
<!-- has been replaced by Glivenko-Cantelli -->
<!--
\bigskip \begin{fact}
\label{glivenko}

The empirical distribution of the number of rejections in the bootstrap samples converges almost surely in $B$ to its true distribution under the strong null:
$$\widehat{F}_{\widehat{\theta}^{(j)}} \xrightarrow[B \to \infty]{A.S.} F_{\widehat{\theta}^{(j)}} $$
\begin{proof}
Since the $B$ bootstrap samples are i.i.d. and $E \big[ \vert \widehat{F}_{\widehat{\theta}^{(j)}} \left( r \right) \vert \big] < \infty$, this follows directly from the SLLN:
%--
\begin{align*}
\widehat{F}_{\widehat{\theta}^{(j)}} \left( r \right) &= \frac{1}{B} \sum_{j^{*}=1}^B 1 \Big\{ \widehat{\theta}^{(j^{*})} \le r \Big\} \\
& \xrightarrow[N \to \infty]{A.S.} E \Big[ 1 \big\{ \widehat{\theta}^{(j)} \le r \big\} \Big] \\
&= P \left( \widehat{\theta}^{(j)} \le r \right) \\
&= F_{\widehat{\theta}^{(j)}}(r)
\end{align*}
\end{proof}
\hl{NOTE TO SELF: For paper, should just invoke Glivenko-Cantelli.}
\end{fact}
-->

<!-------------- Lemma 1 --------------->
\bigskip \begin{lemma}
\label{lemma3}
The number of rejections in the bootstrap samples converges in distribution in $N$ to the number of rejections in samples from the original population under the strong null: $$F_{{ \widehat{\theta} }^{(j)}} \xrightarrow[N \to \infty]{P} F_{{ \widehat{\theta} }^0}$$

\begin{proof}
Define the $r$-family of ``rejection sets'' as all possible configurations of the $W$ test statistics that lead to $r$ rejections: $$\mathcal{A}_r = \Big\{ \left( A_1, \dotsb, A_W \right) \in \mathbb{R}^W : \left( T_1 \in A_1, \dotsb, T_W \in A_W \right) \; \Rightarrow \; \widehat{\theta} = r\Big\}$$
(For example, $\mathcal{A}_5$ contains all the 5-vectors of sets such that, if $\left( T_1 \in A_1, \dotsb, T_5 \in A_5 \right)$, then there are 5 total rejections across the $W$ tests. Suppose, for example, that $T_w \in (-\infty, \infty) \; \forall \; w$ and that each null hypothesis is rejected when $T_w > c_{\alpha}$. Then there are ${W \choose r}$ vectors in $\mathcal{A}_5$, representing the number of ways to select $r$ rejections from $W$ tests. One such vector corresponds to rejecting the first 5 tests and failing to reject all remaining tests:\\ $\left( (c_\alpha, \infty), (c_\alpha, \infty), (c_\alpha, \infty), (c_\alpha, \infty), (c_\alpha, \infty), (-\infty, c_\alpha], \dotsb, (-\infty, c_\alpha] \right)$.)

\bigskip Now consider the distribution of $\widehat{\theta}^{(j)}$:
\begin{align*}
P \left( \widehat{\theta}^{(j)} = r \right) &= P \left( \sum_{w=1}^W 1 \Big\{ T_w^{(j)} > c_{w,\alpha} \Big\} = r \right) \\
&= \sum_{ \left( A_1, \dotsb, A_W \right) \in \mathcal{A} } P \left( T_1^{(j)} \in A_1, \dotsb, T_W^{(j)} \in A_W \right) \\
&\xrightarrow[N \to \infty]{P} \sum_{ \left( A_1, \dotsb, A_W \right) \in \mathcal{A} } P \left( T^0_1 \in A_1, \dotsb, T^0_W \in A_W \right) \tag{\text{from Remark} \ref{CMT_results}}\\
&= P \left( \widehat{\theta}^0 = r \right)
\end{align*}
\end{proof}
\end{lemma}

<!-------------- Theorem 1 --------------->
\bigskip \begin{theorem} 
\label{theorem1}
The empirical distribution of the number of rejections in the bootstrap samples is consistent in $N$ and $B$ for the true distribution of the number of rejections under the strong null:
$$\widehat{F}_{\widehat{\theta}^{(j)}} \xrightarrow[B, N \to \infty]{P} F_{\widehat{\theta}^0} $$
\begin{proof}
Fix an arbitrarily small $\epsilon > 0$ and $\epsilon' > 0$. Remark \ref{glivenko} implies consistency, so there exists a $B^{*}$ such that, for all $B > B^{*}$:
\begin{align*}
P \left( \vert \widehat{F}_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^{(j)} } \vert > \epsilon / 2 \right) &< \epsilon'/2
\end{align*}
Similarly, from Lemma \ref{lemma3}, there exists an $N^{*}$ such that, for all $N > N^{*}$:
\begin{align*}
P \left( \vert F_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^0 } \vert > \epsilon / 2 \right) &< \epsilon' / 2
\end{align*}
Thus, for all $B$ and $N$ such that $B > B^{*}$ and $N > N^{*}$:
\begin{align*}
P \left( \vert \widehat{F}_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^0 } \vert > \epsilon \right) &= P \left( \vert  \widehat{F}_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^{(j)} } + F_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^0 } \vert > \epsilon \right) \\
&\le P \left( \vert \widehat{F}_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^{(j)} } \vert + \vert F_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^0 } \vert > \epsilon \right)  \tag{triangle inequality}\\
&\le P \left( \vert \widehat{F}_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^{(j)} } \vert > \epsilon/2 \; \text{ or } \; \vert F_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^0 } \vert > \epsilon/2 \right) \\
&\le P \left( \vert \widehat{F}_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^{(j)} } \vert > \epsilon/2 \right) + P \left( \vert F_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^0 } \vert > \epsilon/2 \right) \\
&< \epsilon' 
\end{align*}
\end{proof}
\end{theorem}

<!-------------------------- --------------------------------->
<!-------------- SECTION: SIMULATION STUDY --------------->
<!-------------------------- --------------------------------->

\section{Simulation study}

See simulations from 2017-8-1 (written up in "multiple_outcomes (4).pdf"). Maybe use these these scenarios again? Generate at least some data under null to show that power of joint test is <5\%, as intended.

\subsection{Width of 95\% CI generated under null}

\begin{itemize}
\item Figure 1: Mean number of rejections and average of CI limits to show width
\item Figure 2: Average distance of observed rejections from upper CI limit
\end{itemize}

\subsection{Power of joint test}

\begin{itemize}
\item Figure 3 with panels: Prob of rejecting joint null in samples generated under null (should be <5\% for all) and under various alternative scenarios
\end{itemize}

<!-------------------------- --------------------------------->
<!-------------- SECTION: APPLIED EXAMPLE --------------->
<!-------------------------- --------------------------------->

\section{Applied example}

<!--
\newpage
\section*{References}
-->

### 2018-03-01 Project log/multiple_outcomes_project_log.Rmd
Try to fix git syncing

@@
 author: ''
+csl: aje_style.csl
+output:
+  pdf_document: default
+  html_document:
+    df_print: paged
 header-includes:
@@
 - \usepackage{float}
-output: pdf_document
 bibliography: refs_fake.bib
-csl: aje_style.csl
 ---
@@
 
-2018-1-15
-Mysteries: 
-* Understand why FCR resampling doesn't work (these resamples reject more often than the original data).
-I found that prob of rejection even when there is only 1 covariate is too high.
-In my simple script, try resampling residuals and re-attaching them to fitted values.
-Does that work?
-* Is it theoretically valid to use the weird CI for excess hits using the observed hits and the CI under the null?
+\subsection{FWER control methods we should compare for power of joint test}
 
-Since we are bootstrapping and then fitting parametric model (OLS), I think we should use Fox's or Westfall's residual resampling approach (and indeed, this seemed to work in the last simulation).
-Seems like the basic approach automatically violates OLS assumptions, meaning our inference will be wrong and the probability of rejecting the test could be non-nominal.
-Our application of bootstrapping is unusual in that we need to resample in a way that is compatible with parametric assumptions of the test we're going to apply for each outcome.
+\begin{enumerate}
+\item Westfall's single-step minP and maxT methods (pg 47): Easy to code manually
+\item Westfall's step-down free method (pg 66): Moderate to code manually
+\item Romano: Implemented in StepwiseTest package, but needs debugging
+\item Holm (step-down): Easy to code manually; UMP than Bonferroni
+\item Bonferroni (single-step): Easy to code manually
+\end{enumerate}
+
+\subsection{Should we compare our joint test to joint tests based on non-FWER procedures?}
+
+If we reject joint null based on FWER control method, then we can conclude with 95\% confidence that this rejection is real, which implies that the joint null is false.
+
+With $k$-FWER control, we can conclude with 95\% confidence that there are fewer than $k$ false positives, so $R-k+1$ of the observed rejections are real (don't know which ones), where $R$ is the number of adjusted rejections.
+So, as long as $R>k-1$, we can reject the joint null with 95\% confidence.
+PROBLEM: WOULD HAVE TO TRY EVERY $k$??
+
+With FDP control, we can reject the joint null if the proportion of adjusted rejections exceeds the allowed amount ($\gamma$). Again, problem because we'd have to try every $\gamma$?
+
+FDR does not work for a joint test because it uses expected proportion of rejections.
+
+The problem is that $k$-FWER and FDP would probably lead to more powerful joint test than the FWER control method, but not clear which parameters to choose. So maybe we should not worry about them?
+
+\subsection{What we can learn from different procedures}
+
+Say we observe 18 rejections at the 0.05 level. Then, with increasing stringency, we can say:
+\begin{itemize}
+\item Our null CI is $[0, 10]$. "We observed $18-10 = 8$ more hits than we would expect in 95\% of samples taken under the strong null (every null holds)."
+\item $8-$FWER controlling procedure (Romano) rejects 11.
+"There are fewer than 8 false positives, so of these 14 rejections, at least $14-7 = 7$ are real effects, but we don't know which ones." (How should we decide which $k$ to use here?)
+\item FWER-controlling procedure (Bonferroni or preferably Romano/Westfall/another resampling method) rejects 3.
+"All 3 of these rejections are real effects with 95\% confidence", or alternatively: "We reject the strong null with >95\% confidence."
+\end{itemize}
+
+In practice, you could get the first two from the same bootstrap resamples (taken parametrically under the strong null).
+Then you'd have to resample under the original distribution to get the third one.
+
+\subsection{Our two metrics}
 
@@
 
-Also, there might be models for which you can't resample as planned. For example, with multiple regression, we
-
 <!-- NEW SECTION -->
 \section{Possible resampling algorithms}
+
+\textbf{Conclusion:} Since we are bootstrapping and then fitting parametric model (OLS), I think we should use Fox's or Westfall's residual resampling approach (and indeed, this seemed to work in the last simulation).
+Full-case resampling violates the iid error assumption in OLS regression, meaning the probability of rejecting the null within each resample is not $\alpha$.
+Our application of bootstrapping is unusual in that we need to resample in a way that is compatible with parametric assumptions of the test we're going to apply for each outcome.
 
@@
 
-\subsection{CV: Cases when naive bootstrap fails}
-\url{https://stats.stackexchange.com/questions/11210/assumptions-regarding-bootstrap-estimates-of-uncertainty?noredirect=1&lq=1}
-
 \subsection{CV: Applying bootstrap to arbitrary smooth function of the data}
@@
 \section{Notes on past literature}
+
+\subsection{Other "global tests" (i.e., of joint null)}
+
+\begin{itemize}
+\item Fisher's Combination Test aggregates all the p-values instead of just using the smallest one; better than Bonferroni joint test if many small effects, but worse if there are a few large effects.
+Assumes independent tests.
+\end{itemize}
+
+\subsection{Frane (per-family error rates)}
+
+\begin{itemize}
+\item Per-family error rate (PFER): Number of expected false positives 
+\item Controlling PFER implies controlling FWER (so PFER is more stringent)
+\item Bonferroni also controls PFER to $\alpha$, which is partly why it's more "conservative"
+\end{itemize}
+
+Talks about *another interesting interpretation of Bonferroni*, which is that (unlike other FWER-control procedures), it also controls the more stringent "per-family error rate" (PFER), which is the expected number of false positives.
+
+With other FWER procedures that don't control PFER, which is most of them, there is a <5\% probability of at least one false positive, but that could be 1 false positive or it could be 100.
+With Bonferroni, if there are false positives (<5\% chance), it's also the case that the expected number of them is $< 0.05 \times$ (# of tests) with <5\% probability.
+
+So to build upon Tyler's earlier statement, I believe Bonferroni lets us say:
+
+*"We did 100 tests and rejected 20.
+With >95\% confidence, all 20 are true effects (FWER).
+Even if we are in the 5% of instances where we actually do have false positives, there aren't very many of them: only 5 or fewer with 95\% confidence (PFER)."*
+
+\subsection{Yekutieli \& Benjamini}
+
+A parametric bootstrap correction for LMM.
+
+\subsection{Yekutieli \& Benjamini}
+
+A resampling-based procedure for FDR control. Very readable!
+
+\begin{itemize}
+\item Pg. 7: Outlines residual resampling under $H_0$, exactly as we will be doing
+\item Talks a lot about Westfall and also uses subset pivotality
+\item Pg 7: "Subset pivotality is achieved if the distribution of [the p-values of the true null hypotheses] is unaffected by the truth or falsehood of the remaining hypotheses"
+\item Seems to be very closely related to Westfall, but cares about FDR instead of FWER. 
+\end{itemize}
+
+\subsection{van der Laan, Dudoit, Pollard (2004) - not saved}
+
+A resampling-based multiple testing control procedure.
+
+\begin{itemize}
+\item $k-$FWER control procedure where you first reject hypotheses based on a procedure that controls regular FWER, but then also "augment" by rejecting an additional $k=1$ hypotheses (summarized by Romano \& Wolf, pg 1396)
+\end{itemize}
+
+\subsection{Romano \& Wolf}
+
+A resampling-based multiple testing control procedure for FWER, k-FWER, or FDP.
+
+\begin{itemize}
+\item $k-$FWER control: With 95\% confidence, the number of false positives is $< k$
+\item FDP control: With 95\% confidence, the proportion of rejections that are false positives is $< \gamma$
+\item FDR control: with 95\% confidence, the EXPECTED proportion of rejections that are false positives is $< \gamma$ 
+\item They bootstrap under the original data distribution and use this to get quantiles of the distribution of the $k^{th}$-largest test statistic for various subsets of the test statistics, allowing control of the $k-$FWER through a step-down procedure.
+\item Why resample not under the null?
+Since we don't know which hypotheses are true, assume the data really were generated under the null because this is conservative.
+If the data aren't generated under the null, then the critical values will just be more stringent because the kth-largest test stat will be larger (page 1383).
+Basically this "exploits the duality between CIs and hypothesis tests" (pg 1388).
+Apparently resampling not under the null avoids subset pivotality (pg 1388).
+\item Page 1396: Talks about other resampling methods in some detail, including Westfall
+\item Does not use subset pivotality, unlike Westfall
+\end{itemize}
 
@@
 
+Resampling-based procedures for FWER control.
+
 \begin{itemize}
-
-\item I think their algorithm is closest to the one Tyler initially suggested (fix $X$ while resampling $Y$), except that they first center the outcomes (i.e., resample the residuals).
 
@@
 
+\item \hl{Single-step adjusted p-value, aka "minP" method (page 46, 51): For each test $j$, set its adjusted p-value equal to the probability in the resamples (generated under the strong null) that the minimum p-value is less than the observed p-value for test $j$.
+So, it's like a p-value of the p-value (probability of observing a p-value at this small under the strong null)}
+
+\item \hl{Free step-down resampling method (pg 66): Not very intuitive...}
+
+\item Strong FWER control
+
 \item 106-109: More on OLS
@@
 If the actual data are pooled and resampled, then the underlying residual distribution will be estimated using a multi-modal empirical distribution function when the means differ.
-Such resampling violates the first principle of Hall \& Wilson (1991): '...even if the data might be drawn in a way that fails to satisfy $H_0$, resampling should be done in a way that reflects $H_0$'." I DON'T GET WHY UNCENTERED APPROACH VIOLATES H0!
+Such resampling violates the first principle of Hall \& Wilson (1991): '...even if the data might be drawn in a way that fails to satisfy $H_0$, resampling should be done in a way that reflects $H_0$'."
 
@@
 \begin{itemize}
+
+\item \hl{Subset pivotality assumption:} Take any subset of the hypotheses to be tested.
+The distribution of the max test statistic within this group, given that all the nulls in that subset hold, is the same as its distribution given the strong null (all the nulls for ALL the hypothesis tests hold).
+In other words, once you know that all nulls within the subset are true, knowing that all the other nulls are true doesn't change anything about distribution of max test statistic in that subset.
+
+\item Reason for using subset pivotality: To avoid having to resample under every intersection null (i.e., subsets of nulls being true) so that we can just resample under the joint null.
 
@@
 \subsection{Resolved questions}
+Q: Why do datasets created through full-case resampling reject the null too often?
+A: See my Cross Validated question (\url{https://stats.stackexchange.com/questions/323455/why-do-hypothesis-tests-on-resampled-datasets-reject-the-null-too-often}).
+Basically, when we resample full cases, we violate the OLS assumption of iid errors because every repeated pair $(X,Y)$ has exactly the same residual, so the residuals are highly correlated with $X$.
+If we fix $X$ and resample residuals, then even though there are repeated observations, the errors of the model newly fit to the resampled data are iid because we are breaking any association between the resampled residuals and $X$.
+Why do Davison \& Hinkley recommend this approach if it's wrong?
+Because it's only wrong if you need to meet parametric assumptions within each resample.
+Our application of bootstrapping is unusual in that we need to resample in a way that is compatible with parametric assumptions of the test we're going to apply for each outcome.
+
 Q: Why does Blakesley/Westfall’s minP bootstrapping work since it is an extreme order statistic?
@@
 \section{Simulation notes}
+
+Next up: Try regenerating residuals on Sherlock with multiple outcomes and huge sample size. Trying n = 100,000 and B = 2,000 and 5-hr time limit?
+
+Seems like when rejection prob is low, bootstraps do worse? E.g., rho.XY = 0.03 worked perfectly with $n=10,000$, but rho.XY = 0.01 did not.
+
+\subsection{2018-1-23 unsaved local sims}
+
+Revisiting residual resampling for CI: Fix design matrix, then re-generate residuals using $\widehat{\beta}$ and $\widehat{\sigma^2}$.
+(Maybe resampling the residuals would also work, but should we be worried about non-independence?)
+
+\begin{itemize}
+\item Tried a single outcome and different correlation strengths
+\item Compared average rejection probability in bootstraps to average in original data
+\item Need large $n$ for this! Using $n=1000$ didn't work because regression estimates aren't precise enough. Using $n=10,000$ seemed okay.
+\item Not yet assessing CIs for number of rejections because would need multiple outcomes for that
+\end{itemize}
+
+Went back to debug fcr script and found that reattaching residuals works under null.
+
+Under alternative, though, got 25% rejections across all bootstrap resamples vs. 15% across 1000 sim reps in original datasets.
+
+Still got 25% even when generating new residuals and adding them to the original fitted values!!!!!! Whyyyyyyyyyy???????????
+
+\subsection{2018-1-22}
+
+\begin{itemize}
+\item Reduced number of outcomes to 30 to reduce computation time
+\item Now also computing CIs based on raw percentiles (below) scaled by the estimated rate
+\item And am trying with data generated under alternative
+\item Still using uncorrelated outcomes
+\item Rates are again almost perfectly estimated
+\item \hl{We can't use subsampling when data are generated under $H_A$ because the number of rejections will be lower due to lower power.
+So the sampling distribution isn't just wider in the smaller samples; it doesn't even have the same expected value.}
+\end{itemize}
+
+\subsection{2018-1-20 round 2 subsampling}
+
+\begin{itemize}
+\item Running same as below, but with the real code instead of raw quantiles. "covers.correct" uses exactly rate = 0 as below; "covers" uses the rate estimate. 
+\item Rates basically always estimated as 0 now because of larger $B$ 
+\item Coverage now 80% instead of 70%
+\item $\ast$Conclusion: $B=20,000$ instead of $B=2,000$ results in exactly correct rate estimation, at least for data generated under $H_0$. Can probably find middle ground of simulation reps.
+\item Based on local simulations (debug\_subsampling), seems like a CI based just on percentiles of the subsamples is better than the one in the lab handout, where it's centered on the point estimate.
+\end{itemize}
+
+One-sided CI idea to improve coverage a little
+
+\subsection{2018-1-20 subsampling}
+
+\begin{itemize}
+\item Based on the exact variance I derived, when the null is true, the SE of $\widehat{\theta}$ doesn't depend on $n$ because tests maintain $\alpha$ probability of rejecting regardless of $n$.
+So in this case, the rate is exactly 0 (as correctly estimated by the last round of simulations).
+\item So these simulations simplify things by not estimating the rate, but just taking the quantiles of the iterates for the third sample size ($n=900$). Seems like this should work fine. 
+\item $\ast$Giving it the exact rate yields 100% coverage, up from 70% in previous simulation
+\item Sample sizes same as below
+\item Rep time 90th percentile: 1700 seconds with $B=2000$
+\end{itemize}
+
+\subsection{2018-1-19 - results not saved}
+
+\begin{itemize}
+\item Same as 2018-1-18, but increased all sample sizes by order of magnitude
+\item Mean rate exactly 0 with min/max $pm 0.10$ and 25th, 75th percentiles $\pm 0.015$
+\item Still 70% coverage
+\item Used $n=100,000$ and subsample sizes $n_s = (400, 600, 900, 1350)$ to match lab handout, but I have $B = 2,000$ instead of $B = 20,000$
+\end{itemize}
+
+\subsection{2018-1-18 - results not saved}
+
+\begin{itemize}
+\item Trying subsampling with rate estimation in a null scenario and no correlation between outcomes
+\item Only getting 70% coverage
+\item Used $n=10,000$ and subsample sizes $n_s = (40, 60, 90, 135)$ to match lab handout, but I have $B = 2,000$ instead of $B = 20,000$
+\item Each sbatch took ~3.5 hours to run (upper limit)
+\end{itemize}
+
+\subsection{2018-1-16 "under $H_0$ and FCR"}
+
+\begin{itemize}
+\item Joint test plot shows power of our joint test (i.e., resample residuals under $H_0$ and see if observed hits are extreme in this distribution) and compares to power of a naïve Bonferroni joint test (i.e., reject the joint null if any of the 100 Bonferroni-adjusted tests rejects).
+\item Above results make sense: e.g., we are near the nominal joint alpha level (0.05 regardless of alpha level for individual tests) when data were generated under joint null. 
+\item Excess hits plot shows average excess hits (hits above expectation) and CIs averaged across bootstraps.
+The CIs are sometimes completely above the observed excess hits.
+\textbf{This is because the FCR resampling approach is wrong.} See the resolved questions.
+\end{itemize}
 
@@
 
-\item{ \textbf{The apparently too-conservative behavior for the uncorrelated case is a benign artifact} of $\widehat{\theta}$'s discreteness.
+\item {\textbf{The apparently too-conservative behavior for the uncorrelated case is a benign artifact} of $\widehat{\theta}$'s discreteness.
   \begin{itemize}

### 2018-03-01 R code/For Sherlock/analysis.R

@@ -19,13 +19,16 @@ table( s$scen )
 kill = c( names(s)[ grepl( "n.rej.bt", names(s) ) ], "n.rej.5e.04" )
 s = s[ , !names(s) %in% kill ]
 
-# did naive Bonferroni reject?
-s$naive_bonf_rej = s$n.rej.5e.04.1 > 0
+# is coverage different from 95%?
+mean(s$covers)
+mean(s$covers.correct)
+prop.test( sum(s$covers), length(s$covers), p=0.95 )
 
-
-# TEMP: see if reshape problem fixed by random noise
-#s[,11:30] = s[,11:30] + rnorm(n=100, mean=0, sd=0.01)
-#temp = s[,c(10,14,15,16)]
+# what happens we we only 
+s$sim.rep = rep( 1:(nrow(s)/2000), each = 2000 )
+first = s[ !duplicated( s$sim.rep ), ]
+mean(s$n.rej.0.05); mean(s$n.rej.bt.0.05)
+s$covers = 
 
 
 ########################### MAKE DATASETS FOR FIRST PLOT (RESIDUAL RESAMPLING) ###########################
@@ -69,7 +72,7 @@ power$group = paste( "X-Y correlation: ", power$rho.XY )  # for plotting joy
 crit = tidy %>% group_by( scen, alpha, rho.XY, rho.YY ) %>% filter( stat == "crit" & bt.type == "resid" ) %>% summarise( crit = mean(value) )
 crit$group = paste( "X-Y correlation: ", power$rho.XY )  # for plotting joy
 crit$expect = as.numeric( crit$alpha ) * 100
-crit$crit.excess = crit$crit - 
+#crit$crit.excess = crit$crit - 
 
 
 ######## Aggregated Dataset 2: Naive Bonferroni Power ######## 

### 2018-03-01 R code/For Sherlock/doParallel.R

@@ -1,24 +1,31 @@
 
 ########################### LOAD COMMAND-LINE ARGUMENTS  ###########################
 
-# FOR CLUSTER USE
-# load command line arguments
-args = commandArgs(trailingOnly = TRUE)
-jobname = args[1]
-scen = args[2]  # this will be a letter
-boot.reps = as.numeric( args[3] )
+# ~~~ CHECK BEHAVIOR WITH ALPHA VECTOR LONGER THAN 1
 
-# get scen parameters
-setwd("/share/PI/manishad/multTest")
-scen.params = read.csv( "scen_params.csv" )
-p = scen.params[ scen.params$scen.name == scen, ]
 
-# no longer included in parameters because it's a vector
-crit.bonf = 0.05 / p$nY
-alpha = c( crit.bonf, 0.01, 0.05 )
 
-# simulation reps to run within this job
-sim.reps = 10
+# ######### FOR CLUSTER USE #########
+# # load command line arguments
+# args = commandArgs(trailingOnly = TRUE)
+# jobname = args[1]
+# scen = args[2]  # this will be a letter
+# boot.reps = as.numeric( args[3] )
+# 
+# # get scen parameters
+# setwd("/share/PI/manishad/multTest")
+# scen.params = read.csv( "scen_params.csv" )
+# p = scen.params[ scen.params$scen.name == scen, ]
+# 
+# # no longer included in parameters because it's a vector
+# crit.bonf = 0.05 / p$nY
+# alpha = c( crit.bonf, 0.01, 0.05 )
+# 
+# # simulation reps to run within this job
+# # this need to match n.reps.in.doParallel in the genSbatch script
+# sim.reps = 1
+# ######### END OF CLUSTER PART #########
+
 
 ######### FOR LOCAL USE #########
 # setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/Sandbox/2018-1-13")
@@ -31,8 +38,10 @@ rho.YY = c(0)
 rho.XY = c(0)  # null hypothesis: 0
 
 # bootstrap iterates and type
-boot.reps = 200
-bt.type = c( "fcr" )  # fcr: resample under HA; resid: resample under H0
+boot.reps = 500
+sim.reps = 1
+scen = "a"
+bt.type = c( "ha.resid.2" )  # fcr: resample under HA; resid: resample under H0
 
 
 # matrix of scenario parameters
[…]

### 2018-03-01 R code/For Sherlock/functions.R

@@ -115,7 +115,7 @@ sim_data = function( .n, .cor ) {
 # fit regression model for a single outcome Y regressed on all Xs
 # return p-value for the exposure of interest (X1)
 
-fit_model = function( Y.name, .dat, .resid ) {
+fit_model = function( Y.name, .dat, .resid, .sigma, .tval, .intercept ) {
 
   # extract names
   X.names = names( .dat )[ grep( "X", names( .dat ) ) ]
@@ -131,10 +131,27 @@ fit_model = function( Y.name, .dat, .resid ) {
   # extract p-value for exposure of interest (X1)
   m = lm( eval( parse(text=formula) ), data=.dat )
   pval = summary(m)$coefficients[ "X1", 4 ]
-
-  # get residuals if needed for bootstrapping
-  if ( ! .resid ) return( list( pval = pval ) )
-  else if ( .resid ) return( list( pval = pval, resid = residuals(m) ) )
+  
+  if ( ! .resid ) resid.return = NA
+  else resid.return = residuals(m)
+    
+  if ( ! .sigma ) sigma.return = NA
+  else sigma.return = summary(m)$sigma
+  
+  if ( ! .intercept ) intercept.return = NA
+  else intercept.return = coef(m)[["(Intercept)"]]
+  
+  # extract test stat for this test
+  # for use with Romano's code
+  if ( ! .tval ) tval.return = NA
+  else tval.return = summary(m)$coefficients[ "X1", 3 ]
+  
+  return( list( pval = pval,
+                resid = resid.return,
+                sigma = sigma.return, 
+                intercept = intercept.return,
+                tval = tval.return ) )
+  
 }
 
 # test drive
@@ -144,10 +161,12 @@ fit_model = function( Y.name, .dat, .resid ) {
 
 
 ########################### FN: GIVEN DATASET, RETURN STATS ###########################
-
-# .resid: should it return dataset with residuals (for resampling under null)?
-
-dataset_result = function( .dat, .alpha, .resid = FALSE ) {
+# .resid: should it return dataset with residuals (for regenerating residuals)?
+# . sigma: should it return fitted error SD (for regenerating residuals)?
+# .intercept: should it return fitted intercept (for regenerating residuals)?
+dataset_result = function( .dat, .alpha,
+                           .resid = FALSE, .sigma = FALSE, .intercept = FALSE,
+                           .tval = FALSE ) {
 
   # extract names of outcome variables
[…]

### 2018-03-01 R code/For Sherlock/gen_sbatch.R

@@ -10,24 +10,24 @@ setwd(path)
 # path = "~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/Sandbox/2018-1-13"
 # setwd(path)
 
-# from previous simulation
+n = 5e4
+nX = 1
+nY = 100
+rho.XX = 0
+rho.YY = c(0.3)
+rho.XY = 0.008  # null hypothesis: 0
+
 # n = 1000
 # nX = 1
 # nY = 100
 # rho.XX = 0
 # rho.YY = c(0, 0.25, 0.5, 0.75)
-# rho.XY = 0.03  # null hypothesis: 0
-
-n = 1000
-nX = 1
-nY = 100
-rho.XX = 0
-rho.YY = c(0, 0.25, 0.5, 0.75)
-rho.XY = c(0.08, 0.03, 0)  # null hypothesis: 0
+# rho.XY = c(0, 0.03, 0.08)  # null hypothesis: 0
 
 # bootstrap iterates and type
 boot.reps = 2000
-bt.type = c( "resid", "fcr" )  # fcr: resample under HA; resid: resample under H0
+# boot.reps = 2000
+bt.type = c( "ha.resid.2" )  # fcr: resample under HA; resid: resample under H0
 
 
 # matrix of scenario parameters
@@ -52,7 +52,7 @@ source("functions.R")
 
 # number of sbatches to generate (i.e., iterations within each scenario)
 n.reps.per.scen = 1000
-n.reps.in.doParallel = 10
+n.reps.in.doParallel = 1
 n.files = ( n.reps.per.scen / n.reps.in.doParallel ) * n.scen
 
 
@@ -64,7 +64,6 @@ outfile = paste("rm_", 1:n.files, ".out", sep="")
 errorfile = paste("rm_", 1:n.files, ".err", sep="")
 write_path = paste(path, "/sbatch_files/", 1:n.files, ".sbatch", sep="")
 runfile_path = paste(path, "/testRunFile.R", sep="")
-
 
 sbatch_params <- data.frame(jobname,
                             outfile,
@@ -83,13 +82,13 @@ sbatch_params <- data.frame(jobname,
                             stringsAsFactors = F,
                             server_sbatch_path = NA)
 
-# generateSbatch(sbatch_params, runfile_path)
+generateSbatch(sbatch_params, runfile_path)
 
 
 # run them all
 # works
[…]

### 2018-03-01 R code/For Sherlock/stitch_on_sherlock.R

@@ -3,19 +3,51 @@
 
 path = "/share/PI/manishad/multTest"
 setwd(path)
+source("functions.R")
 
+######## FOR STITCHING LONG FILES ########
+
+s = stitch_files(.results.singles.path = "/share/PI/manishad/multTest/sim_results/long",
+                 .results.stitched.write.path = "/share/PI/manishad/multTest/sim_results",
+                 .name.prefix = "long",
+                 .stitch.file.name="stitched.csv")
+
+# write results
+write.csv( s, "stitched.csv" )
+
+
+
+#### Sneak-Peek Analyes
+mean(s$covers.raw)
+mean(s$covers)
+mean(s$covers.correct)
+
+summary(s$rate)
+
+# symmetry of CIs
+mean( abs( s$ci.hi.raw - s$excess ) - abs( s$ci.lo.raw - s$excess ) )
+mean( abs( s$ci.hi.correct - s$excess ) - abs( s$ci.lo.correct - s$excess ) )
+
+
+
+
+
+
+# zip the entire results folder
+setwd("/share/PI/manishad/multTest")
+system( paste("zip -r sim_results.zip sim_results") )
+
+
+######## FOR STITCHING SHORT FILES ########
 source("functions.R")
 s = stitch_files(.results.singles.path = "/share/PI/manishad/multTest/sim_results/short",
                  .results.stitched.write.path = "/share/PI/manishad/multTest/sim_results",
              .name.prefix = "short",
              .stitch.file.name="stitched.csv")
 
-# # write results
-# write.csv( s, "stitched_prepped.csv" )
-# 
-# # zip the entire results folder
-# setwd("/share/PI/manishad/multTest")
-# system( paste("zip -r sim_results.zip sim_results") )
-# 
+# zip the entire results folder
+setwd("/share/PI/manishad/multTest")
+system( paste("zip -r sim_results.zip sim_results") )
+
 
 

### 2018-03-01 R code/For Sherlock/doParallel.R
Test commit: updating doParallel for comparative sim study of joint test

@@ -3,7 +3,7 @@
 
 # ~~~ CHECK BEHAVIOR WITH ALPHA VECTOR LONGER THAN 1
 
-
+ 
 
 # ######### FOR CLUSTER USE #########
 # # load command line arguments

### 2018-03-01 R code/For Sherlock/doParallel.R
h0.parametric method added

@@ -41,7 +41,7 @@ rho.XY = c(0)  # null hypothesis: 0
 boot.reps = 500
 sim.reps = 1
 scen = "a"
-bt.type = c( "ha.resid.2" )  # fcr: resample under HA; resid: resample under H0
+bt.type = c( "h0.parametric" )
 
 
 # matrix of scenario parameters
@@ -112,11 +112,13 @@ for ( j in 1:sim.reps ) {
   
   # get number rejected for observed dataset
   # vector with same length as .alpha
+  # only return residuals and/or model estimates if we need them later for resampling
   samp.res = dataset_result( .dat = d, .alpha = alpha,
                              .resid = ifelse( p$bt.type %in% c("resid", "ha.resid", "ha.resid.2"),
-                                              TRUE, FALSE), # don't return residuals unless we need them later
-                             .sigma = TRUE,
-                             .intercept = TRUE )
+                                              TRUE, FALSE), 
+                             .sigma = ifelse( p$bt.type %in% c( "h0.parametric", "ha.resid.2"),
+                                              TRUE, FALSE),
+                             .intercept = (p$bt.type == "h0.parametric") )
   n.rej = samp.res$rej  # first one is Bonferroni
   names(n.rej) = paste( "n.rej.", as.character(alpha), sep="" )
   
@@ -161,10 +163,8 @@ for ( j in 1:sim.reps ) {
         Xs = as.data.frame( d[ , 1 : length(X.names) ] )
         names(Xs) = X.names
         
-        # regenerate Ys using fitted values
-        new.Ys = matrix( NA, nrow = nrow(d), ncol = length(Y.names) )
-        
         # parametrically generate Ys, setting beta of interest to 0
+        new.Ys = matrix( NA, nrow = nrow(d), ncol = length(Y.names) )
         new.Ys = sapply( 1:ncol(resid),
                              function(x) rnorm( n=n.cells, mean = intercept[x], sd = sigma[x] ) )
         b = as.data.frame( cbind( Xs, new.Ys ) )

### 2018-03-01 R code/For Sherlock/functions.R

@@ -165,7 +165,9 @@ fit_model = function( Y.name, .dat, .resid, .sigma, .tval, .intercept ) {
 # . sigma: should it return fitted error SD (for regenerating residuals)?
 # .intercept: should it return fitted intercept (for regenerating residuals)?
 dataset_result = function( .dat, .alpha,
-                           .resid = FALSE, .sigma = FALSE, .intercept = FALSE,
+                           .resid = FALSE,
+                           .sigma = FALSE,
+                           .intercept = FALSE,
                            .tval = FALSE ) {
 
   # extract names of outcome variables
@@ -205,8 +207,6 @@ dataset_result = function( .dat, .alpha,
   # if we are returning the estimated intercept
   if ( .intercept ) intercepts = as.vector( u[ names(u) == "intercept" ] )
   
-  # ~~~ BOOKMARK: I WAS STEPPING THROUGH THIS FN TO SEE IF INTERCEPTS GET RETURNED PROPERLY
-  
   # if we are returning the test stats for use with Romano
   if ( .tval ) tvals = as.vector( u[ names(u) == "tval" ] )
   
@@ -234,12 +234,16 @@ dataset_result = function( .dat, .alpha,
   if ( ! .sigma ) sigmas.return = NA
   else sigmas.return = sigmas
   
+  if ( ! .intercept ) intercept.return = NA
+  else intercept.return = intercepts
+  
   if ( ! .tval ) tvals.return = NA
   else tvals.return = tvals
   
   return( list( rej = rej,
                 resid = resid.return,
                 sigmas = sigmas.return,
+                intercepts = intercept.return,
                 tvals = tvals.return,
                 pvals = pvals ) )
 }

### 2018-03-01 R code/For Sherlock/doParallel.R
Cleaned up doParallel to accommodate Westfall comparison

@@ -30,15 +30,15 @@
 ######### FOR LOCAL USE #########
 # setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/Sandbox/2018-1-13")
 # p = read.csv("scen_params.csv")  # should be a single row, I think
-n = 1000
+n = 250
 nX = 1
-nY = 100
+nY = 10
 rho.XX = 0
 rho.YY = c(0)
-rho.XY = c(0)  # null hypothesis: 0
+rho.XY = c(0.05)  # null hypothesis: 0
 
 # bootstrap iterates and type
-boot.reps = 500
+boot.reps = 5
 sim.reps = 1
 scen = "a"
 bt.type = c( "h0.parametric" )
@@ -127,7 +127,7 @@ for ( j in 1:sim.reps ) {
   jt.rej.bonf.naive = n.rej[1] > 0
   
   # run all bootstrap iterates
-    r = foreach( icount( boot.reps ), .combine=rbind ) %dopar% {
+    r = foreach( i = 1:boot.reps, .combine=rbind ) %dopar% {
       # draw bootstrap sample 
       ids = sample( 1:nrow(d), replace=TRUE )
       
@@ -165,8 +165,9 @@ for ( j in 1:sim.reps ) {
         
         # parametrically generate Ys, setting beta of interest to 0
         new.Ys = matrix( NA, nrow = nrow(d), ncol = length(Y.names) )
-        new.Ys = sapply( 1:ncol(resid),
-                             function(x) rnorm( n=n.cells, mean = intercept[x], sd = sigma[x] ) )
+        n.cells = length(new.Ys)
+        new.Ys = sapply( 1:ncol(new.Ys),
+                             function(x) rnorm( n = n.cells, mean = intercept[x], sd = sigma[x] ) )
         b = as.data.frame( cbind( Xs, new.Ys ) )
       }
       
@@ -217,73 +218,96 @@ for ( j in 1:sim.reps ) {
         # generate columns of new residuals using the appropriate sigma
         new.resids = sapply( 1:ncol(resid), function(x) rnorm( n=n.cells, sd = sigma[x] ) )
         
-        # # regenerate residuals using fitted values
-        # n.cells = dim(resid)[1] * dim(resid)[2]
-        # # TEST ONLY - LATER WILL NEED TO SAVE VECTOR OF SIGMA-HATS FOR EACH REGRESSION
-        # new.resids = matrix( rnorm( n = n.cells, mean = 0, sd = 1 ),
[…]

### 2018-03-01 R code/For Sherlock/functions.R
Add fns for Westfall

@@ -1,5 +1,121 @@
 
 
+
+######################## FNS FOR WESTFALL's SINGLE-STEP ########################
+
+# returns minP-adjusted p-values (single-step)
+
+# p = original p-values
+# p.bt = bootstrapped p-values (vector) - an m X B matrix
+adjust_minP = function( p, p.bt ) {
+  
+  n.boot = ncol(p.bt)
+  
+  # keep only minimum p-value in each resample
+  minP.bt = apply( p.bt, MARGIN = 2, FUN = min )
+  
+  # for each element of p, get the proportion of resamples
+  #  whose minP was less than the present p-value
+  p.adj = unlist( lapply( p, FUN = function(x) sum( minP.bt <= x ) / n.boot ) )
+  
+  return(p.adj)
+}
+
+# test
+# B = 200
+# n.tests = 10
+# p.bt = matrix( runif(B*n.tests, 0, 1), nrow = n.tests)
+# p = runif( n.tests, 0, .1)
+# 
+# p.adj = adjust_minP( p, p.bt )
+# plot(p,p.adj)
+
+
+######################## FNS FOR WESTFALL's STEP-DOWN ########################
+
+# Westfall textbook, pages 66-67
+adj_Wstep = function( p, p.bt ) {
+  
+  # TEST ONLY
+  # p = structure(c(0.636112475277457, 0.116408752390887, 0.260603159221029, 
+  #                 0.575086408582202, 0.625789849782125), .Dim = c(5L, 1L))
+  # p.bt = structure(c(0.728140113074014, 0.828557135138776, 0.728464927503583, 
+  #                    0.376518353153738, 0.241363541131996, 0.415330227654616, 0.75686321503498, 
+  #                    0.865494617532842, 0.839819273240726, 0.540707408457488, 0.149482739547785, 
+  #                    0.506274953008223, 0.122605643925883, 0.747018941160177, 0.168076477020114, 
+  #                    0.0582566955778811, 0.284459435324328, 0.561050485854955, 0.13839444320949, 
+  #                    0.641335055023296, 0.195313599009888, 0.174692123827015, 0.982035588216878, 
+  #                    0.713525352129964, 0.548770486028363), .Dim = c(5L, 5L), .Dimnames = list(
+  #                      NULL, c("result.1", "result.2", "result.3", "result.4", "result.5"
+  #                      )))
+  
[…]

### 2018-03-02 R code/For Sherlock/doParallel.R
Test

@@ -1,67 +1,62 @@
 
-########################### LOAD COMMAND-LINE ARGUMENTS  ###########################
+######### FOR CLUSTER USE #########
+# load command line arguments
+args = commandArgs(trailingOnly = TRUE)
+jobname = args[1]
+scen = args[2]  # this will be a letter
+boot.reps = as.numeric( args[3] )
 
-# ~~~ CHECK BEHAVIOR WITH ALPHA VECTOR LONGER THAN 1
+# get scen parameters
+setwd("/share/PI/manishad/multTest")
+scen.params = read.csv( "scen_params.csv" )
+p = scen.params[ scen.params$scen.name == scen, ]
 
- 
-
-# ######### FOR CLUSTER USE #########
-# # load command line arguments
-# args = commandArgs(trailingOnly = TRUE)
-# jobname = args[1]
-# scen = args[2]  # this will be a letter
-# boot.reps = as.numeric( args[3] )
-# 
-# # get scen parameters
-# setwd("/share/PI/manishad/multTest")
-# scen.params = read.csv( "scen_params.csv" )
-# p = scen.params[ scen.params$scen.name == scen, ]
-# 
-# # no longer included in parameters because it's a vector
-# crit.bonf = 0.05 / p$nY
-# alpha = c( crit.bonf, 0.01, 0.05 )
-# 
-# # simulation reps to run within this job
-# # this need to match n.reps.in.doParallel in the genSbatch script
-# sim.reps = 1
-# ######### END OF CLUSTER PART #########
-
-
-######### FOR LOCAL USE #########
-# setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/Sandbox/2018-1-13")
-# p = read.csv("scen_params.csv")  # should be a single row, I think
-n = 250
-nX = 1
-nY = 10
-rho.XX = 0
-rho.YY = c(0)
-rho.XY = c(0.05)  # null hypothesis: 0
-
-# bootstrap iterates and type
-boot.reps = 5
-sim.reps = 1
-scen = "a"
-bt.type = c( "h0.parametric" )
-
-
-# matrix of scenario parameters
-scen.params = expand.grid( bt.type, n, nX, nY, rho.XX, rho.YY, rho.XY )
-names(scen.params) = c( "bt.type", "n", "nX", "nY", "rho.XX", "rho.YY", "rho.XY" )
-
-# name the scenarios
-# remove letters that are privileged variables in R
[…]

### 2018-03-02 R code/For Sherlock/functions.R

@@ -131,7 +131,7 @@ cell_corr = function( vname.1, vname.2, .rho.XX, .rho.YY, .rho.XY ) {
     vtype.1 = "outcome"
   }
   
-  if ( length( grep("X", vname.2 ) ) == 1 ) {
+  if ( length( grep("X", vname.2 ) ) == 1 ) { 
     vtype.2 = "covariate"
   } else if ( length( grep("Y", vname.2 ) ) == 1 ) {
     vtype.2 = "outcome"

### 2018-03-02 R code/For Sherlock/gen_sbatch.R

@@ -10,12 +10,12 @@ setwd(path)
 # path = "~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/Sandbox/2018-1-13"
 # setwd(path)
 
-n = 5e4
+n = 500
 nX = 1
 nY = 100
 rho.XX = 0
-rho.YY = c(0.3)
-rho.XY = 0.008  # null hypothesis: 0
+rho.YY = c(0.25)
+rho.XY = 0.03  # null hypothesis: 0
 
 # n = 1000
 # nX = 1
@@ -25,9 +25,8 @@ rho.XY = 0.008  # null hypothesis: 0
 # rho.XY = c(0, 0.03, 0.08)  # null hypothesis: 0
 
 # bootstrap iterates and type
-boot.reps = 2000
-# boot.reps = 2000
-bt.type = c( "ha.resid.2" )  # fcr: resample under HA; resid: resample under H0
+boot.reps = 500
+bt.type = c( "h0.parametric" )
 
 
 # matrix of scenario parameters
@@ -44,14 +43,15 @@ n.scen = length(scen.params[,1])
 write.csv( scen.params, "scen_params.csv" )
 
 
-########################### FN: SIMULATE 1 DATASET ###########################
+########################### GENERATE SBATCHES ###########################
 
 # load functions for generating sbatch files
 source("functions.R")
 
 
 # number of sbatches to generate (i.e., iterations within each scenario)
-n.reps.per.scen = 1000
+# n.reps.per.scen = 1000
+n.reps.per.scen = 1
 n.reps.in.doParallel = 1
 n.files = ( n.reps.per.scen / n.reps.in.doParallel ) * n.scen
 
@@ -65,10 +65,11 @@ errorfile = paste("rm_", 1:n.files, ".err", sep="")
 write_path = paste(path, "/sbatch_files/", 1:n.files, ".sbatch", sep="")
 runfile_path = paste(path, "/testRunFile.R", sep="")
 
+# was 5 hours with 2K bootstrap and n = 5e4
 sbatch_params <- data.frame(jobname,
                             outfile,
                             errorfile,
-                            jobtime = "5:00:00",
+                            jobtime = "1:00:00",
                             quality = "normal",
                             node_number = 1,
                             mem_per_node = 64000,

### 2018-03-02 R code/For Sherlock/functions.R
Allow generating data where only half of X-Y relationships are non-null

@@ -122,7 +122,12 @@ get_crit = function( p.dat, col.p ) {
 # assume that X1 is the covariate of interest
 # assume that none of the other covariates are associated with any outcomes
 
-cell_corr = function( vname.1, vname.2, .rho.XX, .rho.YY, .rho.XY ) {
+# .half: Should only the last half the outcomes have the specified correlation?
+#  (if FALSE, correlation matrix is exchangeable; otherwise first half are 0 and 
+#  second half are rho.XY)
+
+cell_corr = function( vname.1, vname.2, .rho.XX, .rho.YY, .rho.XY,
+                      .nY, .half = FALSE ) {
   
   # use grep to figure out if variables are covariates or outcomes
   if ( length( grep("X", vname.1 ) ) == 1 ) {
@@ -147,15 +152,34 @@ cell_corr = function( vname.1, vname.2, .rho.XX, .rho.YY, .rho.XY ) {
   # case 3: both are outcomes
   if ( vtype.1 == "outcome" & vtype.2 == "outcome" ) return( .rho.YY )
   
+  # browser()
+
   # case 4: one is a covariate and one is an outcome
   if ( vtype.1 != vtype.2 ) {
     # check if this is the covariate of interest
     # equal correlation between X1 and all outcomes
     # all other covariates have correlation 0 with outcome
-    if( "X1" %in% c(vname.1, vname.2) ) return( .rho.XY )  # use 0 for null hypothesis
+    if( "X1" %in% c( vname.1, vname.2 ) ) {
+      # exchangeable case
+      if ( !.half ) return( .rho.XY )
+      
+      # "half are correlated" case
+      if (.half) {
+        # find the one that is the outcome
+        outcome.name = ifelse( vtype.1 == "outcome", vname.1, vname.2 )
+        
+        # extract its number (4 for "Y4")
+        num = substring( outcome.name, first = 2 )
+        
+        # see if its number is greater than halfway or not
+        return( ifelse( num > .nY / 2, .rho.XY, 0 ) )
+        # bookmark
+      }
+    }
+    
+    # if we have multiple covariates and this isn't the one of interest
     else return(0)
   }
-  
 }
 
 # test drive
[…]

### 2018-03-08 R code/For Sherlock/functions.R
Fix qstar issue in Wstep function

@@ -21,7 +21,7 @@ adjust_minP = function( p, p.bt ) {
   return(p.adj)
 }
 
-# test
+# # test
 # B = 200
 # n.tests = 10
 # p.bt = matrix( runif(B*n.tests, 0, 1), nrow = n.tests)
@@ -36,15 +36,15 @@ adjust_minP = function( p, p.bt ) {
 # Westfall textbook, pages 66-67
 adj_Wstep = function( p, p.bt ) {
   
-  # TEST ONLY
-  # p = structure(c(0.636112475277457, 0.116408752390887, 0.260603159221029, 
+  # # TEST ONLY
+  # p = structure(c(0.636112475277457, 0.116408752390887, 0.260603159221029,
   #                 0.575086408582202, 0.625789849782125), .Dim = c(5L, 1L))
-  # p.bt = structure(c(0.728140113074014, 0.828557135138776, 0.728464927503583, 
-  #                    0.376518353153738, 0.241363541131996, 0.415330227654616, 0.75686321503498, 
-  #                    0.865494617532842, 0.839819273240726, 0.540707408457488, 0.149482739547785, 
-  #                    0.506274953008223, 0.122605643925883, 0.747018941160177, 0.168076477020114, 
-  #                    0.0582566955778811, 0.284459435324328, 0.561050485854955, 0.13839444320949, 
-  #                    0.641335055023296, 0.195313599009888, 0.174692123827015, 0.982035588216878, 
+  # p.bt = structure(c(0.728140113074014, 0.828557135138776, 0.728464927503583,
+  #                    0.376518353153738, 0.241363541131996, 0.415330227654616, 0.75686321503498,
+  #                    0.865494617532842, 0.839819273240726, 0.540707408457488, 0.149482739547785,
+  #                    0.506274953008223, 0.122605643925883, 0.747018941160177, 0.168076477020114,
+  #                    0.0582566955778811, 0.284459435324328, 0.561050485854955, 0.13839444320949,
+  #                    0.641335055023296, 0.195313599009888, 0.174692123827015, 0.982035588216878,
   #                    0.713525352129964, 0.548770486028363), .Dim = c(5L, 5L), .Dimnames = list(
   #                      NULL, c("result.1", "result.2", "result.3", "result.4", "result.5"
   #                      )))
[…]

### 2018-03-08 Project log/multiple_outcomes_project_log.Rmd
,

@@
 
+<!-- NEW SECTION -->
+\section{Notes on Ying's MIDUS paper}
+
+\subsection{Variables}
+\begin{itemize}
+\item Parental warmth ($X$): average of separate maternal and paternal warmth scales (so $\in [1,4]$)
+\item {Flourishing (main $Y$)
+  \begin{itemize}
+  \item Continuous version: Sum of standardized subscales for emotional, psychological, and social well-being
+  \item Binary version: Are you in top tertile on enough subscales?
+  \item Count version: For how many subscales are you in top tertile? ($\in [0,13]$?)
+  \end{itemize}
+}
+\item Subdomains comprising flourishing: 13 continuous measures (T3)
+\item Bad outcomes (secondary $Y$s): 7 mental health problems and bad behaviors, all binary (T4)
+\end{itemize}
+
+\subsection{Analyses}
+\begin{itemize}
+\item Continuous flourishing and other continuous $Y$s: Normal GEE (cluster by siblings)
+\item Binary flourishing: Poisson GEE with log link (risk ratios) (\hl{Did log-binomial not converge or other reason for choosing this?})
+\item Count flourishing: Normal GEE (\hl{Why not, e.g., Poisson or NB?})
+\item Bad binary outcomes: Poisson GEE for non-rare (RRs) or logistic regression for common (ORs)
+\end{itemize}
+
+\subsection{Our analyses}
+\begin{itemize}
+\item Multiple imputation for main analyses?
+\item Randomly choose 1 sibling from each sibship to avoid clustering? Then just use OLS and regular log-linear model. Or could do mixed model, but then parametric bootstrap is more complicated.
+\item Maybe treat outcomes as: 3 different flourishing variables + 7 binary bad outcomes? Also addresses model specification angle. 
+\item 
+\end{itemize}
+
 \newpage

### 2018-03-08 R code/For Sherlock/analysis.R

@@ -8,116 +8,78 @@ names(scen.params)[ names(scen.params) == "scen.name" ] = "scen"
 
 # just use a single file instead of stitched one
 s = read.csv("stitched.csv", header=TRUE)
+s = merge(s, scen.params, by = "scen" )
 
 # what percent done are we?
-nrow(s) / (1000 * nrow(scen.params))
+nrow(s) / (500 * nrow(scen.params))
 
 # how many reps per scenario do we have?
 table( s$scen )
 
-# remove columns from individual bootstrap results
-kill = c( names(s)[ grepl( "n.rej.bt", names(s) ) ], "n.rej.5e.04" )
-s = s[ , !names(s) %in% kill ]
-
-# is coverage different from 95%?
-mean(s$covers)
-mean(s$covers.correct)
-prop.test( sum(s$covers), length(s$covers), p=0.95 )
-
-# what happens we we only 
-s$sim.rep = rep( 1:(nrow(s)/2000), each = 2000 )
-first = s[ !duplicated( s$sim.rep ), ]
-mean(s$n.rej.0.05); mean(s$n.rej.bt.0.05)
-s$covers = 
 
 
-########################### MAKE DATASETS FOR FIRST PLOT (RESIDUAL RESAMPLING) ###########################
+########################### COMPARATIVE POWER PLOT ###########################
 
-# http://www.milanor.net/blog/reshape-data-r-tidyr-vs-reshape2/
+# need 1 row per scenario-method combination
+# methods are ours, Wstep, bonf.naive, and minP
+
+# variables to keep:
+# scenario parameters
+#  mean of jt.rej.bonf.naive, jt.rej.minP, jt.rej.Wstep
+#  mean of rej.jt.0.01, rej.jt.0.05
+
+library(reshape2)
+
+
+# 
+( jt.rej.names = c( names(s)[ grep( "jt.rej.", names(s) ) ],
+                    names(s)[ grep( "rej.jt", names(s) ) ] ) )
+  
+# MAKE SURE TO LIST THEM IN SAME ORDER AS JT.REJ.NAMES
+method.names = c("bonf.naive", "minP", "Wstep", "ours.0.01", "ours.0.05")
+
+# reshape wide to long
+# https://stackoverflow.com/questions/12466493/reshaping-multiple-sets-of-measurement-columns-wide-format-into-single-columns
+lp = reshape( s[ , names(s) %in% c(jt.rej.names, "scen") ],
+               varying = jt.rej.names,
[…]

### 2018-03-08 R code/For Sherlock/doParallel.R

@@ -11,13 +11,15 @@ setwd("/share/PI/manishad/multTest")
 scen.params = read.csv( "scen_params.csv" )
 p = scen.params[ scen.params$scen.name == scen, ]
 
+print(p)
+
 # no longer included in parameters because it's a vector
 crit.bonf = 0.05 / p$nY
 alpha = c( crit.bonf, 0.01, 0.05 )
 
 # simulation reps to run within this job
 # this need to match n.reps.in.doParallel in the genSbatch script
-sim.reps = 1
+sim.reps = 5
 ######### END OF CLUSTER PART #########
 
 
@@ -25,23 +27,26 @@ sim.reps = 1
 # # setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/Sandbox/2018-1-13")
 # # p = read.csv("scen_params.csv")  # should be a single row, I think
 # 
-# n = 500
+# n = 100
 # nX = 1
-# nY = 100
+# nY = 40
 # rho.XX = 0
-# rho.YY = c(0)
-# rho.XY = c(0.03)  # null hypothesis: 0
+# rho.YY = c(0.75)
+# rho.XY = c(0.08)  # null hypothesis: 0
+# half = 0
 # 
 # # bootstrap iterates and type
-# boot.reps = 200
-# sim.reps = 1
+# boot.reps = 5
+# sim.reps = 2
 # scen = "a"
 # bt.type = c( "h0.parametric" )
 # 
 # 
 # # matrix of scenario parameters
-# scen.params = expand.grid( bt.type, n, nX, nY, rho.XX, rho.YY, rho.XY )
-# names(scen.params) = c( "bt.type", "n", "nX", "nY", "rho.XX", "rho.YY", "rho.XY" )
+# scen.params = expand.grid( bt.type, n, nX, nY, rho.XX,
+#                            rho.YY, rho.XY, half )
+# names(scen.params) = c( "bt.type", "n", "nX", "nY", "rho.XX",
+#                         "rho.YY", "rho.XY", "half" )
 # 
 # # name the scenarios
 # # remove letters that are privileged variables in R
@@ -98,7 +103,8 @@ for ( j in 1:sim.reps ) {
 
   # make initial dataset from which to bootstrap
   cor = make_corr_mat( .nX = p$nX, .nY = p$nY, .rho.XX = p$rho.XX,
[…]

### 2018-03-08 R code/For Sherlock/gen_sbatch.R

@@ -10,12 +10,13 @@ setwd(path)
 # path = "~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/Sandbox/2018-1-13"
 # setwd(path)
 
-n = 500
+n = 1000
 nX = 1
-nY = 100
+nY = 40
 rho.XX = 0
-rho.YY = c(0.25)
-rho.XY = 0.03  # null hypothesis: 0
+rho.YY = c(0.25, 0.5, 0)
+rho.XY = c(0.02, 0.05, 0) # null hypothesis: 0
+half = c(0, 1)  # exchangeable vs. half-correlated matrix
 
 # n = 1000
 # nX = 1
@@ -25,13 +26,16 @@ rho.XY = 0.03  # null hypothesis: 0
 # rho.XY = c(0, 0.03, 0.08)  # null hypothesis: 0
 
 # bootstrap iterates and type
-boot.reps = 500
+boot.reps = 2000
 bt.type = c( "h0.parametric" )
 
+# ~~~ FOR NEXT TIME, DON'T HAVE SCENARIOS WITH HALF=0 AND HALF=1 WHEN 
+#  RHO.XY=0 BECAUSE SAME THING
+# SCENARIOS Q, R, S
 
 # matrix of scenario parameters
-scen.params = expand.grid( bt.type, n, nX, nY, rho.XX, rho.YY, rho.XY )
-names(scen.params) = c( "bt.type", "n", "nX", "nY", "rho.XX", "rho.YY", "rho.XY" )
+scen.params = expand.grid( bt.type, n, nX, nY, rho.XX, rho.YY, rho.XY, half )
+names(scen.params) = c( "bt.type", "n", "nX", "nY", "rho.XX", "rho.YY", "rho.XY", "half" )
 
 # name the scenarios
 # remove letters that are privileged variables in R
@@ -40,7 +44,7 @@ scen.params$scen.name = letter.names[ 1:dim(scen.params)[1] ]
 n.scen = length(scen.params[,1])
 
 # write the csv file of params (to Sherlock)
-write.csv( scen.params, "scen_params.csv" )
+# write.csv( scen.params, "scen_params.csv" )
 
 
 ########################### GENERATE SBATCHES ###########################
@@ -50,9 +54,8 @@ source("functions.R")
 
 
 # number of sbatches to generate (i.e., iterations within each scenario)
-# n.reps.per.scen = 1000
[…]

### 2018-03-11 Project log/multiple_outcomes_project_log.Rmd
About to add Romano

@@
 
-\subsection{FWER control methods we should compare for power of joint test}
+Mystery #2: Why does Westfall recommend a resampling scheme for logistic regression that doesn't preserve correlation? (pg 215)
+
+Multiple imputation idea: Maybe a general algorithm for resampling under $H_A$ could be something like MI.
+You would take your dataset and copy all the Xs, deleting the Ys.
+Then you would use an off-the-rack parametric MI procedure to impute the "missing data" for each of the resamples (and each time delete the observed data).
+E.g., could do a fully conditional approach where you first regression Y1 on all other variables, including the other outcomes, then regress Y2 on the remaining ones, etc.
+This does invoke assumptions about the form of relationship between the outcomes, but at least wouldn't break any parametric assumptions.
+I think you could do this with MICE's off-the-rack models because it fits logistic regression to binary variables by default (see JStatSoft paper)!
+To resample under $H_0$ this way, could first shuffle the Ys in original dataset?
+That should preserve correlation between Ys.
+
+\section{For Westfall simulations}
+
+* Redo as-is but with 1000 reps/scenario, and have more reps in doParallel
+* Add Holm to simulations -- maybe do it along with Romano?
+* Add Romano to simulations (requires FCR instead of Westfall approach) since it seems to work in practice (\texttt{debug\_fcr}), but remember that this violates OLS assumptions, so isn't technically advisable.
+
+\section{For later simulations}
+
+* Add binary outcomes with logistic regression?
+* To further improve power, could Use something about p-values themselves for joint test?
+
+\section{FWER control methods we should compare for power of joint test}
 
@@
 
-\subsection{Should we compare our joint test to joint tests based on non-FWER procedures?}
-
-If we reject joint null based on FWER control method, then we can conclude with 95\% confidence that this rejection is real, which implies that the joint null is false.
-
-With $k$-FWER control, we can conclude with 95\% confidence that there are fewer than $k$ false positives, so $R-k+1$ of the observed rejections are real (don't know which ones), where $R$ is the number of adjusted rejections.
-So, as long as $R>k-1$, we can reject the joint null with 95\% confidence.
-PROBLEM: WOULD HAVE TO TRY EVERY $k$??
-
-With FDP control, we can reject the joint null if the proportion of adjusted rejections exceeds the allowed amount ($\gamma$). Again, problem because we'd have to try every $\gamma$?
-
-FDR does not work for a joint test because it uses expected proportion of rejections.
-
-The problem is that $k$-FWER and FDP would probably lead to more powerful joint test than the FWER control method, but not clear which parameters to choose. So maybe we should not worry about them?
-
-\subsection{What we can learn from different procedures}
-
-Say we observe 18 rejections at the 0.05 level. Then, with increasing stringency, we can say:
-\begin{itemize}
-\item Our null CI is $[0, 10]$. "We observed $18-10 = 8$ more hits than we would expect in 95\% of samples taken under the strong null (every null holds)."
-\item $8-$FWER controlling procedure (Romano) rejects 11.
-"There are fewer than 8 false positives, so of these 14 rejections, at least $14-7 = 7$ are real effects, but we don't know which ones." (How should we decide which $k$ to use here?)
-\item FWER-controlling procedure (Bonferroni or preferably Romano/Westfall/another resampling method) rejects 3.
-"All 3 of these rejections are real effects with 95\% confidence", or alternatively: "We reject the strong null with >95\% confidence."
-\end{itemize}
-
-In practice, you could get the first two from the same bootstrap resamples (taken parametrically under the strong null).
-Then you'd have to resample under the original distribution to get the third one.
-
-\subsection{Our two metrics}
+\section{Our two metrics}
 
@@
 
-\subsection{New reference}
+\newpage
 
+\section{Resolved questions}
+
+\subsection{What should we do about logistic regression?}
+
+Westfall's algorithm (pg 215) doesn't seem to work.
+He just regresses each outcome on the Xs and then generates from a binomial using the fitted parameters.
+But this doesn't preserve the correlation between the Ys (see \texttt{westfall\_logistic\_regression.R}), so I don't understand.
+Also, he says you can't do case-resampling for logistic regression, but I don't know why -- unless it's the same issue about correlated residuals that seems potentially problematic in OLS.
+
+\subsection{Should we compare our joint test to joint tests based on non-FWER procedures?}
+
+If we reject joint null based on FWER control method, then we can conclude with 95\% confidence that this rejection is real, which implies that the joint null is false.
+
+With $k$-FWER control, we can conclude with 95\% confidence that there are fewer than $k$ false positives, so $R-k+1$ of the observed rejections are real (don't know which ones), where $R$ is the number of adjusted rejections.
+So, as long as $R>k-1$, we can reject the joint null with 95\% confidence.
+PROBLEM: WOULD HAVE TO TRY EVERY $k$??
+
+With FDP control, we can reject the joint null if the proportion of adjusted rejections exceeds the allowed amount ($\gamma$). Again, problem because we'd have to try every $\gamma$?
+
+FDR does not work for a joint test because it uses expected proportion of rejections.
+
+The problem is that $k$-FWER and FDP would probably lead to more powerful joint test than the FWER control method, but not clear which parameters to choose. So maybe we should not worry about them?
+
+\subsection{What we can learn from existing FWER procedures?}
+
+Say we observe 18 rejections at the 0.05 level. Then, with increasing stringency, we can say:
 \begin{itemize}
-\item asdf
+\item Our null CI is $[0, 10]$. "We observed $18-10 = 8$ more hits than we would expect in 95\% of samples taken under the strong null (every null holds)."
+\item $8-$FWER controlling procedure (Romano) rejects 11.
+"There are fewer than 8 false positives, so of these 14 rejections, at least $14-7 = 7$ are real effects, but we don't know which ones." (How should we decide which $k$ to use here?)
+\item FWER-controlling procedure (Bonferroni or preferably Romano/Westfall/another resampling method) rejects 3.
+"All 3 of these rejections are real effects with 95\% confidence", or alternatively: "We reject the strong null with >95\% confidence."
 \end{itemize}
 
-\newpage
-\section{Q \& A}
-\subsection{Open questions}
+In practice, you could get the first two from the same bootstrap resamples (taken parametrically under the strong null).
+Then you'd have to resample under the original distribution to get the third one.
 
-\subsection{Resolved questions}
-Q: Why do datasets created through full-case resampling reject the null too often?
-A: See my Cross Validated question (\url{https://stats.stackexchange.com/questions/323455/why-do-hypothesis-tests-on-resampled-datasets-reject-the-null-too-often}).
-Basically, when we resample full cases, we violate the OLS assumption of iid errors because every repeated pair $(X,Y)$ has exactly the same residual, so the residuals are highly correlated with $X$.
-If we fix $X$ and resample residuals, then even though there are repeated observations, the errors of the model newly fit to the resampled data are iid because we are breaking any association between the resampled residuals and $X$.
-Why do Davison \& Hinkley recommend this approach if it's wrong?
-Because it's only wrong if you need to meet parametric assumptions within each resample.
+\subsection{Why do datasets created through full-case resampling reject the null too often?}
+
+See my Cross Validated question (\url{https://stats.stackexchange.com/questions/323455/why-do-hypothesis-tests-on-resampled-datasets-reject-the-null-too-often}). There are two separate issues:
+
+\begin{enumerate}
+\item You have to center the bootstrap statistics by the estimated statistic in the original data (so do the hypothesis test on $\widehat{\beta}^{(j)} - \widehat{\beta}$, not on $\widehat{\beta}^{(j)}$ itself).
+This is because of the bootstrap analogy principle.
+\item With FCR, we violate the OLS assumption of iid errors because every repeated pair $(X,Y)$ has exactly the same residual, so the residuals are perfectly associated with $X$.
+If we use Westfall's \emph{null} resampling approach where we fix $X$ and resample residuals, then even though there are repeated observations, the \emph{residuals} of the model newly fit to the resampled data are still iid because we are breaking any association between the resampled residuals and $X$.
+This issue only matters if you need to meet parametric assumptions within each resample.
 Our application of bootstrapping is unusual in that we need to resample in a way that is compatible with parametric assumptions of the test we're going to apply for each outcome.
+This is why, e.g., Davison \& Hinkley do specifically recommend FCR for OLS.
+\end{enumerate}
 
-Q: Why does Blakesley/Westfall’s minP bootstrapping work since it is an extreme order statistic?
-A: Probably because minP is not an order statistic of the data themselves.
+It turns out that in the OLS example I tried, only issue #1 actually matter because fixing that issue led to nominal rejections.
+However, issue #2 seems to still be a real concern and might matter more in models less robust to assumption violations than OLS.
 
-Q: If we go the confidence interval route, why not just treat the p-values as our data and resample directly from them to produce CI? 
-A: Because there are only a fixed number of p-values (k), and this is not necessarily asymptotically large.
+\subsection{Why does Blakesley/Westfall’s minP bootstrapping work since it is an extreme order statistic?}
+Probably because minP is not an order statistic of the data themselves.
 
-Q: Why does bootstrapping work with small samples since we need the ECDF in sample to go to the true CDF? Or does it not work in small samples?
+\subsection{If we go the confidence interval route, why not just treat the p-values as our data and resample directly from them to produce CI?}
+Because there are only a fixed number of p-values (k), and this is not necessarily asymptotically large.
+
+\subsection{Why does bootstrapping work with small samples since we need the ECDF in sample to go to the true CDF? Or does it not work in small samples?}
 A: Correct. It does not necessarily work in small samples.
@@
 
-Next up: Try regenerating residuals on Sherlock with multiple outcomes and huge sample size. Trying n = 100,000 and B = 2,000 and 5-hr time limit?
+\subsection{Upcoming simulations}
 
-Seems like when rejection prob is low, bootstraps do worse? E.g., rho.XY = 0.03 worked perfectly with $n=10,000$, but rho.XY = 0.01 did not.
+\subsection{"2018-3-18 with other methods"}
+
+We shelved the CI under the alternative because of the below issues (couldn't find a bootstrapping/resampling method that worked under $H_A$).
+In the present simulations, we compared our joint test of the null (using number of rejections) to other methods (Westfall's minP, Westfall's step-down, Bonferroni).
+
+\begin{itemize}
+\item \textbf{Realized that regenerating residuals (method \texttt{h0.parametric}) actually doesn't work because it loses the correlated structure: we were just generating separate $Y$s.
+Empirically, saw that the null CIs were the same regardless of \texttt{rho.YY}.}
+\item So these simulations used Westfall's reattach-residuals method (method \texttt{resid}), and all results made a lot of sense. 
+\item CI plots shows the averaged CI limits and mean bootstrapped rejections by scenario. 
+\item Seemed weird that both Westfall methods lead to exactly the same joint test, but this is okay.
+It's because the smallest p-values are the ones that change the least, so few p-values shift from being above to below 0.05 when changing from minP to step-down (see folder "Compare Westfall minP vs step-down").
+\end{itemize}
 
@@
 
-Went back to debug fcr script and found that reattaching residuals works under null.
+Went back to debug FCR script and found that reattaching residuals works under null.
 

### 2018-03-11 R code/For Sherlock/analysis.R

@@ -1,7 +1,6 @@
 
 ########################### READ IN DATA ###########################
 
-# LOCAL TEST
 setwd("~/Desktop")
 scen.params = read.csv("scen_params.csv")
 names(scen.params)[ names(scen.params) == "scen.name" ] = "scen"
@@ -17,12 +16,6 @@ nrow(s) / (500 * nrow(scen.params))
 table( s$scen )
 
 
-# PROBLEMS
-#  POWER SHOULD DECREASE WITH INCREASING RHO.XY
-#  CIS SHOULD GET WIDER WITH INCREASING RHO.XY
-#  POINTS IN NULL CI SCREWED UP
-
-
 ########################### COMPARATIVE POWER PLOT ###########################
 
 # need 1 row per scenario-method combination
@@ -34,7 +27,7 @@ table( s$scen )
 #  mean of rej.jt.0.01, rej.jt.0.05
 
 library(reshape2)
-
+library(tidyverse)
 
 # names of joint rejection variables for all methods
 ( jt.rej.names = c( names(s)[ grep( "jt.rej.", names(s) ) ],
@@ -57,8 +50,6 @@ lp = reshape( s[ , names(s) %in% c(jt.rej.names, "scen") ],
 # nrow( lp[ lp$scen == "a", ] ); nrow( s[ s$scen == "a", ]) * length(method.names)
 
 # since each scenario still has multiple rows, take means 
-library(tidyr)
-
 pwr = lp %>% group_by(scen, method) %>%
   summarise( power = mean(jt.rej) )
 
@@ -72,13 +63,12 @@ pwr$group = paste( "X-Y correlation: ", pwr$rho.XY,
                    sep = "" )
 
 # for more plotting joy
-library(car)
-pwr$method.label = recode( pwr$method, 
-                           " 'bonf.naive' = 'NB';
-                           'minP' = 'MP'; 
-                           'ours.0.01' = 'J1';
-                           'ours.0.05' = 'J5';
-                           'Wstep' = 'WS' " )
+pwr$method.label = NA
+pwr$method.label[ pwr$method == "bonf.naive" ] = "NB"
+pwr$method.label[ pwr$method == "minP" ] = "MP"
+pwr$method.label[ pwr$method == "ours.0.01" ] = "J1"
+pwr$method.label[ pwr$method == "ours.0.05" ] = "J5"
+pwr$method.label[ pwr$method == "Wstep" ] = "WS"
 
 
[…]

### 2018-03-11 R code/For Sherlock/doParallel.R

@@ -124,10 +124,6 @@ for ( j in 1:sim.reps ) {
   pvals = samp.res$pvals
   names(n.rej) = paste( "n.rej.", as.character(alpha), sep="" )
   
-  # Bonferroni test of joint null using just original data
-  # i.e., do we reject at least one outcome using Bonferroni threshold?
-  jt.rej.bonf.naive = n.rej[1] > 0
-  
   # run all bootstrap iterates
     r = foreach( i = 1:boot.reps, .combine=rbind ) %dopar% {
       # draw bootstrap sample 
@@ -287,7 +283,7 @@ for ( j in 1:sim.reps ) {
     
     # p-values for observed rejections
     # if we resampled under H0, want prob of observing at least as many rejections as we did
-    if ( p$bt.type == "h0.parametric" ) {
+    if ( p$bt.type == "h0.parametric" | p$bt.type == "resid" ) {
       
       # jt.pval.bonf = sum( n.rej.bt.0.005 >= n.rej[,1] ) / length( n.rej.bt.0.005 )
       jt.pval.0.01 = sum( n.rej.bt.0.01 >= n.rej[["n.rej.0.01"]] ) /
@@ -300,6 +296,18 @@ for ( j in 1:sim.reps ) {
     # rej.jt.bonf = jt.pval.bonf < 0.05
     rej.jt.0.01 = jt.pval.0.01 < 0.05
     rej.jt.0.05 = jt.pval.0.05 < 0.05
+    
+    
+    
+    ######## Bonferroni joint test ########
+    
+    # Bonferroni test of joint null using just original data
+    # i.e., do we reject at least one outcome using Bonferroni threshold?
+    jt.rej.bonf.naive = n.rej[1] > 0
+    
+    
+    ######## Holm joint test ########
+    
     
 
     ######## Westfall's single-step ######## 
@@ -319,11 +327,17 @@ for ( j in 1:sim.reps ) {
     
     jt.rej.Wstep = any( p.adj.stepdown < 0.05 )
     
-    # compare to minP
-    # cbind( p.adj.minP, p.adj.stepdown )
     
-    # should reject more than either procedure above
-    # table( p.adj.stepdown < 0.05 )
+    ######## Romano ########
+    # DO ME
+    # Will need to run this with resampling scheme set to FCR
+    # See compare_joint_tests: I just need to center the boot tvals by the observed ones
+    
+    # regular FWER control
[…]

### 2018-03-11 R code/For Sherlock/gen_sbatch.R

@@ -18,24 +18,14 @@ rho.YY = c(0.25, 0.5, 0)
 rho.XY = c(0.02, 0.05, 0) # null hypothesis: 0
 half = c(0, 1)  # exchangeable vs. half-correlated matrix
 
-# n = 1000
-# nX = 1
-# nY = 100
-# rho.XX = 0
-# rho.YY = c(0, 0.25, 0.5, 0.75)
-# rho.XY = c(0, 0.03, 0.08)  # null hypothesis: 0
-
 # bootstrap iterates and type
 boot.reps = 2000
-bt.type = c( "h0.parametric" )
-
-# ~~~ FOR NEXT TIME, DON'T HAVE SCENARIOS WITH HALF=0 AND HALF=1 WHEN 
-#  RHO.XY=0 BECAUSE SAME THING
-# SCENARIOS Q, R, S
+bt.type = c( "resid" )
 
 # matrix of scenario parameters
 scen.params = expand.grid( bt.type, n, nX, nY, rho.XX, rho.YY, rho.XY, half )
 names(scen.params) = c( "bt.type", "n", "nX", "nY", "rho.XX", "rho.YY", "rho.XY", "half" )
+
 
 # name the scenarios
 # remove letters that are privileged variables in R
@@ -43,8 +33,14 @@ letter.names = c(letters, LETTERS)[ ! c(letters, LETTERS) %in% c("i","T","F") ]
 scen.params$scen.name = letter.names[ 1:dim(scen.params)[1] ]
 n.scen = length(scen.params[,1])
 
+# we don't need scenarios with rho.XY = 0 and half = 1 because redundant with 
+# scenarios where rho.XY = 0 and half = 0
+scen.to.toss = scen.params$scen.name[ scen.params$rho.XY == 0 & scen.params$half == 1 ]
+
+scen.params = scen.params[ ! scen.params$scen.name %in% scen.to.toss, ]
+
 # write the csv file of params (to Sherlock)
-# write.csv( scen.params, "scen_params.csv" )
+write.csv( scen.params, "scen_params.csv" )
 
 
 ########################### GENERATE SBATCHES ###########################
@@ -72,7 +68,7 @@ runfile_path = paste(path, "/testRunFile.R", sep="")
 sbatch_params <- data.frame(jobname,
                             outfile,
                             errorfile,
-                            jobtime = "5:30:00",
+                            jobtime = "0:30:00",
                             quality = "normal",
                             node_number = 1,
[…]

### 2018-03-11 R code/For Sherlock/stitch_on_sherlock.R

@@ -17,9 +17,11 @@ write.csv( s, "stitched.csv" )
 
 # how close are we to being done?
 (nrow(s)/5) / 1500
+table(s$scen)
 
 # move it to Desktop
-# scp mmathur@sherlock:/share/PI/manishad/multTest/sim_results/stitched.csv ~/Desktop
+scp mmathur@sherlock:/share/PI/manishad/multTest/sim_results/stitched.csv ~/Desktop
+scp mmathur@sherlock:/share/PI/manishad/multTest/scen_params.csv ~/Desktop
 
 
 

### 2018-03-12 R code/For Sherlock/doParallel.R
Romano & Holm ready for first cluster test

@@ -31,7 +31,7 @@ sim.reps = 5
 # nX = 1
 # nY = 40
 # rho.XX = 0
-# rho.YY = c(0.75)
+# rho.YY = c(0.25)
 # rho.XY = c(0.08)  # null hypothesis: 0
 # half = 0
 # 
@@ -39,7 +39,7 @@ sim.reps = 5
 # boot.reps = 5
 # sim.reps = 2
 # scen = "a"
-# bt.type = c( "h0.parametric" )
+# bt.type = c( "resid" )
 # 
 # 
 # # matrix of scenario parameters
@@ -119,9 +119,11 @@ for ( j in 1:sim.reps ) {
                                               TRUE, FALSE), 
                              .sigma = ifelse( p$bt.type %in% c( "h0.parametric", "ha.resid.2"),
                                               TRUE, FALSE),
-                             .intercept = (p$bt.type == "h0.parametric") )
+                             .intercept = (p$bt.type == "h0.parametric"),
+                             .tval = TRUE)
   n.rej = samp.res$rej  # first one is Bonferroni
   pvals = samp.res$pvals
+  tvals = samp.res$tvals
   names(n.rej) = paste( "n.rej.", as.character(alpha), sep="" )
   
   # run all bootstrap iterates
@@ -200,7 +202,7 @@ for ( j in 1:sim.reps ) {
       
       
       ##### Bootstrap Under HA #3 - Regenerate Residuals #####
-      # ~~~~ I THINK THIS ALSO LOSES THE CORRELATION???
+      # THIS ACTUALLY LOSES CORRELATION BETWEEN YS
       # re-attach residuals
       if ( p$bt.type == "ha.resid.2" ) {
         # extract residuals from original data
@@ -228,11 +230,15 @@ for ( j in 1:sim.reps ) {
       # important for dataset_result to be able to find the right names
       names(b) = c(X.names, Y.names)
       
-      bt.res = dataset_result( .dat = b, .alpha = alpha, .resid = FALSE, .tval=TRUE )
+      bt.res = dataset_result( .dat = b,
+                               .alpha = alpha,
+                               .resid = FALSE,
+                               .tval=TRUE )
       
       # return all the things
       list( rej = bt.res$rej,
-            pvals = bt.res$pvals )
+            pvals = bt.res$pvals,
+            tvals = bt.res$tvals )
 
       # get number rejected for bootstrap sample
       # we don't need to return residuals for this one
[…]

### 2018-03-12 R code/For Sherlock/doParallel_mice.R
Work on fns for MICE approach for logistic regression

@@ -1,0 +1,333 @@
+
+
+######### FOR CLUSTER USE #########
+# load command line arguments
+args = commandArgs(trailingOnly = TRUE)
+jobname = args[1]
+scen = args[2]  # this will be a letter
+boot.reps = as.numeric( args[3] )
+
+# get scen parameters
+setwd("/share/PI/manishad/multTest")
+scen.params = read.csv( "scen_params.csv" )
+p = scen.params[ scen.params$scen.name == scen, ]
+
+print(p)
+
+# no longer included in parameters because it's a vector
+crit.bonf = 0.05 / p$nY
+alpha = c( crit.bonf, 0.01, 0.05 )
+
+# simulation reps to run within this job
+# this need to match n.reps.in.doParallel in the genSbatch script
+sim.reps = 5
+######### END OF CLUSTER PART #########
+
+
+# ######### FOR LOCAL USE #########
+# # setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/Sandbox/2018-1-13")
+# # p = read.csv("scen_params.csv")  # should be a single row, I think
+#
+# n = 100
+# nX = 1
+# nY = 40
+# rho.XX = 0
+# rho.YY = c(0.25)
+# rho.XY = c(0.08)  # null hypothesis: 0
+# half = 0
+#
+# # bootstrap iterates and type
+# boot.reps = 5
+# sim.reps = 2
+# scen = "a"
+# bt.type = c( "resid" )
+#
+#
+# # matrix of scenario parameters
+# scen.params = expand.grid( bt.type, n, nX, nY, rho.XX,
+#                            rho.YY, rho.XY, half )
+# names(scen.params) = c( "bt.type", "n", "nX", "nY", "rho.XX",
+#                         "rho.YY", "rho.XY", "half" )
+#
+# # name the scenarios
+# # remove letters that are privileged variables in R
+# letter.names = c(letters, LETTERS)[ ! c(letters, LETTERS) %in% c("i","T","F") ]
+# scen.params$scen.name = letter.names[ 1:dim(scen.params)[1] ]
+# p = scen.params
+#
+# # add alpha corresponding to Bonferroni as first one
+# # NOTE THAT CODE ASSUMES BONFERRONI IS THE FIRST ONE
+# crit.bonf = 0.05 / p$nY
+# alpha = c( crit.bonf, 0.01, 0.05 )
+#
[…]

### 2018-03-12 R code/For Sherlock/functions_mice.R

@@ -1,0 +1,631 @@
+
+
+
+######################## FNS FOR WESTFALL's SINGLE-STEP ########################
+
+# returns minP-adjusted p-values (single-step)
+
+# p = original p-values
+# p.bt = bootstrapped p-values (vector) - an m X B matrix
+adjust_minP = function( p, p.bt ) {
+  
+  n.boot = ncol(p.bt)
+  
+  # keep only minimum p-value in each resample
+  minP.bt = apply( p.bt, MARGIN = 2, FUN = min )
+  
+  # for each element of p, get the proportion of resamples
+  #  whose minP was less than the present p-value
+  p.adj = unlist( lapply( p, FUN = function(x) sum( minP.bt <= x ) / n.boot ) )
+  
+  return(p.adj)
+}
+
+# # test
+# B = 200
+# n.tests = 10
+# p.bt = matrix( runif(B*n.tests, 0, 1), nrow = n.tests)
+# p = runif( n.tests, 0, .1)
+# 
+# p.adj = adjust_minP( p, p.bt )
+# plot(p,p.adj)
+
+
+######################## FNS FOR WESTFALL's STEP-DOWN ########################
+
+# Westfall textbook, pages 66-67
+adj_Wstep = function( p, p.bt ) {
+  
+  # # TEST ONLY
+  # p = structure(c(0.636112475277457, 0.116408752390887, 0.260603159221029,
+  #                 0.575086408582202, 0.625789849782125), .Dim = c(5L, 1L))
+  # p.bt = structure(c(0.728140113074014, 0.828557135138776, 0.728464927503583,
+  #                    0.376518353153738, 0.241363541131996, 0.415330227654616, 0.75686321503498,
+  #                    0.865494617532842, 0.839819273240726, 0.540707408457488, 0.149482739547785,
+  #                    0.506274953008223, 0.122605643925883, 0.747018941160177, 0.168076477020114,
+  #                    0.0582566955778811, 0.284459435324328, 0.561050485854955, 0.13839444320949,
+  #                    0.641335055023296, 0.195313599009888, 0.174692123827015, 0.982035588216878,
+  #                    0.713525352129964, 0.548770486028363), .Dim = c(5L, 5L), .Dimnames = list(
+  #                      NULL, c("result.1", "result.2", "result.3", "result.4", "result.5"
+  #                      )))
+  
+  # if using read-in or simulated dat
+  # p = pvals
+  
+  
+  # attach indices to original p-values
[…]

### 2018-03-13 R code/For Sherlock/functions_mice.R
MICE code ready for Sherlock after surviving local test

@@ -240,14 +240,45 @@ make_corr_mat = function( .nX, .nY, .rho.XX, .rho.YY, .rho.XY, .half = FALSE ) {
 # simulate 1 dataset with correlated covariates
 # then simulate 
 
-sim_data = function( .n, .cor ) {
+sim_data = function( .n, .cor,
+                     .Ytype = "normal",
+                     .nX = NULL, .nY = NULL ) # only needed for binary Ys
+  {
 
   vnames = names( .cor )
   
+  # names for variables
+  X.names = paste0( "X", 1 : .nX )
+  Y.names = paste0( "Y", 1 : .nY )
+  
   # simulate the dataset
-  # everything is a standard Normal
-  library(mvtnorm)
-  d = as.data.frame( rmvnorm( n = .n, mean = rep( 0, dim( .cor )[1] ), sigma = as.matrix( .cor ) ) )
+  if ( .Ytype == "normal" ) {
+    # everything is a standard Normal
+    library(mvtnorm)
+    d = as.data.frame( rmvnorm( n = .n, mean = rep( 0, dim( .cor )[1] ), sigma = as.matrix( .cor ) ) )
+    vnames = c( X.names, Y.names )
+  }
+
+  # simulate the dataset
+  else if ( .Ytype == "binary" ) {
+    library(BinNor)
+    sigma.star = compute.sigma.star( no.bin = .nY, no.nor = .nX, prop.vec.bin = rep(0.5, .nY),
+                                   corr.mat = as.matrix(.cor) )
+    
+    d = as.data.frame( 
+      jointly.generate.binary.normal( no.rows = .n, no.bin = .nY, no.nor = .nX,
+                                      prop.vec.bin = rep(0.5, .nY),
+                                      mean.vec.nor = 0,
+                                      var.nor = 1, 
+                                      sigma_star = sigma.star$sigma_star,
+                                      continue.with.warning = TRUE )
+    )
+    
+    # respect ordering from BinNor
+    # opposite of mvrnorm case
+    vnames = c( Y.names, X.names )
+  }
+  
   names(d) = vnames
   
   # return the dataset
@@ -281,11 +312,11 @@ fit_model = function( Y.name, .dat, .resid, .sigma, .tval, .intercept ) {
   
   # ~~~ NEW STUFF
   # is Y binary? Should work for either 0/1 coding or factor coding
-  binary = ( length( unique( .dat[[Y.name]] ) ) == 2 )
[…]

### 2018-03-13 R code/For Sherlock/gen_sbatch_mice.R

@@ -1,0 +1,94 @@
+
+
+########################### SET SIMULATION PARAMETERS MATRIX ###########################
+
+# FOR CLUSTER USE
+path = "/share/PI/manishad/multTest"
+setwd(path)
+
+# FOR LOCAL USE
+# path = "~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/Sandbox/2018-1-13"
+# setwd(path)
+
+n = 1000
+nX = 1
+nY = 40
+rho.XX = 0
+rho.YY = c(0.25, 0.5, 0)
+rho.XY = c(0.02, 0.05, 0) # null hypothesis: 0
+half = c(0, 1)  # exchangeable vs. half-correlated matrix
+
+# bootstrap iterates and type
+boot.reps = 2000
+bt.type = c( "MICE.H0" )
+
+# matrix of scenario parameters
+scen.params = expand.grid( bt.type, n, nX, nY, rho.XX, rho.YY, rho.XY, half )
+names(scen.params) = c( "bt.type", "n", "nX", "nY", "rho.XX", "rho.YY", "rho.XY", "half" )
+
+
+# name the scenarios
+# remove letters that are privileged variables in R
+letter.names = c(letters, LETTERS)[ ! c(letters, LETTERS) %in% c("i","T","F") ]
+scen.params$scen.name = letter.names[ 1:dim(scen.params)[1] ]
+n.scen = length(scen.params[,1])
+
+# we don't need scenarios with rho.XY = 0 and half = 1 because redundant with 
+# scenarios where rho.XY = 0 and half = 0
+scen.to.toss = scen.params$scen.name[ scen.params$rho.XY == 0 & scen.params$half == 1 ]
+
+scen.params = scen.params[ ! scen.params$scen.name %in% scen.to.toss, ]
+
+# write the csv file of params (to Sherlock)
+write.csv( scen.params, "scen_params.csv" )
+
+
+########################### GENERATE SBATCHES ###########################
+
+# load functions for generating sbatch files
+source("functions.R")
+
+
+# number of sbatches to generate (i.e., iterations within each scenario)
+n.reps.per.scen = 500
+n.reps.in.doParallel = 5
+n.files = ( n.reps.per.scen / n.reps.in.doParallel ) * n.scen
+
+
+path = "/share/PI/manishad/multTest"
+
+scen.name = rep( scen.params$scen.name, each = ( n.files / n.scen ) )
[…]

### 2018-05-10 Manuscript/manuscript.Rmd
About to revamp theory in manuscript

@@
 author: ''
+output:
+  pdf_document: default
+  word_document: default
 header-includes:
@@
 - \fancyfoot{}
-- \fancyhead[C]{Multiple Outcomes}
+- \fancyhead[C]{Multiple Testing}
 - \fancyfoot[RO,LE]{\thepage}
@@
 - \usepackage{float}
-output: pdf_document
-bibliography: refs_fake.bib
-csl: aje_style.csl
+bibliography: refs_mo.bib
+csl: apa_style.csl
 ---
@@
 \newtheorem{example}{Example}
-\newtheorem{algorithm}{Algorithm}
+\newtheorem{assump}{Assumption}
 
-<!--
 \doublespacing
@@
 \begin{center}
-\textbf{ \LARGE{Title line one:} } \\ \vspace{5mm}
-\textbf{ \LARGE{title line two} }
+\textbf{ \LARGE{Multiple Testing Research Update \#1} } \\ \vspace{5mm}
+\textbf{ \LARGE{\today} }
 \vspace{10mm}
@@
 
+\tableofcontents
+
+<!--
 \vspace{10mm}
@@
 
-\newpage
 -->
 
-\section{Notes to self}
+\newpage
 
@@
 Let $Z_{nk}$ denote, for the $n^{th}$ subject, the $k^{th}$ random variable.
-Consider a bootstrapping algorithm (which is defined below) that generates, for iterate $j$, a dataset containing the random variables $\left( Z_{n1}^{(j)}, \dotsb, Z_{nK}^{(j)} \right)$ for each subject $n$.
-There are a total of $B$ bootstrapped datasets.
+Consider a resampling algorithm that generates, for iterate $j$, a dataset containing the random vector $\left( Z_{n1}^{(j)}, \dotsb, Z_{nK}^{(j)} \right)$ for each subject $n$.
+There are a total of $B$ resampled datasets.
+We use the superscript $^{(j)}$ to denote random variables, distributions, and statistics in resampled dataset $j$.
 
-We refer to probability density functions as $f$ and cumulative distribution functions as $F$.
+Further suppose that we conduct $W$ hypothesis tests, each with level $\alpha$.
+Denote the $w^{th}$ null hypothesis as $H_{0w}$.
+Let $c_{w,\alpha}$ be the critical value for the test statistic, $T_w$, of the $w^{th}$ test.
+We refer to the vector of test statistics as $\mathbf{T}$.
+<!--Assume that each test concerns a point null hypothesis (\textbf{Assumption 1}) and that each test is parametric with consistent estimates of the tested parameters in the likelihood identifiable from the observed data (\textbf{Assumption 2}).--> We define the "strong null" as the case in which all $W$ point null hypotheses hold and use the superscript $^0$ generally to denote distributions, data, or statistics generated under the strong null.
+
+Define the statistic corresponding to the total number of rejections in $W$ tests in a sample generated under the strong null as $\widehat{\theta}^0 = \sum_{w=1}^W 1 \big\{ T_w^0 > c_{w,\alpha} \big\}$.
+Similarly, the total number of rejections in resample $j$ is $\widehat{\theta}^{(j)} = \sum_{w=1}^W 1 \big\{ T_w^{(j)} > c_{w,\alpha} \big\}$.
+Using $F$ to denote cumulative distribution functions, respectively define the true CDF of the number of rejections under the strong null, its counterpart in the resamples, and its empirical estimator in the resamples as:
+\begin{align*}
+F_{\widehat{\theta}^0}(r) &= P \left( \widehat{\theta}^0 \le r \right) \\
+F_{\widehat{\theta}^{(j)}}(r) &= P \left( \widehat{\theta}^{(j)} \le r \right) \\
+\widehat{F}_{\widehat{\theta}^{(j)}} \left( r \right) &= \frac{1}{B} \sum_{j^{*}=1}^B 1 \Big\{ \widehat{\theta}^{(j^{*})} \le r \Big\}
+\end{align*}
+
+<!----- Old version
+\section{Setting and notation}
+
+Suppose that $K$ random variables are measured on $N$ subjects, with the resulting matrix denoted $\mathbf{Z} \in \mathbb{R}_{N \times K}$.
+Let $Z_{nk}$ denote, for the $n^{th}$ subject, the $k^{th}$ random variable.
+Consider a resampling algorithm that generates, for iterate $j$, a dataset containing the random vector $\left( Z_{n1}^{(j)}, \dotsb, Z_{nK}^{(j)} \right)$ for each subject $n$.
+There are a total of $B$ resampled datasets.
+
+We refer to cumulative distribution functions as $F$.
 Specifically, we refer to the joint distribution of the data as follows.
@@
 
-<!-- Define the joint CDF of $\left( Z_{n1}^{(j)}, \dotsb, Z_{nK}^{(j)} \right)$ as:
+Define the joint CDF of $\left( Z_{n1}^{(j)}, \dotsb, Z_{nK}^{(j)} \right)$ as:
 
@@
 \end{align*}
--->
 
-We will usually refer simply to $\mathbf{Z}$ and $F_{\mathbf{Z}}$ for brevity and will denote their counterparts in bootstrapped dataset $j$ as $\mathbf{Z}^{(j)}$ and $F_{\mathbf{Z}^{(j)}}$.
+We will usually refer simply to $\mathbf{Z}$ and $F_{\mathbf{Z}}$ for brevity and will denote their counterparts in resampled dataset $j$ as $\mathbf{Z}^{(j)}$ and $F_{\mathbf{Z}^{(j)}}$.
 
@@
 We refer to the vector of test statistics as $\mathbf{T}$.
-Assume that each test concerns a point null hypothesis (\textbf{Assumption 1}), that each test statistic is a continuous function of the data with continuous first derivatives, $\mathbf{Z}$ (\textbf{Assumption 2}), and that each test is parametric with consistent estimates of all parameters in the likelihood identifiable from the observed data (\textbf{Assumption 3}).
-We define the "strong null" as the case in which all $W$ point null hypotheses hold and use the superscript $^0$ to denote variables generated under the strong null.
+Assume that each test concerns a point null hypothesis (\textbf{Assumption 1}) and that each test is parametric with consistent estimates of the tested parameters in the likelihood identifiable from the observed data (\textbf{Assumption 2}).
+We define the "strong null" as the case in which all $W$ point null hypotheses hold and use $F_{ \mathbf{Z}^0 }$ to denote the distribution of the data under the strong null.
+We use the superscript $^0$ generally to denote data or statistics generated under the strong null.
 
 Define the statistic corresponding to the total number of rejections in $W$ tests in a sample generated under the strong null as $\widehat{\theta}^0 = \sum_{w=1}^W 1 \big\{ T_w^0 > c_{w,\alpha} \big\}$.
-Similarly, the total number of rejections in bootstrap sample $j$ is $\widehat{\theta}^{(j)} = \sum_{w=1}^W 1 \big\{ T_w^{(j)} > c_{w,\alpha} \big\}$.
+Similarly, the total number of rejections in resample $j$ is $\widehat{\theta}^{(j)} = \sum_{w=1}^W 1 \big\{ T_w^{(j)} > c_{w,\alpha} \big\}$.
 
-Respectively define the true distribution of the number of rejections under the strong null, its counterpart in the bootstrap samples, and its empirical estimator in the bootstrap samples as:
+Respectively define the true distribution of the number of rejections under the strong null, its counterpart in the resamples, and its empirical estimator in the resamples as:
 \begin{align*}
@@
 
+------->
+
 <!------------- --------------------------------->
@@
 
-\section{Results}
+\section{Main results}
 
-We now show that under a parametric bootstrapping algorithm defined below, the empirical distribution of the number of rejections in the bootstrap samples converges in distribution to the true distribution of the number of rejections in sample generated under the strong null.
+We now show that under a certain class of resampling algorithms defined below, the empirical distribution of the number of rejections in the resamples converges to the true distribution of the number of rejections in samples generated under the strong null.
 Figure \ref{roadmap} depicts the structure of the proofs.
@@
 \caption{Overview of theory to follow.
-The dashed arrows denote convergence in probability in either $N$ or both $N$ and $B$.
-The solid arrow denotes almost sure convergence in $B$.
-Parenthetical acronyms refer to remarks, a lemma, and a theorem developed below.}
+Solid arrows denote ordinary limits in $B$, the dashed line denotes almost sure convergence in $N$, and the dotted line denotes convergence in probability in both $N$ and $B$.
+Parenthetical acronyms refer to an assumption, remark, lemma, and theorem developed below.}
 \label{roadmap}
@@
   % the - argument removes the arrowhead
-    (m-1-1) edge[dashed] node [below] {(R1)} (m-1-2)
+    (m-1-1) edge[] node [below] {(A3)} (m-1-2)
 
@@
   % the - argument removes the arrowhead
-    (m-1-1) edge[dashed] node [below] {(L1)} (m-1-2)
+    (m-1-1) edge[] node [below] {(L1)} (m-1-2)
 
     % diagonal line
-	(m-2-1) edge[] node [left] {(R2)} (m-1-1)
+	(m-2-1) edge[dashed] node [left] {(R2)} (m-1-1)
 
-    (m-2-1) edge [dashed] node [below] {(T1)} (m-1-2)
+    (m-2-1) edge [dotted] node [below] {(T1)} (m-1-2)
             ;
@@
 <!-------------- Definition 1 --------------->
-\bigskip \begin{algorithm}[Parametric bootstrap under the strong null] 
+<!-------------- \setcounter{assump}{2} --------------->
+\bigskip \begin{assump}[Resampling under the strong null] 
 \label{cbt_def}
-We will refer throughout to bootstrap samples generated as follows:
-\begin{enumerate}
-\item Use the observed data to obtain consistent point estimates of any unknown parameters in the likelihood. 
-\item For any parameters in the likelihood that are constrained under the null hypothesis, set them equal to their null values (which are uniquely defined under Assumption 1 regarding point null hypotheses).
-For nuisance parameters, set them equal to their consistent estimates.
-\item For any covariates treated as fixed in the likelihood, set them equal to their observed values.
-\item Generate outcome data from the estimated likelihood under the strong null as defined above. 
-\end{enumerate}
-\end{algorithm}
+The resampling algorithm used to generate $\left( Z_{n1}^{(j)}, \dotsb, Z_{nK}^{(j)} \right)$ must generate identically and independently distributed (i.i.d.) observations, and it must ensure that $\mathbf{T}^{(j)} \xrightarrow[B \to \infty]{D} \mathbf{T}^0$, i.e.:
+$$\lim_{B \to \infty} F_{\mathbf{T}^{(j)}} = F_{\mathbf{T}^0}$$
+\end{assump}
 
-(\hl{CITE ME - e.g., Fox, Romano}.)
-Note that unlike some resampling schemes that can induce \hl{XXX} problems, this algorithm guarantees that the bootstrapped data satisfy all parametric assumptions of the hypothesis tests to be conducted, preserving the nominal $\alpha$ of each test.
+This requirement has several critical implications.
+First, the resampled data must preserve the correlation structure of all variables in the dataset, except where the strong null dictates otherwise.
+Otherwise, the distribution of the test statistics will typically not be preserved.
+Second, just as the original data are assumed to respect the parametric assumptions of all $W$ hypothesis tests, the resampled data must be generated in a manner that also respects this parametric structure.
+Otherwise, hypothesis tests conducted on the resampled data may not preserve their nominal $\alpha$-levels, which again affects the distribution of the test statistics.
 
-<!-------------- Example 1 --------------->
+<!-------------- Example 1 ------------->
 \bigskip
-\begin{example}[Parametric bootstrap for multiple linear regression] 
-Suppose that one of the hypothesis tests is conducted using an OLS model regressing the $w^{th}$ variable on the first variable, adjusting for the second variable:
-$$Z_{nw} = \beta_0 + \beta_1 Z_{n1} + \beta_2 Z_{n2} + \epsilon_{nw}$$
-with $H_0: \beta_1 = 0$ and $\epsilon \sim N(0, \sigma^2)$. Then we can apply Algorithm \ref{cbt_def} as follows:
-\begin{enumerate}
-\item Fit the model to the observed data to estimate the three nuisance parameters as $\widehat{\beta}_0$, $\widehat{\beta}_2$, and $\widehat{\sigma}^2$. 
-\item For each bootstrap iterate, set the design matrix equal to that of the observed data.
-\item Set $\beta_1 = 0$ (invoking the strong null), and set nuisance parameters $\beta_0 = \widehat{\beta}_0$, $\beta_2 = \widehat{\beta}_2$, and $\sigma^2 = \widehat{\sigma}^2$. 
-\item Thus, for each bootstrap iterate, generate the bootstrapped outcomes as $Z_{nw}^{(j)} = \widehat{\beta}_0 + \widehat{\beta}_2 Z_{n2} + \epsilon^{(j)}_{nw}$, where $\epsilon^{(j)}_{nw} \sim N(0,\widehat{\sigma}^2)$.
-\end{enumerate}
+\begin{example}[A valid resampling algorithm for multiple linear regression] 
+\label{ex:ols}
+\normalfont
+Suppose we conduct $W=2$ hypothesis tests using ordinary least squares (OLS) regressions of the first and second variables on the third variable, adjusting for the fourth:
+\begin{align*}
+Z_{n1} &= \beta_0 + \beta_1 Z_{n3} + \beta_2 Z_{n4} + \epsilon_{n1}\\
+Z_{n2} &= \alpha_0 + \alpha_1 Z_{n3} + \alpha_2 Z_{n4} + \epsilon_{n2}
+\end{align*}
+with null hypotheses $H_{01}: \beta_1 = 0$ and $H_{02}: \alpha_1 = 0$, and assuming $\epsilon_{nw} \sim N(0, \sigma^2_w)$ for $w \in \{1,2\}$.
+A resampling algorithm \hl{[@westfalltext]} satisifying Assumption \ref{cbt_def} could proceed by first fixing the covariates $Z_{n3}$ and $Z_{n4}$ for all observations while setting the "outcomes" equal to a vector of residuals resampled with replacement.
+That is, letting $n'$ denote an observation sampled with replacement and letting hats denote the usual OLS estimates obtained from the original dataset, the resampled variables for observation $n$ are:
+\begin{align*}
+Z_{n1}^{(j)} &:= \left( \widehat{Z}_{n'1} - Z_{n'1} \right)\\
+Z_{n2}^{(j)} &:= \left( \widehat{Z}_{n'2} - Z_{n'2} \right)\\
+Z_{n3}^{(j)} &:= Z_{n3}\\
+Z_{n4}^{(j)} &:= Z_{n4}
+\end{align*}
+This resampling algorithm fulfills Assumption \ref{cbt_def} (\hl{proof in progress}).
+Note that even if the original data were not generated under the strong null, this resampling algorithm guarantees that $E \big[ Z_{nw}^{(j)} \; \vert \; Z_{n3} \big] = E \big[ Z_{nw}^{(j)} \big]$ for $w \in \{1,2\}$, implying the strong null.
+This algorithm also retains the distributional assumption on $\epsilon_{n1}$ and $\epsilon_{n2}$.
 \end{example}
 
-<!-------------- Example 2 --------------->
+<!-------------- Example 2  ------------->
 \bigskip
-\begin{example}[Parametric bootstrap for linear mixed model] 
-The approach in Definition \ref{cbt_def} can accommodate parametric models with correlated observations, such as a linear mixed model with random intercepts by subject.
-One could obtain consistent estimates of the random intercepts using, for example, empirical Bayesian estimation, and these estimates can be plugged into the estimated likelihood under the strong null.
+\begin{example}[An invalid resampling approach for multiple linear regression]
+\normalfont
+We now return to the setting of Example \ref{ex:ols}.
+An alternative resampling approach that is intuitive, but in fact violates Assumption \ref{cbt_def}, is to again fix $Z_{n3}$ and $Z_{n4}$ but to resample with replacement the outcome vectors $\left( Z_{n'1}, Z_{n'2} \right)$ rather than the residuals:
+\begin{align*}
+Z_{n1}^{(j)} &:= {Z}_{n'1}\\
+Z_{n2}^{(j)} &:= {Z}_{n'2}\\
+Z_{n3}^{(j)} &:= Z_{n3}\\
+Z_{n4}^{(j)} &:= Z_{n4}
+\end{align*}
+Although this approach indeed enforces the strong null and preserves the correlation between the outcomes, it fails to preserve the correlation between the outcomes and the adjusted covariate, $Z_{n4}$, and thus does not recover the distribution of $\left( \widehat{\beta}_1^0, \widehat{\alpha}_1^0 \right)$.
+\end{example}
+
+<!-------------- Example 2 ------------->
+\bigskip
+\begin{example}[Another invalid resampling algorithm for multiple linear regression] 
+\normalfont
+Another incorrect alternative would be to bootstrap parametrically while constraining $\beta_1 = \alpha_1 = 0$ to enforce the strong null:
+\begin{align*}
+Z_{n1}^{(j)} &:= \widehat{\beta}_0 + \widehat{\beta}_2 Z_{n4} + \epsilon_{n1}^{(j)}\\
+Z_{n2}^{(j)} &:= \widehat{\alpha}_0 + \widehat{\alpha}_2 Z_{n4} + \epsilon_{n2}^{(j)}\\
+Z_{n3}^{(j)} &:= Z_{n3}\\
+Z_{n4}^{(j)} &:= Z_{n4}
+\end{align*}
+where $\epsilon_{nw}^{(j)} \sim_{i.i.d.} N(0, \widehat{\sigma}^2_{\epsilon_{nw}})$ for $w \in \{ 1,2\}$.
+However, this sequential algorithm fails to preserve the correlation between $Z_{n1}$ and $Z_{n2}$, in turn failing to recover the distribution of $\left( \widehat{\beta}_1^0, \widehat{\alpha}_1^0 \right)$.
 \end{example}
@@
 \label{CMT_results}
-The distribution of the bootstrapped test statistics is consistent in $N$ for the true distribution of the test statistics under the strong null.
-By repeatedly invoking the Continuous Mapping Theorem (CMT):
-\begin{align}
-\label{consistent}
-f_{\mathbf{Z}^{(j)}} &\xrightarrow[N \to \infty]{P} f_{\mathbf{Z}^0} \tag{CMT; Assumption 3} \\
-f_{\mathbf{T}^{(j)}} &\xrightarrow[N \to \infty]{P} f_{\mathbf{T}^0} \tag{CMT; Assumption 2} \\
-F_{\mathbf{T}^{(j)}} &\xrightarrow[N \to \infty]{P} F_{\mathbf{T}^0} \tag{CMT; integration preserves continuity} \\
-\end{align}
+\normalfont
+For Assumption \ref{cbt_def} to hold, it is sufficient for $\mathbf{T}$ to be a continuous function of $\mathbf{Z}$ and for $\mathbf{Z}^{(j)} \xrightarrow[B \to \infty]{D} \mathbf{Z}^0$, i.e., $\lim_{B \to \infty} F_{\mathbf{Z}^{(j)}} = F_{\mathbf{Z}^0}$.
+Note that this condition is \emph{not} necessary for Assumption \ref{cbt_def} to hold; for example, under the residual resampling scheme of Example \ref{ex:ols}, we have $E \big[ Z_{n1}^{(j)} \big] = E \big[Z_{n2}^{(j)} \big] = 0$ regardless of $E \big[ Z_{n1} \big]$ and $E \big[Z_{n2}\big]$.
 \end{remark}
@@
 \label{glivenko}
-By the i.i.d construction of the bootstrap samples and the Glivenko-Cantelli theorem, the empirical distribution of the number of rejections in the bootstrap samples converges almost surely in $B$ to its true distribution under the strong null:
+\normalfont
+By the i.i.d.
+construction of the resampled data (Assumption \ref{cbt_def}) and the Glivenko-Cantelli theorem, the empirical distribution of the number of rejections in the resamples converges almost surely in $B$ to its true distribution:
 $$\widehat{F}_{\widehat{\theta}^{(j)}} \xrightarrow[B \to \infty]{A.S.} F_{\widehat{\theta}^{(j)}} $$
@@
 \label{lemma3}
-The number of rejections in the bootstrap samples converges in distribution in $N$ to the number of rejections in samples from the original population under the strong null: $$F_{{ \widehat{\theta} }^{(j)}} \xrightarrow[N \to \infty]{P} F_{{ \widehat{\theta} }^0}$$
+The number of rejections in the resamples converges in distribution in $N$ to the number of rejections in samples from the original population under the strong null:
+$$\lim_{B \to \infty} F_{{ \widehat{\theta} }^{(j)}} = F_{{ \widehat{\theta}^0 }}$$
 
@@
 \begin{align*}
-P \left( \widehat{\theta}^{(j)} = r \right) &= P \left( \sum_{w=1}^W 1 \Big\{ T_w^{(j)} > c_{w,\alpha} \Big\} = r \right) \\
-&= \sum_{ \left( A_1, \dotsb, A_W \right) \in \mathcal{A} } P \left( T_1^{(j)} \in A_1, \dotsb, T_W^{(j)} \in A_W \right) \\
-&\xrightarrow[N \to \infty]{P} \sum_{ \left( A_1, \dotsb, A_W \right) \in \mathcal{A} } P \left( T^0_1 \in A_1, \dotsb, T^0_W \in A_W \right) \tag{\text{from Remark} \ref{CMT_results}}\\
+\lim_{B \to \infty} P \left( \widehat{\theta}^{(j)} = r \right) &= \lim_{B \to \infty} P \left( \sum_{w=1}^W 1 \Big\{ T_w^{(j)} > c_{w,\alpha} \Big\} = r \right) \\
+&= \lim_{B \to \infty} \sum_{ \left( A_1, \dotsb, A_W \right) \in \mathcal{A} } P \left( T_1^{(j)} \in A_1, \dotsb, T_W^{(j)} \in A_W \right) \\
+& \sum_{ \left( A_1, \dotsb, A_W \right) \in \mathcal{A} } P \left( T^0_1 \in A_1, \dotsb, T^0_W \in A_W \right) \tag{\text{from Assumption} \ref{cbt_def}}\\
 &= P \left( \widehat{\theta}^0 = r \right)
@@
 \label{theorem1}
-The empirical distribution of the number of rejections in the bootstrap samples is consistent in $N$ and $B$ for the true distribution of the number of rejections under the strong null:
+The empirical distribution of the number of rejections in the resamples is consistent in $N$ and $B$ for the true distribution of the number of rejections under the strong null:
 $$\widehat{F}_{\widehat{\theta}^{(j)}} \xrightarrow[B, N \to \infty]{P} F_{\widehat{\theta}^0} $$
 \begin{proof}
-Fix an arbitrarily small $\epsilon > 0$ and $\epsilon' > 0$. Remark \ref{glivenko} implies consistency, so there exists a $B^{*}$ such that, for all $B > B^{*}$:
+Fix an arbitrarily small $\epsilon > 0$ and $\epsilon' > 0$. Remark \ref{glivenko} implies convergence in probability, so there exists a $B^{*}$ such that, for all $B > B^{*}$:
 \begin{align*}
-P \left( \vert \widehat{F}_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^{(j)} } \vert > \epsilon / 2 \right) &< \epsilon'/2
+P \left( \vert \widehat{F}_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^{(j)} } \vert > \epsilon / 2 \right) &< \epsilon'
 \end{align*}
@@
 \begin{align*}
-P \left( \vert F_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^0 } \vert > \epsilon / 2 \right) &< \epsilon' / 2
+\left\vert F_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^0 } \right\vert &< \epsilon / 2
 \end{align*}
@@
 &\le P \left( \vert \widehat{F}_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^{(j)} } \vert > \epsilon/2 \right) + P \left( \vert F_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^0 } \vert > \epsilon/2 \right) \\
-&< \epsilon' 
+&< \epsilon' + 0
 \end{align*}
@@
 \end{theorem}
+
+<!-------------------------- --------------------------------->
+<!-------------- SECTION: GIBBS --------------->
+<!-------------------------- --------------------------------->
+
+\section{A general Gibbs resampling algorithm}
+
+For simple modeling choices and study designs, it is fairly easy to specify a resampling algorithm that satisfies Assumption \ref{cbt_def} (for example, @westfalltext provide numerous such algorithms).
+However, designing valid ad hoc algorithms can be challenging for more complex settings, such as generalized linear models with non-identity links [@westfalltext].
+Thus, here, we outline a general resampling algorithm that can capture complex dependence structures for \hl{arbitrary} analysis models, as long as the user is willing to assume a parametric structure for the relationships between outcomes\footnote{This method also assumes correct specification of the relationships between the outcomes and the covariates, but these assumptions are already implicit in the analysis model, so do not represent additional assumptions.}.
+
+Let $\mathbf{Y} = \left( \mathbf{Y}_1, \cdots, \mathbf{Y}_p \right)$ be a random vector of $p$ outcomes and $\mathbf{C}$ be a random vector of auxiliary covariates whose relationships with $\mathbf{Y}$ are not restricted under the strong null.
+Let $\mathbf{Y}^{\text{obs}}$ and $\mathbf{C}^{\text{obs}}$ denote their realized counterparts in the observed dataset.
+Let $\boldsymbol{\alpha}$ be a vector of all parameters characterizing the joint distribution of $\left( \mathbf{Y}, \mathbf{C} \right)$.
+(For example, if the joint distribution is conveniently specified in terms of a series of conditional densities, then $\boldsymbol{\alpha}$ could comprise regression parameters of each outcome in $\mathbf{Y}$ on each covariate in $\mathbf{C}$ and on the other outcomes.) Using $f^0$ to denote PDFs under the strong null, assume a uniform prior on $\boldsymbol{\alpha}$, such that $f^0 \left( \boldsymbol{\alpha} \right) \propto 1$.
+Then, fixing $\mathbf{C}$ as in Example \ref{ex:ols}, the posterior predictive distribution for a new vector of outcomes is:
+
+\begin{align*}
+f^0 \left( \mathbf{Y}^{(j)} \; \vert \; \mathbf{Y}^{\text{obs}}, \mathbf{C}^{\text{obs}} \right) &= \int_{A} f^0 \left( \mathbf{Y}^{(j)} \; \vert \; \mathbf{Y}^{\text{obs}}, \mathbf{C}^{\text{obs}}, \boldsymbol{\alpha} \right) f^0 \left( \boldsymbol{\alpha} \; \vert \; \mathbf{Y}^{\text{obs}}, \mathbf{C}^{\text{obs}} \right) \partial \boldsymbol{\alpha} \\
+&\propto \int_{A} f^0 \left( \mathbf{Y}^{(j)} \; \vert \; \mathbf{Y}^{\text{obs}}, \mathbf{C}^{\text{obs}}, \boldsymbol{\alpha} \right) f^0 \left( \mathbf{Y}^{\text{obs}}, \mathbf{C}^{\text{obs}} \; \vert \; \boldsymbol{\alpha} \right) \partial \boldsymbol{\alpha} \tag{uniform prior}
+\end{align*}
+
+To draw $\mathbf{Y}^{(j)}$, a generic Gibbs sampler could proceed as follows [@gelman]:
+\begin{enumerate}
+\item{Specify fully conditional distributions for each outcome $Y_i$:
+\begin{align*}
+Y_1 &\sim f^0 \left( Y_1 \; \vert \; \boldsymbol{\alpha}_1, \mathbf{C}^{\text{obs}}, \mathbf{Y}_{-1} \right) \\
+&\;\;\vdots \notag \\
+Y_p &\sim f^0 \left( Y_p \; \vert \; \boldsymbol{\alpha}_p, \mathbf{C}^{\text{obs}}, \mathbf{Y}_{-p} \right) \\
+\end{align*}
+where $\mathbf{Y}_{-i}$ represents all outcomes in $\mathbf{Y}$ except for $Y_i$ itself, and where $\boldsymbol{\alpha}_i$ represents the subset of parameters in $\boldsymbol{\alpha}$ that characterize the joint distribution of $\left( Y_i, \mathbf{C} \right)$.
+}
+
+\item{Obtain initial values of $\boldsymbol{\alpha}$ (i.e., $\boldsymbol{\alpha}^{(1)}$) for the first Gibbs iterate, for example by using MLEs from the observed data.}
+
+\item{For iterate $j$, successively draw parameters and predictive values for each outcome:
+\begin{align*}
+\boldsymbol{\alpha}_1^{(j)} &\sim f^0 \left( \boldsymbol{\alpha}_1 \; \vert \; \mathbf{Y}^{(j-1)}, \mathbf{C}^{\text{obs}} \right) \\
+Y_1^{(j)} &\sim f^0 \left( Y_1 \; \vert \; \boldsymbol{\alpha}_1^{(j)}, \mathbf{Y}_{-1}^{(j-1)}, \mathbf{C}^{\text{obs}} \right) \\
+&\;\;\vdots \notag \\
+\boldsymbol{\alpha}_p^{(j)} &\sim f^0 \left( \boldsymbol{\alpha}_p \; \vert \; \mathbf{Y}^{(j-1)}, \mathbf{C}^{\text{obs}} \right) \\
+Y_p^{(j)} &\sim f^0 \left( Y_p \; \vert \; \boldsymbol{\alpha}_p^{(j)}, \mathbf{Y}_{-p}^{(j-1)}, \mathbf{C}^{\text{obs}} \right)
+\end{align*}
+}
+\end{enumerate}
+
+By the \hl{usual} results (\hl{LAME}) regarding Gibbs sampling and Bayesian estimation of the posterior predictive distribution [@gelman], this algorithm satisfies the convergence criterion in Remark \ref{CMT_results}.
+
+In practice, this algorithm can be easily implemented by repurposing existing software for missing data imputation by chained equations (e.g., @micepackage).
+To do so, the observed dataset is copied while setting to missing all values of $\mathbf{Y}$, and this new dataset with "missing" values is concatenated with the observed one, yielding a pseudo-dataset with $2N$ observations (half of which have missing outcomes).
+Then, existing software can be used to create $B$ imputations of the "missing" data, with the imputed portion of each of $B$ imputations (the last $N$ observations of the pseudo-datasets) becoming the resampled datasets.
+Critically, the user must be careful to specify the "imputation" models such that the strong null is enforced; in contrast, typical software defaults use all variables in the dataset to impute all other variables, which would not enforce the strong null.
 
@@
 
-See simulations from 2017-8-1 (written up in "multiple_outcomes (4).pdf").
-Maybe use these these scenarios again?
-Generate at least some data under null to show that power of joint test is <5\%, as intended.
+\subsection{Methods}
 
-\subsection{Width of 95\% CI generated under null}
+We generated multivariate standard normal data, comprising 1 covariate ($X$) and 40 outcomes ($Y_1, \dotsb, Y_{40}$).
+The sample size was fixed at $N = 1000$.
+The correlation between each pair of outcomes was $\rho_{YY}$.
+The correlation between $X$ and a fixed proportion, $q$, of outcomes was $\rho_{XY}$ (with remaining pairs having correlation 0).
+In a full-factorial design, we manipulated:
 
 \begin{itemize}
-\item Figure 1: Mean number of rejections and average of CI limits to show width
-\item Figure 2: Average distance of observed rejections from upper CI limit
+\item $\rho_{XY} \in \{ 0, 0.02, 0.03, 0.04, 0.05, 0.10 \}$
+\item $\rho_{YY} \in \{ 0, 0.25, 0.50 \}$
+\item $q \in \{ 0.50, 1 \}$
+\item $\alpha \in \{ 0.01, 0.05 \}$
 \end{itemize}
 
-\subsection{Power of joint test}
+For each of $1000$ simulation reps per scenario, the simulation algorithm proceeded as follows:
 
-\begin{itemize}
-\item Figure 3 with panels: Prob of rejecting joint null in samples generated under null (should be <5\% for all) and under various alternative scenarios
-\end{itemize}
+\begin{enumerate}
+\item Generate an observed dataset according to the scenario.
+\item Compute $\widehat{\theta}$: For each outcome $Y_i$, regress $Y_i$ on $X$ and conduct a $t$-test at level $\alpha$ on the coefficient for $X$. 
+\item{For each resampling iterate $j$ (with $B = 2000$), apply the algorithm in Example \ref{ex:ols}:
+	\begin{enumerate}
+	  \item Fix $X$ for all observations in the dataset.
+    \item Draw observation index $n'$ randomly with replacement from $\{ 1, \dotsb, N \}$.
+For observation $n$, set $Y_{nw}^{(j)} = \widehat{Y}_{n'w} - Y_{n'w}$ (the regression residual for observation $n'$ in the original dataset).
+    \item Compute $\widehat{\theta}^{(j)}$: Regress $Y_{w}^{(j)}$ on $X$ and conduct a $t$-test at level $\alpha$ on the coefficient for $X$. 
+    \end{enumerate}
+}
+\item Use the quantiles of $\left( \widehat{\theta}^{(1)}, \dotsb, \widehat{\theta}^{(B)} \right)$ to construct a 95\% CI for $\theta_0$ and to check if $\widehat{\theta}$ is above the $1-\alpha$ quantile.
+\end{enumerate}
+
+\subsection{Power to test the strong null}
+
+Figure 2 shows the power of joint tests of the strong null constructed using existing familywise error rate (FWER) procedures, for which we defined rejection of the strong null as rejecting at least one of the 40 FWER-corrected hypothesis tests.
+We compared two non-resampling methods (Bonferroni and Holm), four resampling methods that use data resampled under the strong null (Westfall minP[@westfalltext], Westfall step-down[@westfalltext], our method with $\alpha=0.05$, and our method with $\alpha=0.01$), one resampling method using data resampled without constraints (Romano[@romano]), and our highly experimental procedure involving rejection based on the mean log $p$-value.
+
+For resampling methods using data generated under the null, we used the algorithm in Example \ref{ex:ols}.
+For the single method using data generated without constraints (Romano), we used a classical nonparametric bootstrap in which we simply resampled entire rows of the dataset with replacement and conducted hypothesis tests of the bootstrapped estimates versus those in the original dataset (rather than 0) to recover the null distribution [@hall;@chernick].
+However, note that the classical bootstrap does not guarantee correct inference in the fitted OLS models because the exact repetition of entire rows of data induces correlation between $X$ and the residuals; we chose this approach because we are unaware of an unconstrained resampling algorithm for OLS that preserves parametric assumptions.
+(In contrast, residual resampling as in Example \ref{ex:ols} circumvents this problem because entire rows are not repeated.) Nevertheless, the joint test based on Romano's method under the classical bootstrap showed nominal rejection rates in the null scenarios (Figure 2, upper left panel).
+
+![Power of joint tests based on various FWER control procedures versus our methods.](joint_test.png)
+
+\subsection{Observed rejections vs. 95\% confidence interval under strong null}
+
+Figure 3 shows these results. (As we discussed, for the next iteration of this writeup, I'll remove the scenarios with small effect sizes; they are just included here for completeness.)
+
+![Average number of rejections and confidence limits in datasets resampled under the null when the original data were generated under various null and alternative scenarios.](null_ci.png)
 
@@
 
-<!--
-\newpage
 \section*{References}
--->

### 2018-05-10 R code/For Sherlock/analysis.R

@@ -16,15 +16,11 @@ nrow(s) / (500 * nrow(scen.params))
 table( s$scen )
 
 
-########################### COMPARATIVE POWER PLOT ###########################
+
+########################### DATA PREP ###########################
 
 # need 1 row per scenario-method combination
 # methods are ours, Wstep, bonf.naive, and minP
-
-# variables to keep:
-# scenario parameters
-#  mean of jt.rej.bonf.naive, jt.rej.minP, jt.rej.Wstep
-#  mean of rej.jt.0.01, rej.jt.0.05
 
 library(reshape2)
 library(tidyverse)
@@ -33,8 +29,15 @@ library(tidyverse)
 ( jt.rej.names = c( names(s)[ grep( "jt.rej.", names(s) ) ],
                     names(s)[ grep( "rej.jt", names(s) ) ] ) )
   
-# MAKE SURE TO LIST THEM IN SAME ORDER AS JT.REJ.NAMES
-method.names = c("bonf.naive", "minP", "Wstep", "ours.0.01", "ours.0.05")
+# MAKE SURE TO LIST THEM IN SAME ORDER  AS JT.REJ.NAMES
+method.names = c("bonf.naive",
+                 "holm",
+                 "minP",
+                 "Wstep",
+                 "Romano",
+                 "meanP",
+                 "ours.0.01",
+                 "ours.0.05")
 
 # reshape wide to long
 # https://stackoverflow.com/questions/12466493/reshaping-multiple-sets-of-measurement-columns-wide-format-into-single-columns
@@ -49,6 +52,7 @@ lp = reshape( s[ , names(s) %in% c(jt.rej.names, "scen") ],
 # long data should have 1 row per method-scenario combination
 # nrow( lp[ lp$scen == "a", ] ); nrow( s[ s$scen == "a", ]) * length(method.names)
 
+
 # since each scenario still has multiple rows, take means 
 pwr = lp %>% group_by(scen, method) %>%
   summarise( power = mean(jt.rej) )
@@ -58,20 +62,48 @@ pwr = merge(pwr, scen.params, by = "scen" )
 # for plotting joy
 pwr$group = paste( "X-Y correlation: ", pwr$rho.XY,
                    " for ",
-                   ifelse( pwr$half, "half of", "all" ),
-                   " pairs", 
+                   pwr$prop.corr * 100,
+                   "% of pairs", 
[…]

### 2018-05-10 R code/For Sherlock/doParallel.R

@@ -20,6 +20,22 @@ alpha = c( crit.bonf, 0.01, 0.05 )
 # simulation reps to run within this job
 # this need to match n.reps.in.doParallel in the genSbatch script
 sim.reps = 5
+
+# EDITED FOR C++ ISSUE WITH PACKAGE INSTALLATION
+library(doParallel, lib.loc = "/share/PI/manishad/Rpackages/")
+library(foreach, lib.loc = "/share/PI/manishad/Rpackages/")
+library(mvtnorm, lib.loc = "/share/PI/manishad/Rpackages/")
+library(StepwiseTest, lib.loc = "/share/PI/manishad/Rpackages/")  # Romano
+library(matrixcalc, lib.loc = "/share/PI/manishad/Rpackages/")
+
+source("functions.R")
+
+# names of methods that resample under H0 vs. HA
+h0.methods = c("Y", "resid", "h0.parametric")
+ha.methods = c("fcr", "ha.resid", "ha.resid.2")
+
+# set the number of cores
+registerDoParallel(cores=16)
 ######### END OF CLUSTER PART #########
 
 
@@ -33,20 +49,20 @@ sim.reps = 5
 # rho.XX = 0
 # rho.YY = c(0.25)
 # rho.XY = c(0.08)  # null hypothesis: 0
-# half = 0
+# prop.corr = 0.05
 # 
 # # bootstrap iterates and type
 # boot.reps = 5
 # sim.reps = 2
 # scen = "a"
-# bt.type = c( "resid" )
+# bt.type = c( "ha.resid" )
 # 
 # 
 # # matrix of scenario parameters
 # scen.params = expand.grid( bt.type, n, nX, nY, rho.XX,
-#                            rho.YY, rho.XY, half )
+#                            rho.YY, rho.XY, prop.corr )
 # names(scen.params) = c( "bt.type", "n", "nX", "nY", "rho.XX",
-#                         "rho.YY", "rho.XY", "half" )
+#                         "rho.YY", "rho.XY", "prop.corr" )
 # 
 # # name the scenarios
 # # remove letters that are privileged variables in R
@@ -59,8 +75,22 @@ sim.reps = 5
 # crit.bonf = 0.05 / p$nY
 # alpha = c( crit.bonf, 0.01, 0.05 )
 # 
-# # for sourcing functions later
+# library(doParallel)
+# library(foreach)
+# library(mvtnorm)
[…]

### 2018-05-10 R code/For Sherlock/doParallel_mice.R

@@ -19,7 +19,7 @@ alpha = c( crit.bonf, 0.01, 0.05 )
 
 # simulation reps to run within this job
 # this need to match n.reps.in.doParallel in the genSbatch script
-sim.reps = 5
+sim.reps = 1
 ######### END OF CLUSTER PART #########
 
 
@@ -73,10 +73,12 @@ sim.reps = 5
 # runs all bootstrapping iterations
 # writes 1-row csv with number rejected and bootstrap CI
 
+# need to load all packages here, not just in functions.R?
 library(doParallel)
 library(foreach)
 library(mvtnorm)
 library(BinNor)
+library(mice, lib.loc = "/share/PI/manishad/Rpackages/")
 
 # ~~~ LATER COULD PUT MICE FUNCTIONS IN REGULAR FNS THING
 #  BECAUSE THEY SHOULD BE BACKWARD-COMPATIBLE
@@ -134,11 +136,14 @@ for ( j in 1:sim.reps ) {
   
   ######## Begin MICE sampling under H0 ########
   if ( p$bt.type == "MICE.H0" ) {
+    
+    cat("Flag 1: entered imputation loop")
+    
     # double the dataset
     d2 = rbind(d, d)
     
     # make second half of dataset missing
-    d2[ (n+1) : nrow(d2), names(d2) %in% Y.names ] = NA
+    d2[ (p$n+1) : nrow(d2), names(d2) %in% Y.names ] = NA
     
     # recode binaries as factors to avoid whiny MICE
     for (y in Y.names) {
@@ -146,7 +151,6 @@ for ( j in 1:sim.reps ) {
     }
     
     # first fit empty MICE to get predictor matrix
-    library(mice)
     ini = mice(d2, maxit = 0)
     pred = ini$predictorMatrix
     
@@ -158,6 +162,8 @@ for ( j in 1:sim.reps ) {
     method = ini$method
     method[ method != "logreg" & method != "" ] = "logreg"
     imp = mice( d2, pred = pred, method = method, m = boot.reps )
+    
+    cat("Flag 2: finished imputation")
   }
   ######## End MICE sampling under H0 ########
   
@@ -168,7 +174,7 @@ for ( j in 1:sim.reps ) {
       
       # MICE has already generated all the datasets
       # so here, just extract the single imputed dataset that we need
-      b = complete(imp,i)[ (n+1) : (2*n), ]
+      b = complete(imp,i)[ (p$n+1) : (2*p$n), ]
 
[…]

### 2018-05-10 R code/For Sherlock/functions.R

@@ -1,6 +1,5 @@
 
 
-
 ######################## FNS FOR WESTFALL's SINGLE-STEP ########################
 
 # returns minP-adjusted p-values (single-step)
@@ -133,9 +132,18 @@ get_crit = function( p.dat, col.p ) {
 # .half: Should only the last half the outcomes have the specified correlation?
 #  (if FALSE, correlation matrix is exchangeable; otherwise first half are 0 and 
 #  second half are rho.XY)
+# This is only included for backward compatibility because now we have the more flexible .prop.corr thing
 
-cell_corr = function( vname.1, vname.2, .rho.XX, .rho.YY, .rho.XY,
-                      .nY, .half = FALSE ) {
+# .prop.corr: What proportion of Ys should be correlated with X?
+
+cell_corr = function( vname.1,
+                      vname.2,
+                      .rho.XX,
+                      .rho.YY,
+                      .rho.XY,
+                      .nY,
+                      #.half = 0,
+                      .prop.corr = 1) {
   
   # use grep to figure out if variables are covariates or outcomes
   if ( length( grep("X", vname.1 ) ) == 1 ) {
@@ -168,20 +176,41 @@ cell_corr = function( vname.1, vname.2, .rho.XX, .rho.YY, .rho.XY,
     # equal correlation between X1 and all outcomes
     # all other covariates have correlation 0 with outcome
     if( "X1" %in% c( vname.1, vname.2 ) ) {
-      # exchangeable case
-      if ( !.half ) return( .rho.XY )
       
-      # "half are correlated" case
-      if (.half) {
+      # ##### ONLY FOR BACKWARD COMPATIBILITY #####
+      # # exchangeable case
+      # if ( .half == 0 ) return( .rho.XY )
+      # 
+      # # "half are correlated" case
+      # if ( .half == 1 ) {
+      #   # find the one that is the outcome
+      #   outcome.name = ifelse( vtype.1 == "outcome", vname.1, vname.2 )
+      #   
+      #   # extract its number (4 for "Y4")
+      #   num = substring( outcome.name, first = 2 )
+      #   
+      #   # see if its number is greater than halfway or not
+      #   return( ifelse( num > .nY / 2, .rho.XY, 0 ) )
+      # }
+      # ##### END PART FOR BACKWARD COMPATIBILITY #####
+      
+      # ~~~ MODIFIED
[…]

### 2018-05-10 R code/For Sherlock/gen_sbatch.R

@@ -15,29 +15,45 @@ nX = 1
 nY = 40
 rho.XX = 0
 rho.YY = c(0.25, 0.5, 0)
-rho.XY = c(0.02, 0.05, 0) # null hypothesis: 0
-half = c(0, 1)  # exchangeable vs. half-correlated matrix
+rho.XY = c(0, 0.02, 0.03, 0.04, 0.05, 0.10) # null hypothesis: 0
+prop.corr = c(0.5, 1)  # exchangeable vs. half-correlated matrix
 
 # bootstrap iterates and type
 boot.reps = 2000
-bt.type = c( "resid" )
+bt.type = c( "resid", "ha.resid" )
+
 
 # matrix of scenario parameters
-scen.params = expand.grid( bt.type, n, nX, nY, rho.XX, rho.YY, rho.XY, half )
-names(scen.params) = c( "bt.type", "n", "nX", "nY", "rho.XX", "rho.YY", "rho.XY", "half" )
+scen.params = expand.grid( bt.type, n, nX, nY, rho.XX, rho.YY, rho.XY, prop.corr )
+
+# give letters to scenarios
+names(scen.params) = c( "bt.type", "n", "nX", "nY", "rho.XX", "rho.YY", "rho.XY", "prop.corr" )
 
 
-# name the scenarios
+######## Remove Some Scenarios ########
+# for rhoXY = 0.10, only want info for confidence intervals
+# so cut out the ha.resid ones
+scen.params = scen.params[ scen.params$rho.XY != 0.10 |
+                             scen.params$rho.XY == 0.10 & scen.params$bt.type == "resid", ]
+
+# for rho.XY = 0, don't need to manipulate prop.corr because all 0 anyway
+scen.params = scen.params[ scen.params$rho.XY != 0 |
+                             scen.params$rho.XY == 0 & scen.params$prop.corr == 1, ]
+
+
+######## Name the Scenarios ########
+# if merging results with other simulations, set this to the last letter already used
+start.at = "a"
+
 # remove letters that are privileged variables in R
-letter.names = c(letters, LETTERS)[ ! c(letters, LETTERS) %in% c("i","T","F") ]
-scen.params$scen.name = letter.names[ 1:dim(scen.params)[1] ]
+# letter.names = c(letters, LETTERS, paste(letters, letters, sep="") )
+# letter.names = letter.names[ ! c(letters, LETTERS) %in% c("i","T","F") ]  # exclude privileged characters
[…]

### 2018-05-10 R code/For Sherlock/gen_sbatch_mice.R

@@ -14,12 +14,17 @@ n = 1000
 nX = 1
 nY = 40
 rho.XX = 0
-rho.YY = c(0.25, 0.5, 0)
-rho.XY = c(0.02, 0.05, 0) # null hypothesis: 0
-half = c(0, 1)  # exchangeable vs. half-correlated matrix
+rho.YY = c(0.25, 0)
+rho.XY = c(0.05, 0) # null hypothesis: 0
+half = c(0)  # exchangeable vs. half-correlated matrix
+
+# scenarios from OLS
+# rho.YY = c(0.25, 0.5, 0)
+# rho.XY = c(0.02, 0.05, 0) # null hypothesis: 0
+# half = c(0, 1)  # exchangeable vs. half-correlated matrix
 
 # bootstrap iterates and type
-boot.reps = 2000
+boot.reps = 1000
 bt.type = c( "MICE.H0" )
 
 # matrix of scenario parameters
@@ -51,7 +56,7 @@ source("functions.R")
 
 # number of sbatches to generate (i.e., iterations within each scenario)
 n.reps.per.scen = 500
-n.reps.in.doParallel = 5
+n.reps.in.doParallel = 1
 n.files = ( n.reps.per.scen / n.reps.in.doParallel ) * n.scen
 
 
@@ -68,7 +73,7 @@ runfile_path = paste(path, "/testRunFile.R", sep="")
 sbatch_params <- data.frame(jobname,
                             outfile,
                             errorfile,
-                            jobtime = "0:30:00",
+                            jobtime = "6:00:00",
                             quality = "normal",
                             node_number = 1,
                             mem_per_node = 64000,

### 2018-05-10 R code/For Sherlock/stitch_on_sherlock.R

@@ -16,7 +16,7 @@ s = stitch_files(.results.singles.path = "/share/PI/manishad/multTest/sim_result
 write.csv( s, "stitched.csv" )
 
 # how close are we to being done?
-(nrow(s)/5) / 1500
+(nrow(s)/5) / 12000
 table(s$scen)
 
 # move it to Desktop
@@ -24,39 +24,3 @@ scp mmathur@sherlock:/share/PI/manishad/multTest/sim_results/stitched.csv ~/Desk
 scp mmathur@sherlock:/share/PI/manishad/multTest/scen_params.csv ~/Desktop
 
 
-
-
-# #### Sneak-Peek Analyes
-# mean(s$covers.raw)
-# mean(s$covers)
-# mean(s$covers.correct)
-# 
-# summary(s$rate)
-# 
-# # symmetry of CIs
-# mean( abs( s$ci.hi.raw - s$excess ) - abs( s$ci.lo.raw - s$excess ) )
-# mean( abs( s$ci.hi.correct - s$excess ) - abs( s$ci.lo.correct - s$excess ) )
-# 
-# 
-# 
-# 
-# 
-# 
-# # zip the entire results folder
-# setwd("/share/PI/manishad/multTest")
-# system( paste("zip -r sim_results.zip sim_results") )
-# 
-# 
-# ######## FOR STITCHING SHORT FILES ########
-# source("functions.R")
-# s = stitch_files(.results.singles.path = "/share/PI/manishad/multTest/sim_results/short",
-#                  .results.stitched.write.path = "/share/PI/manishad/multTest/sim_results",
-#              .name.prefix = "short",
-#              .stitch.file.name="stitched.csv")
-# 
-# # zip the entire results folder
-# setwd("/share/PI/manishad/multTest")
-# system( paste("zip -r sim_results.zip sim_results") )
-# 
-# 
-# 

### 2018-05-13 Manuscript/manuscript.Rmd
Revising ms

@@
 
+\newcommand*{\vertbar}{\rule[-1ex]{0.5pt}{2.5ex}}
+
 \newpage
@@
 <!-------------------------- --------------------------------->
-<!-------------- SECTION: SETTING AND NOTATION --------------->
+<!------------- SECTION: SETTING AND NOTATION -------------->
 <!-------------------------- --------------------------------->
+
+\section{Setting and notation}
+
+To cover:
+\begin{itemize}
+\item Draw data structure for us and for Freedman
+\item Distance metric
+\item Various distribution functions
+\item Matrices
+\item Assumptions
+\end{itemize}
+
+Assumptions:
+\begin{itemize}
+\item Ones borrowed directly from Freedman
+\item All regressions have same design matrix and sample size. Do they all need to have same exposure of interest?
+\item Residuals have mean 0
+\item Independent observation
+\end{itemize}
+
+Superscripts denote lengths. Subscripts denote indices. $[\cdot]_{ij}$ denotes the $i,j^{th}$ element of a matrix, and $[\cdot]_{i}$ denotes the $i^{th}$ element of a vector.
+
+$\beta^W$: the $W$-vector of coefficients of interest for each regression.
+For clarity, we use the superscript to remind the reader that this is not the usual $p$-vector of coefficients from a single multiple regression model, but rather a vector of single coefficient estimates from $W$ different regression models.
+
+$\widehat{\beta}^W$: the W-vector of coefficient estimates of interest for each regression
+
+$\epsilon_w$: the $N$-vector of true errors for the $w^{th}$ regression
+
+$\epsilon_{kw}$: the scalar true error for observation $k$ in regression $w$
+
+$X$: the $N \times p$ design matrix
+
+$\underset{p \times N} A = \left( X'X \right)^{-1}X'$
+
+$\Psi_n(F)$: the distribution of $\sqrt{n} \left( \widehat{\beta}^W - \beta^W \right)$
+
+$p$: number of parameters. See Lemma 2.1'
+
+$\sigma^2_w$: see Lemma 2.1'
+
+<!-------------------------- LEMMA 8.9' ---------------------------->
+The following development closely follows @freedman, who considered the asymptotic validity of residual resampling in recovering the sampling distribtribution for a single multiple linear regression model.
+Here, we extend this work to consider the sampling distribution of coefficient estimations from several multiple linear regression models sharing the same design matrix, but with different outcomes.
+Several of the following lemmas parallel those in @bickel.
+\begin{lemma}[Lemma 8.9']
+Let $C$ and $D$ be $W \times N$ random matrices, and let $B$ be a fixed $N$-vector. Let $G_C$ and $G_D$ be the marginal distribution functions of the $W$-vectors $CB$ and $DB$, respectively. Then: 
+\begin{align*}
+d_2^W \left( G_C, G_D \right)^2 &\le \text{tr} \{ BB' E \big[ (C-D)'(C-D) \big] \}
+\end{align*}
+
+\begin{proof}
+\begin{align*}
+d_2^W \left( G_C, G_D \right)^2 &\le E \Big[ || CB - DB ||^2 \Big]\\
+&= E \Big[ \text{tr} \{ \underbrace{(CB-DB)}_{1 \times W} \underbrace{(CB-DB)'}_{W \times 1} \} \Big] \\
+&= E \Big[ \text{tr} \{ \underbrace{(C-D)}_{W \times N} \underbrace{BB'}_{N \times N} \underbrace{(C-D)'}_{N \times W} \} \Big] \\
+&= E \Big[ \text{tr} \{ BB' (C-D)'(C-D) \} \Big] \tag{trace cyclical invariance}\\
+&= \text{tr} \{ E \Big[ BB' (C-D)'(C-D) \Big] \}  \\
+&= \text{tr} \{ BB' E [ \underbrace{ (C-D)'(C-D) }_{N \times N} ] \} \tag{B is fixed} \\
+\end{align*}
+\textcolor{red}{In the inequality, the LHS is the infimum of the expectation over all possible pairs of random variables, whereas the RHS is the expectation for a particular pair of random variables (see similar step in Bickel \& Freedman pg 1214, Lemma 8.9.}
+\end{proof}
+\end{lemma}
+
+<!-------------------------- THEOREM 2.1'---------------------------->
+The next theorem bounds the distance between the true sampling distribution of the test statistics and the estimated sampling distribution in the resamples in terms of the distance between the sampling distribution of the true errors and the resampled residuals.
+
+\begin{theorem}[Theorem 2.1']
+\label{bound_thm}
+Let $\underset{N \times 1} B$ be the $i^{th}$ column of $(X'X)^{-1}X'$ (i.e., corresponding to the $i^{th}$ regression covariate). Then:
+\begin{align*}
+d_2^W \left( \Psi(F), \Psi(F_N) \right)^2 &\le N \cdot \text{tr} \{ BB' \} \cdot d_2^W \left( F, F_N \right)^2
+\end{align*}
+
+\begin{proof}
+
+In general for multiple regression, we have $\widehat{\beta} - \beta = \left( X'X\right)^{-1} X' \epsilon$.
+Letting:
+
+\begin{align*}
+C = \underbrace{
+\begin{bmatrix}
+    \text{---} \hspace{-0.2cm} & \epsilon_1' & \hspace{-0.2cm} \text{---} \\
+    & \vdots & \\
+    \text{---} \hspace{-0.2cm} & \epsilon_W' & \hspace{-0.2cm} \text{---}
+\end{bmatrix} }_{W \times N}
+\end{align*}
+
+we can express $\Psi_N(F)$ as the law of the $W$-vector:
+
+\begin{align*}
+\sqrt{N} \left( \widehat{\beta}^W - \beta^W \right) &= \sqrt{N} \begin{bmatrix}
+    [ (X'X)^{-1}X' \epsilon_1 ]_{i}\\
+    \vdots\\
+    [ (X'X)^{-1}X' \epsilon_W ]_{i}
+\end{bmatrix} = \sqrt{N} \begin{bmatrix}
+    \epsilon_1' B \\
+    \vdots\\
+    \epsilon_W' B
+\end{bmatrix} =
+\sqrt{N} \cdot CB
+\end{align*}
+
+whose $w^{th}$ element pertains to the $i^{th}$ regression coefficient in the $w^{th}$ regression.
+Now let $D$ be the counterpart of $C$ with the residual vectors, $\widehat{\epsilon}_w$, in place of the true errors, $\epsilon_w$.
+Then the entries of matrix $(C-D)'(C-D)$ are:
+\begin{align*}
+[(C-D)'(C-D)]_{kj} &= \sum_{w=1}^W \left( \epsilon_{jw} - \widehat{\epsilon}_{jw} \right) \left( \epsilon_{kw} - \widehat{\epsilon}_{kw} \right)
+\end{align*}
+
+We have $E[ \left( \epsilon_{jw} - \widehat{\epsilon}_{jw} \right) \left( \epsilon_{kw} - \widehat{\epsilon}_{kw} \right) ] = \text{Cov} \left( \epsilon_{jw} - \widehat{\epsilon}_{jw}, \epsilon_{kw} - \widehat{\epsilon}_{kw} \right)$, but for all $k \ne j$, the covariance is 0 because the observations are independent.
+Thus, $E[(C-D)'(C-D)]$ is a diagonal matrix equal to $I^{N} \cdot E \Big[ \sum_{w=1}^W \left( \epsilon_{jw} - \widehat{\epsilon}_{jw} \right)^2 \Big] = I^{N} \cdot d_2^W \left( \widehat{F}_n, F\right)$.
+\textcolor{red}{I took the latter equality from B\&F Lemma 8.9 but don't understand it.
+I traced it back to the proof of Lemma 8.6 (pg 1213), in which this equality is stated to hold "without loss of generality in view of Lemma 8.1", but I still don't understand.} The result then follows immediately from applying Lemma 8.9', setting $G_C = F$, $G_D = \widehat{F}_n$ and $B, C,$ and $D$ as defined above and pulling the scalar $\sqrt{N}$ outside the squared distance.
+\end{proof}
+\end{theorem}
+
+To apply the bound in Theorem 2.1', we will first bound the term on the right-hand side through a triangle inequality argument, which applies because $d_2^W(\cdot, cdot)$ is a metric [@bickel]:
+
+\begin{align}
+\label{triangle}
+d_2^W \left( \widehat{F}_N, F \right)^2 \le d_2^W \left( \widehat{F}_N, F_N \right)^2 + d_2^W \left( F_N, F \right)^2 
+\end{align}
+The first term on the RHS relates the empirical distribution of the residuals to the empirical distribution of the true errors (which are discrete distributions taking $N$ values); the second term relates the latter empirical distribution to the true error distribution (which is continuous).
+
+\begin{lemma}[Lemma 2.1']
+\label{first_bound}
+Concerning the first term in Equation (\ref{triangle}):
+\begin{align*}
+d_2^W \left( \widehat{F}_N, F_N\right) \xrightarrow[N \to \infty]{P} 0
+\end{align*}
+
+\begin{proof}
+As in Definition XXX, let $U \sim \widehat{F}_N$ and $V \sim F_N$ be two random variables in $\mathbb{R}^W$ that follow the empirical marginal distributions of the residuals and true errors.
+Denote their elements $\left( U_1, \dotsb, U_W \right)$ and $\left( V_1, \dotsb, V_W \right)$.
+\textcolor{red}{Let $(\widetilde{U}, \widetilde{V})$ be a special choice of $(U,V)$ that follow not only the marginal empirical distributions $\widehat{F}_N$ and $F_N$, but also the empirical \emph{joint} distribution of the residuals and the true errors.}
+
+\begin{align*}
+d_2^W \left( \widehat{F}_N, F_N\right) & := \inf\limits_{ \substack{U \sim \widehat{F}_N \\ V \sim F_N} } E \big[ || U - V ||^2 \big] \\
+& \textcolor{red}{\le E \Big[ || \widetilde{U} - \widetilde{V} ||^2 \Big] \tag{\text{choose one element from inf set}} }\\
+& = \frac{1}{N} \sum_{n=1}^N \underbrace{ \sum_{w=1}^W \left( \widehat{\epsilon}_{nw} - \epsilon_{nw} \right)^2 }_{||\cdot||^2 \text{ of a $W$-vector}} \tag{\text{joint ECDF expectation}}\\
+& = \frac{1}{N} \sum_{w=1}^W \underbrace{ \sum_{n=1}^N \left( \widehat{\epsilon}_{nw} - \epsilon_{nw} \right)^2 }_{||\cdot||^2 \text{ of an $N$-vector}}\\
+&= \frac{1}{N} \sum_{w=1}^W || \widehat{\epsilon}_w - \epsilon_w ||^2 \\
+\Rightarrow E \Big[ d_2^W \left( \widehat{F}_N, F_N\right) \Big] &= \frac{p}{N} \sum_{w=1}^W \sigma^2_w \tag{\text{Freedman Eq. (2.2)}} \\
+& \xrightarrow[N \to \infty]{P} 0
+\end{align*}
+where $p$ is the number of parameters and $\sigma^2_W$ is XXX.
+The interchange of summations in lines 3-4 is used to turn a norm involving residuals from different regressions, summed over $N$ observations, into a norm involving residuals of observations within a regression, summed over $W$ regressions.
+The latter is more convenient because it allows application of existing theory for a single multiple regression model.
+
+\end{proof}
+\end{lemma}
+
+\begin{lemma}
+\label{second_bound}
+Concerning the expectation of the second term in Equation (\ref{triangle}):
+\begin{align*}
+E \Big[ d_2^W \left( \widehat{F}_N, F \right)^2 \Big] \xrightarrow[N \to \infty]{P} 0
+\end{align*}
+
+\begin{proof}
+Letting $P_N$ denote an empirical probability, \textcolor{red}{the joint cumulative distribution corresponding to} $F_N$ is:
+\begin{align*}
+P_N \left( \epsilon_{n1} \le c_1, \dotsb, \epsilon_{nW} \le c_W \right) &= \frac{1}{N} \sum_{n=1}^N 1 \big\{ \epsilon_{n1} \le c_1, \dotsb, \epsilon_{nW} \le c_W \big\} \\
+&\xrightarrow[N \to \infty]{p} P \left( \epsilon_{n1} \le c_1, \dotsb, \epsilon_{nW} \le c_W \right) \tag{SLLN}
+\end{align*}
+which is the joint cumulative distribution corresponding to $F$. By @bickel's Lemma 8.3, this convergence in probability implies the desired result.
+\end{proof}
+\end{lemma}
+
+\begin{theorem}[Lemma 2.4']
+\begin{align*}
+\widehat{\beta}^{*W} \xrightarrow[N \to \infty]{D} \widehat{\beta}^W
+\end{align*}
+
+\begin{proof}
+
+\textcolor{red}{First, note that $\text{tr} \{ BB' \} \ge 0$ because XXX.} Thus, combining Theorem \ref{bound_thm} with the triangle inequality in Equation (\ref{triangle}) yields:
+\begin{align*}
+d_2^W \left( \Psi(F), \Psi(F_n) \right)^2 &\le N \cdot \text{tr} \{ BB' \} \cdot \left( d_2^W \left( \widehat{F}_N, F_N \right)^2 + d_2^W \left( F_N, F \right)^2 \right)
+\end{align*}
+
+The term $N \cdot \text{tr} \{ BB' \} \xrightarrow[N \to \infty]{P} k$ by Assumption XXX.
+The convergence in $r^{th}$ mean of Lemma \ref{first_bound} implies that $d_2^W \left( \widehat{F}_N, F \right) \xrightarrow[N \to \infty]{P} 0$.
+Lastly, by Lemma \ref{second_bound}, $d_2^W \left( F_N, F \right)^2 \xrightarrow[N \to \infty]{P} 0$.
+We therefore have $d_2^W \left( \Psi(F), \Psi(F_n) \right)^2 \xrightarrow[N \to \infty]{P} 0$.
+The desired result then follows from @bickel's Lemma 8.3 relating convergence of the Mallow's metric to convergence in distribution.
+\end{proof}
+\end{theorem}
+
+<!----
+\begin{lemma}[Lemma 2.4']
+\begin{align*}
+\frac{1}{N} \sum_{w=1}^W || \widehat{\epsilon}_w - \epsilon_w ||^2 \xrightarrow[N \to \infty]{p} 0
+\end{align*}
+
+\begin{proof}
+This follows immediately from applying @freedman's Lemma 2.4 to each of the $W$ terms. 
+\end{proof}
+\end{lemma}
+
+BOOKMARK: THIS LEMMA SHOULD BE THAT THE EXPECTATION IS GOING TO ZERO. LAST LINE WITH SIGMA^2 IS WRONG B/C SHOULD BE EXPECTATION. SEE FREEDMAN. 
+---->
+
+<!-------------------------- --------------------------------->
+<!------------- SECTION: MAIN RESULTS -------------->
+<!-------------------------- --------------------------------->
+
+\section{Results}
+
+<!--
+<!-------------------------- --------------------------------->
+<!---------- SECTION: OLD MATERIAL (BEFORE 5/4/2018) ----------->
+<!-------------------------- --------------------------------->
+\newpage
+\section{OLD MATERIAL}
 
@@
 with null hypotheses $H_{01}: \beta_1 = 0$ and $H_{02}: \alpha_1 = 0$, and assuming $\epsilon_{nw} \sim N(0, \sigma^2_w)$ for $w \in \{1,2\}$.
-A resampling algorithm \hl{[@westfalltext]} satisifying Assumption \ref{cbt_def} could proceed by first fixing the covariates $Z_{n3}$ and $Z_{n4}$ for all observations while setting the "outcomes" equal to a vector of residuals resampled with replacement.
+A resampling algorithm \textcolor{red}{[@westfalltext]} satisifying Assumption \ref{cbt_def} could proceed by first fixing the covariates $Z_{n3}$ and $Z_{n4}$ for all observations while setting the "outcomes" equal to a vector of residuals resampled with replacement.
 That is, letting $n'$ denote an observation sampled with replacement and letting hats denote the usual OLS estimates obtained from the original dataset, the resampled variables for observation $n$ are:
@@
 Critically, the user must be careful to specify the "imputation" models such that the strong null is enforced; in contrast, typical software defaults use all variables in the dataset to impute all other variables, which would not enforce the strong null.
+-->
 

### 2018-05-13 Project log/multiple_outcomes_project_log.Rmd
Revising ms

@@
 - \usepackage{float}
+- \usepackage{amsmath}
 bibliography: refs_fake.bib
@@
 
+<!-- NEW SECTION -->
+\section{Proof sketch following Freedman paper}
+
+\begin{enumerate}
+\item{ Define multivariate Mallows metric, $\tilde{d}_2$. Prove that it's still a metric and that Eq. (2.3) still holds. Since it's a metric, triangle inequality (2.6) still holds:
+$$\frac{1}{2} \tilde{d}_2(F_n, F)^2 \le \tilde{d}_2(\widehat{F}_n, F_n) + \tilde{d}_2( F_n, F)$$}
+
+\item{ $\bigstar$ Show that Theorem 2.1 still holds for $\tilde{d}_2$, assuming the same design matrix for all regressions:
+$$\tilde{d}_2^p( \Psi (F_n), \Psi (F)) \le n \cdot \text{trace} [X(n)'X(n)]^{-1} \cdot d_2(F,G)$$
+}
+
+\item{Work on each piece of the triangle inequality:
+	\begin{enumerate}
+    	\item $\tilde{d}_2(F_n, F) \to 0$ because we can still apply SLLN to joint distribution (analog to B\&F Lemmas 8.3/8.4).
+        \item $\tilde{d}_2( \widehat{F}_n, F_n) \le \frac{p}{n} \sum_{w=1}^W \sigma^2_w \to 0$ (analog to Lemmas 2.1/2.2).
+    \end{enumerate}
+}
+
+\item{Lemma 2.4 still holds because each term $\to 0$ by B\&F Lemma 2.4:
+$$\frac{1}{n} \sum_{w=1}^W \| \widehat{\epsilon}_{w} (n) - \widehat{\epsilon}_{w} (n) \|^2 \to 0$$
+}
+
+\item{ Lemma 2.5 still holds using our Eq. (2.3), our Lemma 2.4, and the fact that $\widehat{F}_n = \tilde{F}_n$ for us:
+
+$$d_2( \widehat{F}_n, F_n ) \to 0$$}
+
+\item{$\bigstar$ Show that $\Psi (F_n)$ is close to $\Psi (\widehat{F}_n)$ by writing difference between vector of coefficient estimates in terms of residuals and resampled residuals (loose analog to Theorem 2.2).}
+\end{enumerate}
+
 \section{Stream of consciousness}
 
-Mystery #2: Why does Westfall recommend a resampling scheme for logistic regression that doesn't preserve correlation? (pg 215)
+MICE thing doesn't work because we're regressing $Y_2$ on $Y_1$, and if data are generated under $H_A$, these are correlated because both outcomes are correlated with $X$.
+This remains true even if we leave $X$ out of the imputation model.
 
-Multiple imputation idea: Maybe a general algorithm for resampling under $H_A$ could be something like MI.
-You would take your dataset and copy all the Xs, deleting the Ys.
-Then you would use an off-the-rack parametric MI procedure to impute the "missing data" for each of the resamples (and each time delete the observed data).
-E.g., could do a fully conditional approach where you first regression Y1 on all other variables, including the other outcomes, then regress Y2 on the remaining ones, etc.
-This does invoke assumptions about the form of relationship between the outcomes, but at least wouldn't break any parametric assumptions.
-I think you could do this with MICE's off-the-rack models because it fits logistic regression to binary variables by default (see JStatSoft paper)!
-To resample under $H_0$ this way, could first shuffle the Ys in original dataset?
-That should preserve correlation between Ys.
+FCR-center works for OLS but not logistic regression because for the latter, SE depends on value of parameter. (Is this subset pivotality?)
 
-\section{For Westfall simulations}
+Marginally vs. conditionally: See simulation writeups.
 
-* Redo as-is but with 1000 reps/scenario, and have more reps in doParallel
-* Add Holm to simulations -- maybe do it along with Romano?
-* Add Romano to simulations (requires FCR instead of Westfall approach) since it seems to work in practice (\texttt{debug\_fcr}), but remember that this violates OLS assumptions, so isn't technically advisable.
-
-\section{For later simulations}
-
-* Add binary outcomes with logistic regression?
-* To further improve power, could Use something about p-values themselves for joint test?
-
-\section{FWER control methods we should compare for power of joint test}
-
-\begin{enumerate}
-\item Westfall's single-step minP and maxT methods (pg 47): Easy to code manually
-\item Westfall's step-down free method (pg 66): Moderate to code manually
-\item Romano: Implemented in StepwiseTest package, but needs debugging
-\item Holm (step-down): Easy to code manually; UMP than Bonferroni
-\item Bonferroni (single-step): Easy to code manually
-\end{enumerate}
+Bootstrapping brainteaser: Observations really are independent in FCR resamples. I think the previous issues with this were actually not due to failed assumptions. 
+\url{https://stats.stackexchange.com/questions/339237/are-observations-independent-in-bootstrapped-resamples}
 
@@
 \item Talks a lot about Westfall and also uses subset pivotality
-\item Pg 7: "Subset pivotality is achieved if the distribution of [the p-values of the true null hypotheses] is unaffected by the truth or falsehood of the remaining hypotheses"
+\item Pg 7 (subset pivotality): When we resample under the strong null, the distribution of $p$-values for the null hypotheses that are actually \emph{false} is different from what it is in real life.
+That's because we are resampling such that \emph{all} nulls are true, whereas in real life some might be false.
+The goal is to use the resampled distribution of $p$-values \emph{for the nulls that are true} to approximate its true distribution.
+This only holds if the distribution of that subset of $p$-values isn't affected by the truth or falsehood of the remaining hypotheses.
 \item Seems to be very closely related to Westfall, but cares about FDR instead of FWER. 
@@
 \item Why resample not under the null?
-Since we don't know which hypotheses are true, assume the data really were generated under the null because this is conservative.
-If the data aren't generated under the null, then the critical values will just be more stringent because the kth-largest test stat will be larger (page 1383).
-Basically this "exploits the duality between CIs and hypothesis tests" (pg 1388).
 Apparently resampling not under the null avoids subset pivotality (pg 1388).
+Resampling under $H_A$ "exploits the duality between CIs and hypothesis tests" (pg 1388).
+Even though we're resampling under $H_A$, the \emph{critical value} is still under the null distribution.
+(See Example 2.1, where the critical value is clearly for the distribution when $\mu=0$.
+Also note that the $A$ subscript on the critical values, for example on page 1384, indicates that it's the critical value when all $s$ nulls are true.) Thus, I think he must be centering the bootstrapped test stats by the observed one in order to get the null distribution of the critical values.
+This is corroborated by an email exchange I had with author of StepwiseTest package (see file "Romano Wolf StepwiseTest package question email thread").
 \item Page 1396: Talks about other resampling methods in some detail, including Westfall
@@
 
-\subsection{What should we do about logistic regression?}
+\subsection{What does Westfall do for logistic regression?}
 
@@
 He just regresses each outcome on the Xs and then generates from a binomial using the fitted parameters.
-But this doesn't preserve the correlation between the Ys (see \texttt{westfall\_logistic\_regression.R}), so I don't understand.
-Also, he says you can't do case-resampling for logistic regression, but I don't know why -- unless it's the same issue about correlated residuals that seems potentially problematic in OLS.
+But this doesn't preserve the correlation between the Ys unless you condition on all covariates such that the Ys are independent (see \texttt{westfall\_logistic\_regression.R}).
 
@@
 \newpage
+\subsection{Exact variance under null}
+
+Let $p_w$ be the $w^{th}$ observed p-value. Then, under the strong null, we have:
+\begin{align*}
+\text{Var}\left( \widehat{\theta} \right) &= W \alpha \left( 1 - \alpha \right) + 2 \sum_{1 \le i < j \le W} P \left( p_i < \alpha, p_j < \alpha \right) - \alpha^2
+\end{align*}
+
+\begin{align*}
+\text{Var} \left( \widehat{\theta} \right) &= \text{Var} \left( \sum_{w=1}^W 1 \big\{ p_w < \alpha \big\} \right) \\
+&= \sum_{w=1}^W \text{Var} \left( 1 \big\{ p_w < \alpha \big\} \right) + 2 \sum_{1 \le i < j \le W} \text{Cov} \left( 1 \big\{ p_i < \alpha \big\}, 1 \big\{ p_j < \alpha \big\} \right) \\
+%--
+&= W \alpha \left( 1-\alpha \right) + 2 \sum_{1 \le i < j \le W} E \big[ 1 \big\{ p_i < \alpha, p_j < \alpha \big\} \big] - E \big[ 1 \big\{ p_i < \alpha \big\} \big] E \big[ 1 \big\{ p_j < \alpha \big\} \big] \\
+%--
+&= W \alpha \left( 1-\alpha \right) + 2 \sum_{1 \le i < j \le W} \Big[ \underbrace{P \left( p_i < \alpha, p_j < \alpha \right)}_{=\alpha^2 \text{ under independence}} - \alpha^2 \Big]
+\end{align*}
+
+<!-- NEW SECTION -->
+\newpage
 \section{Simulation notes}
 
-\subsection{Upcoming simulations}
+\subsection{"Marginally vs conditionally"}
 
-\subsection{"2018-3-18 with other methods"}
+Goal was to test the hypothesis that if outcomes are independent conditional on \emph{adjusted} covariate C, then the test stats for X are independent even though the outcomes are marginally correlated.
+However, if the outcomes are independent only conditional on \emph{unadjusted} U (as in below simulations), then the test stats for X are correlated.
+
+\subsection{"Validate FCR-center resampling"}
+
+Since realizing that MICE algorithm wouldn't work, I tried to validate FCR-center (i.e., do FCR, but then center each resampled test statistic by its counterpart in the original data) by generating data with 1 standard normal covariate and 1 unmeasured non-confounder that induces correlation between $Y_1$ (normal) and $Y_2$ (either normal or binary depending on scenario).
+The empirical (Sherlock) results for both scenarios are in \texttt{stitched.csv}.
+
+The simulation had two separate stages:
+
+\begin{enumerate}
+\item Locally, use \texttt{fcr\_center\_validation.R} to actually generate data under $H_0$ (i.e., override the beta for $X$ and set it to 0) for $10,000$ simulation reps and check the correlation of the $p$-values.
+This gives us the "true" correlations under $H_0$.
+\item On Sherlock, use \texttt{fcr\_center\_validation\_for\_sherlock.R} to generate, for each of $2,500$ simulation reps per scenario, an original dataset under $H_A$ and then use FCR-resample to produce $B=2,000$ resamples under $H_0$.
+Fit the analysis model to each resample.
+Compute the average correlation of the $p$-values for each simulation rep.
+\end{enumerate}
+
+\underline{Results:} When both outcomes are normal, FCR-center exactly recovers the correct $p$-value correlation (true correlation: 0.70, empirical correlation: 0.70).
+However, when 1 outcome is normal and 1 is binary, FCR-center does not work (true correlation: 0.31, empirical correlation: 0.20).
+
+\underline{Explanation:} I think this is because CI and hypothesis test are directly equivalent for OLS, but not for logistic regression.
+That's because for OLS, the SE is independent of the value of the parameter $\beta$ itself: it's just $\sigma^2(X'X)^{-1}$, and $\sigma$^2 is independent of $\beta$ for any fixed model specification.
+In contrast, with logistic regression, the SE depends on $\beta$, which is why its MLEs require iteration.
+Basically this is the same reason that the SE for a proportion is different when computed under $H_0$ vs.
+$H_A$, so the CI and test might not coincide.
+I did a separate simulation to check this in which I avoided any bootstrapping.
+I just directly generated data under $H_0$ or $H_A$ using \texttt{fcr\_center\_validation.R} and looked at whether the "true" SE of the coefficient estimates was the same for $H_0$ and $H_A$.
+Below, the 0.06 and 0.08 show the difference in SEs under $H_0$ and $H_A$ for a binary outcome.
+
+\hl{IS THIS THE FABLED SUBSET PIVOTALITY?}
+
+\begin{table}[H]
+\centering
+\label{my-label}
+\begin{tabular}{@{}llll@{}}
+\toprule
+Y2 type & $H_0$ & $SE(\widehat{\beta}_{Y_1})$ & $SE(\widehat{\beta}_{Y_2})$ \\ \midrule
+Binary  & T    & 0.10      & 0.06      \\
+Binary  & F    & 0.10      & 0.08      \\ \midrule
+Normal  & T    & 0.10      & 0.10      \\
+Normal  & F    & 0.10      & 0.10        \\ \bottomrule
+\end{tabular}
+\end{table}
+
+\subsection{2018-4-7 unsaved local simulations using \texttt{mice\_validation.R} }
+
+"2 normal outcomes":
+
+Critical: 
+If Y1 and Y2 are correlated because of U, but U is adjusted in the models for Y1 and Y2, then the test stats and p-values AREN’T correlated.
+
+If U is NOT adjusted in analysis (implicit in our current simulations), then the test stats and p-values ARE correlated.
+
+\subsection{2018-4-3}
+
+Reran all methods to compare. This is written up in the file "2018-4-3 Multiple Outcomes update #1".
+
+\subsection{2018-3-16 $\ast$real Romano}
+
+Resampling parametrically under original data distribution (\texttt{ha.resid}) and centering bootstrapped test stats by observed one in order to apply Romano.
+
+\subsection{"2018-3-15 add Romano-H0"}
+
+This is the same as the previous one, except adding Holm and a version of Romano that still uses \emph{null} resamples (\texttt{h0.parametric}).
+Note that this is not actually what he recommends in his paper (he resamples from the original data distribution, but then presumably centers the bootstrapped test statistics by the observed one).
+
+For paper, could combine these results (minus Romano) with previous one in order to have 1000 simulations per method (except Holm because it wasn't in the previous simulations).
+
+\subsection{"2018-3-18 $\ast$with other methods"}
 
@@
 
-<!--
-\newpage
-Here is a test reference[@evalue].
-Use semicolons to separate multiple references.
-To change reference style, Google for a .csl style sheet for journal of interest of one with similar reference style.
--->
-
 <!-- NEW SECTION -->

### 2018-05-13 Manuscript/manuscript.Rmd
About to remove old material

@@
 <!-------------------------- --------------------------------->
+<!------------- SECTION: TO DO -------------->
+<!-------------------------- --------------------------------->
+
+\section{To do}
+
+\begin{itemize}
+\item Connect the last theorem with number of rejections
+\end{itemize}
+
+<!-------------------------- --------------------------------->
 <!------------- SECTION: SETTING AND NOTATION -------------->
@@
 \item Ones borrowed directly from Freedman
+\item Covariates are mean-centered?
 \item All regressions have same design matrix and sample size. Do they all need to have same exposure of interest?
@@
 $\sigma^2_w$: see Lemma 2.1'
+
+<!-------------------------- REGULARITY ASSUMPTION ---------------------------->
+\begin{assump}
+\label{regularity_assump}
+As a regularity condition on the design matrix, assume that for some constant $k>0$:
+\begin{align*}
+N \cdot B'B &\xrightarrow[N \to \infty]{P} k \\
+\Leftrightarrow N \sum_{n=1}^N [(X'X)^{-1}X']^2_{ij} &\xrightarrow[n \to \infty]{P} k
+\end{align*}
+\end{assump}
+
+\begin{remark}
+To provide intuition, consider the assumption in the context of simple linear regression \textcolor{red}{($p=2$)} with a mean-centered exposure of interest, $X_1$.
+Define $S_1 = \sum_{n=1}^N X_{n1}$ and $S_2 = \sum_{n=1}^N X_{n1}^2$.
+Then we have:
+\begin{align*}
+X'X &= \begin{bmatrix}
+N & S_1 \\
+S_1 & S_2 
+\end{bmatrix}\\ \\
+(X'X)^{-1} &= \frac{1}{N S_2 - S_1^2} \begin{bmatrix}
+S_2 & -S_1 \\
+-S_1 & N 
+\end{bmatrix}\\ \\
+(X'X)^{-1}X' &= \frac{1}{N S_2 - S_1^2} \begin{bmatrix}
+S_2 - X_{11} S_1 & \dotsb & S_2 - X_{N1} S_1 \\
+-S_1 + X_{11}N & \dotsb & -S_1 + X_{N1} N 
+\end{bmatrix}\\
+\end{align*}
+
+The regularity condition concerns the term:
+\begin{align*}
+N \sum_{n=1}^N B_n^2 &= N \sum_{n=1}^N [(X'X)^{-1}X']^2_{n1} \\
+&= \frac{1}{ N S_2^2 - 2 S_2 S_1^2 + \frac{1}{N} S_2} \sum_{n=1}^N \left( X_{n1} N - S_1 \right)^2 \\
+&= \frac{1}{ N S_2^2 - 2 S_2 S_1^2 + \frac{1}{N} S_2} \sum_{n=1}^N \left( N^2 X_{n1}^2 - 2N X_{n1} S_1 + S_1^2 \right) \\
+&= \frac{N^2 S_2 - 2N S_1^2 + NS_1^2 }{N S_2^2 - 2 S_2 S_1^2 + \frac{1}{N} S_2} \\
+&= \frac{N^2 S_2 - NS_1^2 }{N S_2^2 - 2 S_2 S_1^2 + \frac{1}{N} S_2} \\
+&= \frac{\frac{S_2}{N} - \frac{S_1^2}{N^2} }{ \frac{S_2^2}{N^2} - 2 \frac{S_2}{N} \frac{S_1^2}{N^2} + \frac{S_2}{N^2} } \tag{$\div N^3$} \\
+&\xrightarrow[N \to \infty]{P} 1 / \sigma^2_1
+\end{align*}
+The last line follows because, by mean-centering, we have $S_1/N \xrightarrow[N \to \infty]{P} 0$ and $S_2/N \xrightarrow[N \to \infty]{P} \sigma^2_1$.
+Thus, for simple linear regression, Assumption \ref{regularity_assump} holds if $\sigma^2_1 > 0$.
+\end{remark}
 

### 2018-05-15 Manuscript/manuscript.Rmd
About to update main text

--- 2018-05-10 Manuscript/manuscript.Rmd
@@
 \begin{center}
-\textbf{ \LARGE{Multiple Testing Research Update \#1} } \\ \vspace{5mm}
+\textbf{ \LARGE{Multiple Testing Research Update \#2} } \\ \vspace{5mm}
 \textbf{ \LARGE{\today} }
@@
 <!-------------------------- --------------------------------->
-<!-------------- SECTION: SETTING AND NOTATION --------------->
+<!------------- SECTION: TO DO -------------->
+<!-------------------------- --------------------------------->
+
+\section{To do}
+
+\begin{itemize}
+\item \textbf{Explain why residual resampling works creates samples generated under strong null}
+\item \textbf{Think more broadly about proof and why it had to be so complicated}
+\item \textbf{Check Remark 2 again}
+\item \textbf{Fix references}
+\item Connect the last theorem with number of rejections
+\item Re-incorporate old material (e.g. general presentation)
+\item Incorporate any of TVW's old comments
+\end{itemize}
+
+Say somewhere in paper:
+\begin{itemize}
+\item Unlike Freedman, we assume the resample size is same as original sample size ($N$)
+\end{itemize}
+
+<!-------------------------- --------------------------------->
+<!------------- SECTION: INTRODUCTION -------------->
+<!-------------------------- --------------------------------->
+
+\section{Introduction}
+
+Contributions of this paper:
+
+\begin{itemize}
+\item The new metrics
+\item Theoretical justification for an existing resampling algorithm
+\item Joint test power comparison of existing methods and its interesting results wrt the "naïve" methods (they're actually pretty good because differences emerge mainly in adjustment of the larger p-values)
+\end{itemize}
+
+<!-------------------------- --------------------------------->
+<!------------- SECTION: SETTING AND NOTATION -------------->
 <!-------------------------- --------------------------------->
@@
 
-<!----- Old version
-\section{Setting and notation}
-
-Suppose that $K$ random variables are measured on $N$ subjects, with the resulting matrix denoted $\mathbf{Z} \in \mathbb{R}_{N \times K}$.
-Let $Z_{nk}$ denote, for the $n^{th}$ subject, the $k^{th}$ random variable.
-Consider a resampling algorithm that generates, for iterate $j$, a dataset containing the random vector $\left( Z_{n1}^{(j)}, \dotsb, Z_{nK}^{(j)} \right)$ for each subject $n$.
-There are a total of $B$ resampled datasets.
-
-We refer to cumulative distribution functions as $F$.
-Specifically, we refer to the joint distribution of the data as follows.
-With $\mathcal{B} \in \mathbb{R}_{N \times K}$ denoting a matrix of constants with entries $\left( b_{11}, \dotsb, b_{NK} \right)$, let:
-\begin{align*}
-F_{\mathbf{Z}} \left( \mathbf{B} \right) &= P \left( Z_{11} \le b_{11}, \dotsb, Z_{NK} \le b_{NK} \right)
-\end{align*}
-
-Define the joint CDF of $\left( Z_{n1}^{(j)}, \dotsb, Z_{nK}^{(j)} \right)$ as:
-
-\begin{align*}
-F_{\left( Z_{n1}^{(j)}, \dotsb, Z_{nK}^{(j)} \right)} \left( r_1, \dotsb, r_K \right) &= P \left( Z_{n1}^{(j)} < r_1, \dotsb, Z_{nK}^{(j)} < r_K \right)
-\end{align*}
-
-We will usually refer simply to $\mathbf{Z}$ and $F_{\mathbf{Z}}$ for brevity and will denote their counterparts in resampled dataset $j$ as $\mathbf{Z}^{(j)}$ and $F_{\mathbf{Z}^{(j)}}$.
-
-Further suppose that we conduct $W$ hypothesis tests, each with level $\alpha$.
-Let $c_{w,\alpha}$ be the critical value for the test statistic, $T_w$, of the $w^{th}$ test.
-We refer to the vector of test statistics as $\mathbf{T}$.
-Assume that each test concerns a point null hypothesis (\textbf{Assumption 1}) and that each test is parametric with consistent estimates of the tested parameters in the likelihood identifiable from the observed data (\textbf{Assumption 2}).
-We define the "strong null" as the case in which all $W$ point null hypotheses hold and use $F_{ \mathbf{Z}^0 }$ to denote the distribution of the data under the strong null.
-We use the superscript $^0$ generally to denote data or statistics generated under the strong null.
-
-Define the statistic corresponding to the total number of rejections in $W$ tests in a sample generated under the strong null as $\widehat{\theta}^0 = \sum_{w=1}^W 1 \big\{ T_w^0 > c_{w,\alpha} \big\}$.
-Similarly, the total number of rejections in resample $j$ is $\widehat{\theta}^{(j)} = \sum_{w=1}^W 1 \big\{ T_w^{(j)} > c_{w,\alpha} \big\}$.
-
-Respectively define the true distribution of the number of rejections under the strong null, its counterpart in the resamples, and its empirical estimator in the resamples as:
-\begin{align*}
-F_{\widehat{\theta}^0}(r) &= P \left( \widehat{\theta}^0 \le r \right) \\
-F_{\widehat{\theta}^{(j)}}(r) &= P \left( \widehat{\theta}^{(j)} \le r \right) \\
-\widehat{F}_{\widehat{\theta}^{(j)}} \left( r \right) &= \frac{1}{B} \sum_{j^{*}=1}^B 1 \Big\{ \widehat{\theta}^{(j^{*})} \le r \Big\}
-\end{align*}
-
-------->
-
 <!------------- --------------------------------->
@@
 We now show that under a certain class of resampling algorithms defined below, the empirical distribution of the number of rejections in the resamples converges to the true distribution of the number of rejections in samples generated under the strong null.
-Figure \ref{roadmap} depicts the structure of the proofs.
-
-\begin{figure}[H]
-\caption{Overview of theory to follow.
-Solid arrows denote ordinary limits in $B$, the dashed line denotes almost sure convergence in $N$, and the dotted line denotes convergence in probability in both $N$ and $B$.
-Parenthetical acronyms refer to an assumption, remark, lemma, and theorem developed below.}
-\label{roadmap}
-\begin{center}
-\begin{tikzpicture}
-  \matrix (m) [matrix of math nodes,row sep=3em,column sep=4em,minimum width=2em]
-  {
-     F_{ \mathbf{T}^{(j)} } & F_{ \mathbf{T}^0 } \\
-      %& F_{ \mathbf{Z} }
-      \\};
-  \path[-stealth]
-  % the - argument removes the arrowhead
-    (m-1-1) edge[] node [below] {(A3)} (m-1-2)
-
-    % diagonal line
-	%(m-1-1) edge[dashed] node [below] {(C1)} (m-2-2)
-
-    %(m-1-2) edge[dotted] node [right] {(L2)} (m-2-2)
-            ;
-\end{tikzpicture} % pic 1
-\qquad
-\begin{tikzpicture}
-  \matrix (m) [matrix of math nodes,row sep=3em,column sep=4em,minimum width=2em]
-  {
-     F_{ \widehat{\theta}^{(j)} } & F_{ \widehat{\theta}^0 }  \\
-      \widehat{F}_{ \widehat{\theta}^{(j)} } & \\};
-  \path[-stealth]
-  % the - argument removes the arrowhead
-    (m-1-1) edge[] node [below] {(L1)} (m-1-2)
-
-    % diagonal line
-	(m-2-1) edge[dashed] node [left] {(R2)} (m-1-1)
-
-    (m-2-1) edge [dotted] node [below] {(T1)} (m-1-2)
-            ;
-\end{tikzpicture}
-\end{center}
-\end{figure}
 
@@
 \label{cbt_def}
-The resampling algorithm used to generate $\left( Z_{n1}^{(j)}, \dotsb, Z_{nK}^{(j)} \right)$ must generate identically and independently distributed (i.i.d.) observations, and it must ensure that $\mathbf{T}^{(j)} \xrightarrow[B \to \infty]{D} \mathbf{T}^0$, i.e.:
+The resampling algorithm used to generate $\left( Z_{n1}^{(j)}, \dotsb, Z_{nK}^{(j)} \right)$ must generate observations that are independently and identically distributed (i.i.d.), that fulfill the parametric assumptions of all hypothesis tests, and it must ensure that $\mathbf{T}^{(j)} \xrightarrow[B \to \infty]{D} \mathbf{T}^0$, i.e.:
 $$\lim_{B \to \infty} F_{\mathbf{T}^{(j)}} = F_{\mathbf{T}^0}$$
@@
 \end{align*}
-This resampling algorithm fulfills Assumption \ref{cbt_def} (\hl{proof in progress}).
+This resampling algorithm fulfills Assumption \ref{cbt_def} (Appendix).
 Note that even if the original data were not generated under the strong null, this resampling algorithm guarantees that $E \big[ Z_{nw}^{(j)} \; \vert \; Z_{n3} \big] = E \big[ Z_{nw}^{(j)} \big]$ for $w \in \{1,2\}$, implying the strong null.
@@
 
-<!-------------- Remark 2 --------------->
-\bigskip \begin{remark}
-\label{glivenko}
-\normalfont
-By the i.i.d.
-construction of the resampled data (Assumption \ref{cbt_def}) and the Glivenko-Cantelli theorem, the empirical distribution of the number of rejections in the resamples converges almost surely in $B$ to its true distribution:
-$$\widehat{F}_{\widehat{\theta}^{(j)}} \xrightarrow[B \to \infty]{A.S.} F_{\widehat{\theta}^{(j)}} $$
-\end{remark}
-
-<!-------------- Fact 2 --------------->
-<!-- has been replaced by Glivenko-Cantelli -->
-<!--
-\bigskip \begin{fact}
-\label{glivenko}
-
-The empirical distribution of the number of rejections in the bootstrap samples converges almost surely in $B$ to its true distribution under the strong null:
-$$\widehat{F}_{\widehat{\theta}^{(j)}} \xrightarrow[B \to \infty]{A.S.} F_{\widehat{\theta}^{(j)}} $$
-\begin{proof}
-Since the $B$ bootstrap samples are i.i.d. and $E \big[ \vert \widehat{F}_{\widehat{\theta}^{(j)}} \left( r \right) \vert \big] < \infty$, this follows directly from the SLLN:
-%--
-\begin{align*}
-\widehat{F}_{\widehat{\theta}^{(j)}} \left( r \right) &= \frac{1}{B} \sum_{j^{*}=1}^B 1 \Big\{ \widehat{\theta}^{(j^{*})} \le r \Big\} \\
-& \xrightarrow[N \to \infty]{A.S.} E \Big[ 1 \big\{ \widehat{\theta}^{(j)} \le r \big\} \Big] \\
-&= P \left( \widehat{\theta}^{(j)} \le r \right) \\
-&= F_{\widehat{\theta}^{(j)}}(r)
-\end{align*}
-\end{proof}
-\hl{NOTE TO SELF: For paper, should just invoke Glivenko-Cantelli.}
-\end{fact}
--->
-
 <!-------------- Lemma 1 --------------->
@@
 \end{lemma}
-
-<!-------------- Theorem 1 --------------->
-\bigskip \begin{theorem} 
-\label{theorem1}
-The empirical distribution of the number of rejections in the resamples is consistent in $N$ and $B$ for the true distribution of the number of rejections under the strong null:
-$$\widehat{F}_{\widehat{\theta}^{(j)}} \xrightarrow[B, N \to \infty]{P} F_{\widehat{\theta}^0} $$
-\begin{proof}
-Fix an arbitrarily small $\epsilon > 0$ and $\epsilon' > 0$. Remark \ref{glivenko} implies convergence in probability, so there exists a $B^{*}$ such that, for all $B > B^{*}$:
-\begin{align*}
-P \left( \vert \widehat{F}_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^{(j)} } \vert > \epsilon / 2 \right) &< \epsilon'
-\end{align*}
-Similarly, from Lemma \ref{lemma3}, there exists an $N^{*}$ such that, for all $N > N^{*}$:
-\begin{align*}
-\left\vert F_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^0 } \right\vert &< \epsilon / 2
-\end{align*}
-Thus, for all $B$ and $N$ such that $B > B^{*}$ and $N > N^{*}$:
-\begin{align*}
-P \left( \vert \widehat{F}_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^0 } \vert > \epsilon \right) &= P \left( \vert  \widehat{F}_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^{(j)} } + F_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^0 } \vert > \epsilon \right) \\
-&\le P \left( \vert \widehat{F}_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^{(j)} } \vert + \vert F_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^0 } \vert > \epsilon \right)  \tag{triangle inequality}\\
-&\le P \left( \vert \widehat{F}_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^{(j)} } \vert > \epsilon/2 \; \text{ or } \; \vert F_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^0 } \vert > \epsilon/2 \right) \\
-&\le P \left( \vert \widehat{F}_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^{(j)} } \vert > \epsilon/2 \right) + P \left( \vert F_{ \widehat{\theta}^{(j)} } - F_{ \widehat{\theta}^0 } \vert > \epsilon/2 \right) \\
-&< \epsilon' + 0
-\end{align*}
-\end{proof}
-\end{theorem}
-
-<!-------------------------- --------------------------------->
-<!-------------- SECTION: GIBBS --------------->
-<!-------------------------- --------------------------------->
-
-\section{A general Gibbs resampling algorithm}
-
-For simple modeling choices and study designs, it is fairly easy to specify a resampling algorithm that satisfies Assumption \ref{cbt_def} (for example, @westfalltext provide numerous such algorithms).
-However, designing valid ad hoc algorithms can be challenging for more complex settings, such as generalized linear models with non-identity links [@westfalltext].
-Thus, here, we outline a general resampling algorithm that can capture complex dependence structures for \hl{arbitrary} analysis models, as long as the user is willing to assume a parametric structure for the relationships between outcomes\footnote{This method also assumes correct specification of the relationships between the outcomes and the covariates, but these assumptions are already implicit in the analysis model, so do not represent additional assumptions.}.
-
-Let $\mathbf{Y} = \left( \mathbf{Y}_1, \cdots, \mathbf{Y}_p \right)$ be a random vector of $p$ outcomes and $\mathbf{C}$ be a random vector of auxiliary covariates whose relationships with $\mathbf{Y}$ are not restricted under the strong null.
-Let $\mathbf{Y}^{\text{obs}}$ and $\mathbf{C}^{\text{obs}}$ denote their realized counterparts in the observed dataset.
-Let $\boldsymbol{\alpha}$ be a vector of all parameters characterizing the joint distribution of $\left( \mathbf{Y}, \mathbf{C} \right)$.
-(For example, if the joint distribution is conveniently specified in terms of a series of conditional densities, then $\boldsymbol{\alpha}$ could comprise regression parameters of each outcome in $\mathbf{Y}$ on each covariate in $\mathbf{C}$ and on the other outcomes.) Using $f^0$ to denote PDFs under the strong null, assume a uniform prior on $\boldsymbol{\alpha}$, such that $f^0 \left( \boldsymbol{\alpha} \right) \propto 1$.
-Then, fixing $\mathbf{C}$ as in Example \ref{ex:ols}, the posterior predictive distribution for a new vector of outcomes is:
-
-\begin{align*}
-f^0 \left( \mathbf{Y}^{(j)} \; \vert \; \mathbf{Y}^{\text{obs}}, \mathbf{C}^{\text{obs}} \right) &= \int_{A} f^0 \left( \mathbf{Y}^{(j)} \; \vert \; \mathbf{Y}^{\text{obs}}, \mathbf{C}^{\text{obs}}, \boldsymbol{\alpha} \right) f^0 \left( \boldsymbol{\alpha} \; \vert \; \mathbf{Y}^{\text{obs}}, \mathbf{C}^{\text{obs}} \right) \partial \boldsymbol{\alpha} \\
-&\propto \int_{A} f^0 \left( \mathbf{Y}^{(j)} \; \vert \; \mathbf{Y}^{\text{obs}}, \mathbf{C}^{\text{obs}}, \boldsymbol{\alpha} \right) f^0 \left( \mathbf{Y}^{\text{obs}}, \mathbf{C}^{\text{obs}} \; \vert \; \boldsymbol{\alpha} \right) \partial \boldsymbol{\alpha} \tag{uniform prior}
-\end{align*}
-
-To draw $\mathbf{Y}^{(j)}$, a generic Gibbs sampler could proceed as follows [@gelman]:
-\begin{enumerate}
-\item{Specify fully conditional distributions for each outcome $Y_i$:
-\begin{align*}
-Y_1 &\sim f^0 \left( Y_1 \; \vert \; \boldsymbol{\alpha}_1, \mathbf{C}^{\text{obs}}, \mathbf{Y}_{-1} \right) \\
-&\;\;\vdots \notag \\
-Y_p &\sim f^0 \left( Y_p \; \vert \; \boldsymbol{\alpha}_p, \mathbf{C}^{\text{obs}}, \mathbf{Y}_{-p} \right) \\
-\end{align*}
-where $\mathbf{Y}_{-i}$ represents all outcomes in $\mathbf{Y}$ except for $Y_i$ itself, and where $\boldsymbol{\alpha}_i$ represents the subset of parameters in $\boldsymbol{\alpha}$ that characterize the joint distribution of $\left( Y_i, \mathbf{C} \right)$.
-}
-
-\item{Obtain initial values of $\boldsymbol{\alpha}$ (i.e., $\boldsymbol{\alpha}^{(1)}$) for the first Gibbs iterate, for example by using MLEs from the observed data.}
-
-\item{For iterate $j$, successively draw parameters and predictive values for each outcome:
-\begin{align*}
-\boldsymbol{\alpha}_1^{(j)} &\sim f^0 \left( \boldsymbol{\alpha}_1 \; \vert \; \mathbf{Y}^{(j-1)}, \mathbf{C}^{\text{obs}} \right) \\
-Y_1^{(j)} &\sim f^0 \left( Y_1 \; \vert \; \boldsymbol{\alpha}_1^{(j)}, \mathbf{Y}_{-1}^{(j-1)}, \mathbf{C}^{\text{obs}} \right) \\
-&\;\;\vdots \notag \\
-\boldsymbol{\alpha}_p^{(j)} &\sim f^0 \left( \boldsymbol{\alpha}_p \; \vert \; \mathbf{Y}^{(j-1)}, \mathbf{C}^{\text{obs}} \right) \\
-Y_p^{(j)} &\sim f^0 \left( Y_p \; \vert \; \boldsymbol{\alpha}_p^{(j)}, \mathbf{Y}_{-p}^{(j-1)}, \mathbf{C}^{\text{obs}} \right)
-\end{align*}
-}
-\end{enumerate}
-
-By the \hl{usual} results (\hl{LAME}) regarding Gibbs sampling and Bayesian estimation of the posterior predictive distribution [@gelman], this algorithm satisfies the convergence criterion in Remark \ref{CMT_results}.
-
-In practice, this algorithm can be easily implemented by repurposing existing software for missing data imputation by chained equations (e.g., @micepackage).
-To do so, the observed dataset is copied while setting to missing all values of $\mathbf{Y}$, and this new dataset with "missing" values is concatenated with the observed one, yielding a pseudo-dataset with $2N$ observations (half of which have missing outcomes).
-Then, existing software can be used to create $B$ imputations of the "missing" data, with the imputed portion of each of $B$ imputations (the last $N$ observations of the pseudo-datasets) becoming the resampled datasets.
-Critically, the user must be careful to specify the "imputation" models such that the strong null is enforced; in contrast, typical software defaults use all variables in the dataset to impute all other variables, which would not enforce the strong null.
 
@@
 
+<!-------------------------- --------------------------------->
+<!-------------- SECTION: APPENDIX --------------->
+<!-------------------------- --------------------------------->
+
+\newpage
+
+\begin{center}
+\textbf{ \LARGE{Appendix} } \\ \vspace{5mm}
+\vspace{10mm}
+\end{center}
+
+\section{Validity of residual resampling for OLS regression}
+
+\subsection{Regression setting and notation}
+
+Here we modify notation in the main text to accommodate the OLS setting.
+In general, we will usually use superscripts to denote lengths and subscripts to denote indices.
+$[\cdot]_{ij}$ denotes the $(i,j)^{th}$ element of a matrix, and $[\cdot]_{i}$ denotes the $i^{th}$ element of a vector.
+Suppose that each of $W$ outcome variables is regressed on the same $N \times p$ design matrix, $X$.
+Assume that $X$ contains an intercept term, such that the residuals have mean 0, and that all other covariates are mean-centered.
+Let $\epsilon_w = \left( \epsilon_{1w}, \cdots, \epsilon_{Nw} \right)$ denote the $N$-vector of true errors for the $w^{th}$ regression and $\widehat{\epsilon}_w$ its estimated counterpart (the residuals).
+Let $\sigma^2_w = E \big[ \epsilon_{nw}^2 | X \big]$ as usual.
+Let $\beta^W = \left( \beta_1, \dotsb, \beta_W \right)$ be a vector containing, for each of $W$ regression models, the single coefficient of interest, and let $\widehat{\beta}^W$ denote its sample estimate.
+(We use the superscript to remind the reader that this is not the usual $p$-vector of coefficients from a single multiple regression model.) The $W$ test statistics of interest are the usual $Z$-statistics $\left( \frac{\widehat{\beta}_1}{  \widehat{\sigma}^2_1 \left( X'X\right)^{-1} }, \dotsb, \frac{\widehat{\beta}_W}{  \widehat{\sigma}^2_W \left( X'X\right)^{-1} } \right)$.
+
+Additionally, in order to control the
+
+assume the following regularity condition on the design matrix, which allows :
+
+bookmark
+
+<!-------------------------- REGULARITY ASSUMPTION ---------------------------->
+
+\begin{assump}
+\label{regularity_assump}
+Let $i \in \{ 2, \dotsb, p\}$ be the index of the regression covariate of interest and $B \in \mathbb{R}^{N\times1}$ be the transposed $i^{th}$ row of $(X'X)^{-1}X'$, or equivalently the $i^{th}$ column of $X (X'X)^{-1}$.
+Assume that for some constant $k>0$:
+\begin{align*}
+N \cdot B'B &\xrightarrow[N \to \infty]{P} k \\
+\Leftrightarrow N \sum_{n=1}^N [(X'X)^{-1}X']^2_{in} &\xrightarrow[N \to \infty]{P} k
+\end{align*}
+\end{assump}
+
+\begin{remark}
+To provide intuition, consider the assumption in the context of simple linear regressions ($p=2$), with $X_2$ denoting the exposure of interest.
+Define $S_1 = \sum_{n=1}^N X_{n2}$ and $S_2 = \sum_{n=1}^N X_{n2}^2$.
+Then, for each of $W$ such models, we have:
+\begin{align*}
+X'X &= \begin{bmatrix}
+N & S_1 \\
+S_1 & S_2 
+\end{bmatrix}\\ \\
+(X'X)^{-1} &= \frac{1}{N S_2 - S_1^2} \begin{bmatrix}
+S_2 & -S_1 \\
+-S_1 & N 
+\end{bmatrix}\\ \\
+(X'X)^{-1}X' &= \frac{1}{N S_2 - S_1^2} \begin{bmatrix}
+S_2 - X_{11} S_1 & \dotsb & S_2 - X_{N1} S_1 \\
+-S_1 + X_{11}N & \dotsb & -S_1 + X_{N1} N 
+\end{bmatrix}\\
+\end{align*}
+
+The regularity condition concerns the term:
+\begin{align*}
+N \sum_{n=1}^N B_n^2 &= N \sum_{n=1}^N [(X'X)^{-1}X']^2_{2n} \\
+&= \frac{1}{ N S_2^2 - 2 S_2 S_1^2 + \frac{1}{N} S_2} \sum_{n=1}^N \left( X_{n2} N - S_1 \right)^2 \\
+&= \frac{1}{ N S_2^2 - 2 S_2 S_1^2 + \frac{1}{N} S_2} \sum_{n=1}^N \left( N^2 X_{n2}^2 - 2N X_{n2} S_1 + S_1^2 \right) \\
+&= \frac{N^2 S_2 - 2N S_1^2 + NS_1^2 }{N S_2^2 - 2 S_2 S_1^2 + \frac{1}{N} S_2} \\
+&= \frac{N^2 S_2 - NS_1^2 }{N S_2^2 - 2 S_2 S_1^2 + \frac{1}{N} S_2} \\
+&= \frac{\frac{S_2}{N} - \frac{S_1^2}{N^2} }{ \frac{S_2^2}{N^2} - 2 \frac{S_2}{N} \frac{S_1^2}{N^2} + \frac{S_2}{N^2} } \tag{$\div N^3$} \\
+&\xrightarrow[N \to \infty]{P} 1 / \sigma^2_w
+\end{align*}
+The last line follows because, by mean-centering, we have $S_1/N \xrightarrow[N \to \infty]{P} 0$ and $S_2/N \xrightarrow[N \to \infty]{P} \sigma^2_w$.
+Thus, for simple linear regressions, Assumption \ref{regularity_assump} holds if $\sigma^2_w > 0$ for all $w \in \{ 1, \cdots, W\}$.
+\end{remark}
+
+We will consider validity of the bootstrap in terms of convergence on the Mallows-Wasserstein metric, a conventional choice that is defined as follows [@dasgupta; @freedman].
+
+\begin{definition}
+\label{mallows}
+Let $G_A$ and $G_B$ be arbitrary marginal distribution functions for random vectors $A \in \mathbb{R}^W$ and $B \in \mathbb{R}^W$, respectively.
+Then the Mallows-Wasserstein distance between $G_A$ and $G_B$ is the infimum, taken over all possible joint distributions for $(A,B)$ such that $A \sim G_A$ and $B \sim G_B$ marginally, of the expected squared Euclidean norm between $A$ and $B$:
+
+\begin{align*}
+d_2^W \left( G_A, G_B \right) & := \inf\limits_{ \substack{A \sim G_A \\ B \sim G_B} } E \big[ || A - B ||^2 \big]
+\end{align*}
+\end{definition}
+
+We proceed to prove that the residual-resampling bootstrap is consistent with respect to the Mallows-Wasserstein metric in a development that closely follows @freedman, who considered the asymptotic validity of residual resampling in recovering the sampling distribution of a $p$-vector of coefficient estimates from a single multiple linear regression model.
+Here, we extend this work to consider the sampling distribution of $\widehat{\beta}^W$.
+We first establish a series of lemmas paralleling those of @freedman and @bickel.
+
+<!-------------------------- LEMMA 8.9' ---------------------------->
+
+\begin{lemma}[Lemma 8.9']
+\label{trace_bound}
+Let $C$ and $D$ be $W \times N$ random matrices, and let $B$ be a fixed $N$-vector. Let $G_C$ and $G_D$ be the marginal distribution functions of the $W$-vectors $CB$ and $DB$, respectively. Then: 
+\begin{align*}
+d_2^W \left( G_C, G_D \right)^2 &\le \text{tr} \{ BB' \cdot E \big[ (C-D)'(C-D) \big] \}
+\end{align*}
+
+\begin{proof}
+\begin{align*}
+d_2^W \left( G_C, G_D \right)^2 &\le E \Big[ || CB - DB ||^2 \Big]\\
+&= E \Big[ \text{tr} \{ \underbrace{(CB-DB)}_{W \times 1} \underbrace{(CB-DB)'}_{1 \times W} \} \Big] \\
+&= E \Big[ \text{tr} \{ \underbrace{(C-D)}_{W \times N} \underbrace{BB'}_{N \times N} \underbrace{(C-D)'}_{N \times W} \} \Big] \\
+&= E \Big[ \text{tr} \{ \underbrace{ BB' (C-D)'(C-D) }_{N \times N} \} \Big] \tag{trace cyclical invariance}\\
+&= \text{tr} \{ E \Big[ BB' (C-D)'(C-D) \Big] \}  \\
+&= \text{tr} \{ BB' \cdot E [ \underbrace{ (C-D)'(C-D) }_{N \times N} ] \} \tag{$B$ is fixed} \\
+\end{align*}
+\textcolor{red}{In the inequality, the LHS is the infimum of the expectation over all possible pairs of random variables with the marginal distributions $G_C$ and $G_D$, whereas the RHS is the expectation for a particular pair of random variables (see similar step in Bickel \& Freedman pg 1214, Lemma 8.9).}
+\end{proof}
+\end{lemma}
+
+<!-------------------------- THEOREM 2.1'---------------------------->
+The next theorem bounds the distance between the true sampling distribution of the test statistics and the estimated sampling distribution in the resamples in terms of the distance between the sampling distribution of the true errors and the resampled residuals.
+
+\begin{theorem}[Theorem 2.1']
+\label{bound_thm}
+Paralleling @freedman's notation, let $F$ denote the true distribution function of the true errors, $\left( \epsilon_{n1}, \cdots, \epsilon_{nW} \right)$, and let $\widehat{F}_N$ denote the empirical distribution function of the residuals, which is used to approximate $F$ in residual resampling.
+Let $\Psi(F)$ denote the distribution of the standardized coefficient estimates, $\sqrt{N} \left( \widehat{\beta}^W - \beta^W \right)$ that are constructed as a function of the true error distribution; $\Psi(F)$ therefore represents the true sampling distribution to which a valid bootstrapped sampling distribution must converge.
+In contrast, let $\Psi(\widehat{F}_N)$ the distribution of the standardized coefficient estimates in which the empirical distribution of the residuals is substituted for the true distribution, i.e.
+$\sqrt{N} \left( \widehat{\beta}^{W(j)} - \widehat{\beta}^W \right)$\footnote{Note that the resampled test statistics are BOOKMARK, }.
+
+As in Assumption \label{regularity_assump}, let $B \in \mathbb{R}^{N\times1}$ be the $i^{th}$ column of $(X'X)^{-1}X'$. Then:
+\begin{align*}
+d_2^W \left( \Psi(F), \Psi(\widehat{F}_N) \right)^2 &\le N \cdot \text{tr} \{ BB' \} \cdot d_2^W \left( F, \widehat{F}_N \right)^2
+\end{align*}
+\begin{proof}
+In general for multiple regression, we have $\widehat{\beta} - \beta = \left( X'X\right)^{-1} X' \epsilon$.
+Letting:
+\begin{align*}
+C \in \mathbb{R}^{W \times N} =
+\begin{bmatrix}
+    \text{---} \hspace{-0.2cm} & \epsilon_1' & \hspace{-0.2cm} \text{---} \\
+    & \vdots & \\
+    \text{---} \hspace{-0.2cm} & \epsilon_W' & \hspace{-0.2cm} \text{---}
+\end{bmatrix}
+\end{align*}
+we can express $\Psi(F)$ as the distribution of the $W$-vector: 
+\begin{align*}
+\sqrt{N} \left( \widehat{\beta}^W - \beta^W \right) &= \sqrt{N} \begin{bmatrix}
+    [ (X'X)^{-1}X' \epsilon_1 ]_{i}\\
+    \vdots\\
+    [ (X'X)^{-1}X' \epsilon_W ]_{i}
+\end{bmatrix} = \sqrt{N} \begin{bmatrix}
+    \epsilon_1' B \\
+    \vdots\\
+    \epsilon_W' B
+\end{bmatrix} =
+\sqrt{N} \cdot CB
+\end{align*}
+whose $w^{th}$ element pertains to the $i^{th}$ regression coefficient in the $w^{th}$ regression.
+Now let $D$ be the counterpart of $C$ with the residual vectors, $\widehat{\epsilon}_w$, in place of the true errors, $\epsilon_w$.
+Then the entries of matrix $(C-D)'(C-D)$ are:
+\begin{align*}
+[(C-D)'(C-D)]_{kj} &= \sum_{w=1}^W [(C-D)']_{kw} [C-D]_{wj} \\
+&= \sum_{w=1}^W [C-D]_{wk} [C-D]_{wj} \\
+&= \sum_{w=1}^W \left( \epsilon_{jw} - \widehat{\epsilon}_{jw} \right) \left( \epsilon_{kw} - \widehat{\epsilon}_{kw} \right)
+\end{align*}
+We have $E[ \left( \epsilon_{jw} - \widehat{\epsilon}_{jw} \right) \left( \epsilon_{kw} - \widehat{\epsilon}_{kw} \right) ] = \text{Cov} \left( \epsilon_{jw} - \widehat{\epsilon}_{jw}, \epsilon_{kw} - \widehat{\epsilon}_{kw} \right)$, but for all $k \ne j$, the covariance is 0 because the observations are independent.
+Thus, letting $I^{N}$ denote the $N \times N$ identity matrix, we have that $E[(C-D)'(C-D)]$ is a diagonal matrix equal to $I^{N} \cdot E \Big[ \sum_{w=1}^W \left( \epsilon_{jw} - \widehat{\epsilon}_{jw} \right)^2 \Big] = I^{N} \cdot d_2^W \left( \widehat{F}_N, F\right)$.
+\textcolor{red}{I took the latter equality from B\&F Lemma 8.9 but don't understand it.
+I traced it back to the proof of Lemma 8.6 (pg 1213), in which this equality is stated to hold "without loss of generality in view of Lemma 8.1", but I still don't understand.} The result then follows immediately from applying Lemma \ref{trace_bound}, setting $G_C = F$, $G_D = \widehat{F}_N$ and $B, C,$ and $D$ as defined above and pulling the scalar $\sqrt{N}$ outside the squared distance.
+\end{proof}
+\end{theorem}
+
+To apply the bound in Theorem \ref{bound_thm}, we will first bound the term on the right-hand side using a triangle inequality, which applies because $d_2^W(\cdot, \cdot)$ is a metric [@bickel].
+To this end,  let $F_N$ denote the empirical distribution function of the true error vector, $\epsilon^W$.
+Then we have the following triangle inequality:
+\begin{align}
+\label{triangle}
+d_2^W \left( \widehat{F}_N, F \right)^2 \le d_2^W \left( \widehat{F}_N, F_N \right)^2 + d_2^W \left( F_N, F \right)^2 
+\end{align}
+The first term on the RHS relates the empirical distribution of the residuals to the empirical distribution of the true errors (which are discrete distributions taking $N$ values); the second term relates the latter empirical distribution to the true error distribution (which is continuous by standard OLS assumptions).
+
+\begin{lemma}[Lemma 2.1']
+\label{first_bound}
+Concerning the first term on the RHS of Equation (\ref{triangle}):
+\begin{align*}
+d_2^W \left( \widehat{F}_N, F_N\right)^2 \xrightarrow[N \to \infty]{P} 0
+\end{align*}
+
+\begin{proof}
+As in Definition \ref{mallows}, let $U \sim \widehat{F}_N$ and $V \sim F_N$ be arbitrary random variables in $\mathbb{R}^W$ that follow the empirical marginal distributions of the residuals and true errors.
+Denote their elements $\left( U_1, \dotsb, U_W \right)$ and $\left( V_1, \dotsb, V_W \right)$.
+\textcolor{red}{Let $(\widetilde{U}, \widetilde{V})$ be a special choice of $(U,V)$ that follow not only the marginal empirical distributions $\widehat{F}_N$ and $F_N$, but also the empirical \emph{joint} distribution of the residuals and the true errors.}
+
+\begin{align*}
+d_2^W \left( \widehat{F}_N, F_N\right)^2 & := \inf\limits_{ \substack{U \sim \widehat{F}_N \\ V \sim F_N} } E \big[ || U - V ||^2 \big] \\
+& \textcolor{red}{\le E \Big[ || \widetilde{U} - \widetilde{V} ||^2 \Big] \tag{\text{choose one element from inf set}} }\\
+& = \frac{1}{N} \sum_{n=1}^N \underbrace{ \sum_{w=1}^W \left( \widehat{\epsilon}_{nw} - \epsilon_{nw} \right)^2 }_{||\cdot||^2 \text{ of a $W$-vector}} \tag{\text{joint ECDF expectation}}\\
+& = \frac{1}{N} \sum_{w=1}^W \underbrace{ \sum_{n=1}^N \left( \widehat{\epsilon}_{nw} - \epsilon_{nw} \right)^2 }_{||\cdot||^2 \text{ of an $N$-vector}}\\
+&= \frac{1}{N} \sum_{w=1}^W || \widehat{\epsilon}_w - \epsilon_w ||^2 \\
+\Rightarrow E \Big[ d_2^W \left( \widehat{F}_N, F_N \right)^2 \Big] &= \frac{p}{N} \sum_{w=1}^W \sigma^2_w \tag{\text{Freedman Eq. (2.2)}} \\
+& \xrightarrow[N \to \infty]{} 0
+\end{align*}
+The interchange of summations in lines 3-4 is used to express $W$ norms involving residuals from different regressions, summed over $N$ observations, as $N$ norms involving residuals of observations within a regression, summed over $W$ regressions.
+The latter is more convenient because it allows application of existing theory for a single multiple regression model.
+
+\end{proof}
+\end{lemma}
+
+\begin{lemma}
+\label{second_bound}
+Concerning the expectation of the second term on the RHS of Equation (\ref{triangle}):
+\begin{align*}
+E \Big[ d_2^W \left( F_N, F \right)^2 \Big] \xrightarrow[N \to \infty]{P} 0
+\end{align*}
+
+\begin{proof}
+Letting $P_N$ denote an empirical probability, $F_N$ can be expressed as:
+\begin{align*}
+P_N \left( \epsilon_{n1} \le c_1, \dotsb, \epsilon_{nW} \le c_W \right) &= \frac{1}{N} \sum_{n=1}^N 1 \big\{ \epsilon_{n1} \le c_1, \dotsb, \epsilon_{nW} \le c_W \big\} \\
+&\xrightarrow[N \to \infty]{P} P \left( \epsilon_{n1} \le c_1, \dotsb, \epsilon_{nW} \le c_W \right) \tag{SLLN}
+\end{align*}
+which is the joint cumulative distribution corresponding to $F$. By @bickel's Lemma 8.3, this convergence in probability implies the desired result.
+\end{proof}
+\end{lemma}
+
+\begin{theorem}
+The residual bootstrap is weakly consistent under the Mallows-Wasserstein metric for the OLS coefficient estimates [@dasgupta]; that is:
+\begin{align*}
+d_2^W \left( \Psi(F), \Psi(\widehat{F}_N) \right)^2 \xrightarrow[N \to \infty]{P} 0
+\end{align*}
+
+\begin{proof}
+
+Combining Theorem \ref{bound_thm} with the triangle inequality in Equation (\ref{triangle}) and observing that $\text{tr} \{ BB' \} = \sum_{n=1}^N B_N^2 \ge 0$ yields:
+\begin{align*}
+d_2^W \left( \Psi(F), \Psi(\widehat{F}_N) \right)^2 &\le N \cdot \text{tr} \{ BB' \} \cdot \left( d_2^W \left( \widehat{F}_N, F_N \right)^2 + d_2^W \left( \widehat{F}_N, F \right)^2 \right)
+\end{align*}
+
+The term $N \cdot \text{tr} \{ BB' \} \xrightarrow[N \to \infty]{P} k$ by Assumption \ref{regularity_assump}.
+By Lemma \ref{first_bound}, $d_2^W \left( \widehat{F}_N, F \right)^2 \xrightarrow[N \to \infty]{P} 0$.
+Last, the convergence in $r^{th}$ mean of Lemma \ref{second_bound} implies that $d_2^W \left( F_N, F \right)^2 \xrightarrow[N \to \infty]{P} 0$, so the desired result holds.
+
+Big picture (bookmark)
+
+This shows that test stats generated as a function of the ECDF of the residuals ($\widehat{F}_N$) converge to test stats generated from the true error distribution ($F$).
+The bootstrap residual distribution can be made arbitrarily close to $\widehat{F}_N$ by taking $B \to \infty$, so by convention [@freedman], we ignore this source of error.
+
+Null issue
+
+\section{In progress: connection to number of rejections}
+The estimated standard errors are $\left( (X'X)^{-1}\widehat{\sigma}^2_1, \dotsb, (X'X)^{-1}\widehat{\sigma}^2_W \right)$, which converge to a point mass at $\left( (X'X)^{-1}\sigma^2_1, \dotsb, (X'X)^{-1}\sigma^2_W \right)$, so the desired convergence result holds for the joint distribution of the test statistics $\frac{ \widehat{\beta}_w }{ (X'X)^{-1} \widehat{\sigma}^2_w }$ (see @freedman's Theorem 2.2)
+\end{proof}
+\end{theorem}
+
+* Getting to number of rejections:
+
+\newpage
 \section*{References}

### 2018-05-16 R code/For Sherlock/doParallel_mice.R
Clean up

@@ -1,371 +1,0 @@
-
-######### FOR CLUSTER USE #########
-# load command line arguments
-args = commandArgs(trailingOnly = TRUE)
-jobname = args[1]
-scen = args[2]  # this will be a letter
-boot.reps = as.numeric( args[3] )
-
-# get scen parameters
-setwd("/share/PI/manishad/multTest")
-scen.params = read.csv( "scen_params.csv" )
-p = scen.params[ scen.params$scen.name == scen, ]
-
-print(p)
-
-# no longer included in parameters because it's a vector
-crit.bonf = 0.05 / p$nY
-alpha = c( crit.bonf, 0.01, 0.05 )
-
-# simulation reps to run within this job
-# this need to match n.reps.in.doParallel in the genSbatch script
-sim.reps = 1
-######### END OF CLUSTER PART #########
-
-
-# ######### FOR LOCAL USE #########
-# # setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/Sandbox/2018-1-13")
-# # p = read.csv("scen_params.csv")  # should be a single row, I think
-# 
-# n = 100
-# nX = 1
-# nY = 4
-# rho.XX = 0
-# rho.YY = c(0.25)
-# rho.XY = c(0.08)  # null hypothesis: 0
-# half = 0
-# 
-# # bootstrap iterates and type
-# boot.reps = 5
-# sim.reps = 2
-# scen = "a"
-# bt.type = c( "MICE.H0" )
-# 
-# 
-# # matrix of scenario parameters
-# scen.params = expand.grid( bt.type, n, nX, nY, rho.XX,
-#                            rho.YY, rho.XY, half )
-# names(scen.params) = c( "bt.type", "n", "nX", "nY", "rho.XX",
-#                         "rho.YY", "rho.XY", "half" )
-# 
-# # name the scenarios
-# # remove letters that are privileged variables in R
-# letter.names = c(letters, LETTERS)[ ! c(letters, LETTERS) %in% c("i","T","F") ]
-# scen.params$scen.name = letter.names[ 1:dim(scen.params)[1] ]
-# p = scen.params
-# 
-# # add alpha corresponding to Bonferroni as first one
-# # NOTE THAT CODE ASSUMES BONFERRONI IS THE FIRST ONE
-# crit.bonf = 0.05 / p$nY
-# alpha = c( crit.bonf, 0.01, 0.05 )
-# 
[…]

### 2018-05-16 R code/For Sherlock/functions_mice.R

@@ -1,662 +1,0 @@
-
-
-
-######################## FNS FOR WESTFALL's SINGLE-STEP ########################
-
-# returns minP-adjusted p-values (single-step)
-
-# p = original p-values
-# p.bt = bootstrapped p-values (vector) - an m X B matrix
-adjust_minP = function( p, p.bt ) {
-  
-  n.boot = ncol(p.bt)
-  
-  # keep only minimum p-value in each resample
-  minP.bt = apply( p.bt, MARGIN = 2, FUN = min )
-  
-  # for each element of p, get the proportion of resamples
-  #  whose minP was less than the present p-value
-  p.adj = unlist( lapply( p, FUN = function(x) sum( minP.bt <= x ) / n.boot ) )
-  
-  return(p.adj)
-}
-
-# # test
-# B = 200
-# n.tests = 10
-# p.bt = matrix( runif(B*n.tests, 0, 1), nrow = n.tests)
-# p = runif( n.tests, 0, .1)
-# 
-# p.adj = adjust_minP( p, p.bt )
-# plot(p,p.adj)
-
-
-######################## FNS FOR WESTFALL's STEP-DOWN ########################
-
-# Westfall textbook, pages 66-67
-adj_Wstep = function( p, p.bt ) {
-  
-  # # TEST ONLY
-  # p = structure(c(0.636112475277457, 0.116408752390887, 0.260603159221029,
-  #                 0.575086408582202, 0.625789849782125), .Dim = c(5L, 1L))
-  # p.bt = structure(c(0.728140113074014, 0.828557135138776, 0.728464927503583,
-  #                    0.376518353153738, 0.241363541131996, 0.415330227654616, 0.75686321503498,
-  #                    0.865494617532842, 0.839819273240726, 0.540707408457488, 0.149482739547785,
-  #                    0.506274953008223, 0.122605643925883, 0.747018941160177, 0.168076477020114,
-  #                    0.0582566955778811, 0.284459435324328, 0.561050485854955, 0.13839444320949,
-  #                    0.641335055023296, 0.195313599009888, 0.174692123827015, 0.982035588216878,
-  #                    0.713525352129964, 0.548770486028363), .Dim = c(5L, 5L), .Dimnames = list(
-  #                      NULL, c("result.1", "result.2", "result.3", "result.4", "result.5"
-  #                      )))
-  
-  # if using read-in or simulated dat
-  # p = pvals
-  
-  
[…]

### 2018-05-16 R code/For Sherlock/gen_sbatch_mice.R

@@ -1,99 +1,0 @@
-
-
-########################### SET SIMULATION PARAMETERS MATRIX ###########################
-
-# FOR CLUSTER USE
-path = "/share/PI/manishad/multTest"
-setwd(path)
-
-# FOR LOCAL USE
-# path = "~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/Sandbox/2018-1-13"
-# setwd(path)
-
-n = 1000
-nX = 1
-nY = 40
-rho.XX = 0
-rho.YY = c(0.25, 0)
-rho.XY = c(0.05, 0) # null hypothesis: 0
-half = c(0)  # exchangeable vs. half-correlated matrix
-
-# scenarios from OLS
-# rho.YY = c(0.25, 0.5, 0)
-# rho.XY = c(0.02, 0.05, 0) # null hypothesis: 0
-# half = c(0, 1)  # exchangeable vs. half-correlated matrix
-
-# bootstrap iterates and type
-boot.reps = 1000
-bt.type = c( "MICE.H0" )
-
-# matrix of scenario parameters
-scen.params = expand.grid( bt.type, n, nX, nY, rho.XX, rho.YY, rho.XY, half )
-names(scen.params) = c( "bt.type", "n", "nX", "nY", "rho.XX", "rho.YY", "rho.XY", "half" )
-
-
-# name the scenarios
-# remove letters that are privileged variables in R
-letter.names = c(letters, LETTERS)[ ! c(letters, LETTERS) %in% c("i","T","F") ]
-scen.params$scen.name = letter.names[ 1:dim(scen.params)[1] ]
-n.scen = length(scen.params[,1])
-
-# we don't need scenarios with rho.XY = 0 and half = 1 because redundant with 
-# scenarios where rho.XY = 0 and half = 0
-scen.to.toss = scen.params$scen.name[ scen.params$rho.XY == 0 & scen.params$half == 1 ]
-
-scen.params = scen.params[ ! scen.params$scen.name %in% scen.to.toss, ]
-
-# write the csv file of params (to Sherlock)
-write.csv( scen.params, "scen_params.csv" )
-
-
-########################### GENERATE SBATCHES ###########################
-
-# load functions for generating sbatch files
-source("functions.R")
-
-
-# number of sbatches to generate (i.e., iterations within each scenario)
-n.reps.per.scen = 500
-n.reps.in.doParallel = 1
[…]

### 2018-05-18 Manuscript/manuscript.Rmd
Running Freedman resampling

@@
 author: ''
+header-includes:
+- \usepackage[T1]{fontenc}
+- \usepackage{microtype}
+- \usepackage[margin=1in]{geometry}
+- \usepackage{fancyhdr}
+- \usepackage{fancyhdr}
+- \pagestyle{fancy}
+- \fancyhead{}
+- \fancyfoot{}
+- \fancyhead[C]{Statistics for Multisite Replications}
+- \fancyfoot[RO,LE]{\thepage}
+- \usepackage{titlesec}
+- \usepackage{booktabs}
+- \usepackage{lettrine}
+- \usepackage{paralist}
+- \usepackage{setspace}\doublespacing
+- \usepackage{natbib}
+- \setcitestyle{apalike}
+- \usepackage{url}
+- \usepackage{parskip}
+- \usepackage{color,soul}
+- \usepackage{amsmath,amsfonts,amsthm, textcomp}
 output:
-  pdf_document: default
-  word_document: default
+  pdf_document:
+bibliography: refs_mo.bib
+---
+
+<!---
+---
+title: ''
+author: ''
 header-includes:
@@
 - \usepackage{float}
+- \usepackage{natbib}
+- \setcitestyle{apalike}
+output:
+  pdf_document:
+    citation_package: natbib
 bibliography: refs_mo.bib
-csl: apa_style.csl
 ---
+--->
 
@@
 \begin{center}
-\textbf{ \LARGE{Multiple Testing Research Update \#2} } \\ \vspace{5mm}
+\textbf{ \LARGE{Multiple Research Update \#2} } \\ \vspace{5mm}
 \textbf{ \LARGE{\today} }
@@
 <!-------------------------- --------------------------------->
+
+\citet{westfalltext}
 
@@
 with null hypotheses $H_{01}: \beta_1 = 0$ and $H_{02}: \alpha_1 = 0$, and assuming $\epsilon_{nw} \sim N(0, \sigma^2_w)$ for $w \in \{1,2\}$.
-A resampling algorithm \textcolor{red}{[@westfalltext]} satisifying Assumption \ref{cbt_def} could proceed by first fixing the covariates $Z_{n3}$ and $Z_{n4}$ for all observations while setting the "outcomes" equal to a vector of residuals resampled with replacement.
+A resampling algorithm \textcolor{red}{[@freedman]} satisifying Assumption \ref{cbt_def} could proceed by first fixing the covariates $Z_{n3}$ and $Z_{n4}$ for all observations while setting the "outcomes" equal to the fitted values plus a vector of residuals resampled with replacement.
 That is, letting $n'$ denote an observation sampled with replacement and letting hats denote the usual OLS estimates obtained from the original dataset, the resampled variables for observation $n$ are:
 \begin{align*}
-Z_{n1}^{(j)} &:= \left( \widehat{Z}_{n'1} - Z_{n'1} \right)\\
-Z_{n2}^{(j)} &:= \left( \widehat{Z}_{n'2} - Z_{n'2} \right)\\
+Z_{n1}^{(j)} &:= \widehat{Z}_{n1} + \left( \widehat{Z}_{n'1} - Z_{n'1} \right)\\
+Z_{n2}^{(j)} &:= \widehat{Z}_{n2} + \left( \widehat{Z}_{n'2} - Z_{n'2} \right)\\
 Z_{n3}^{(j)} &:= Z_{n3}\\
@@
 \end{align*}
+Then, as usual in bootstrap hypothesis testing [@hall], the test statistics are computed using $H_{01}: \beta_1 = \widehat{\beta}_1$ and $H_{02}: \alpha_1 = \widehat{\alpha}_1$ in order to recover the null distribution.
 This resampling algorithm fulfills Assumption \ref{cbt_def} (Appendix).
@@
 \section*{References}
+
+\bibliography{refs.bib}
+\bibliographystyle{apacite}

### 2018-05-18 R code/For Sherlock/analysis.R

@@ -10,7 +10,7 @@ s = read.csv("stitched.csv", header=TRUE)
 s = merge(s, scen.params, by = "scen" )
 
 # what percent done are we?
-nrow(s) / (500 * nrow(scen.params))
+nrow(s) / (1000 * nrow(scen.params))
 
 # how many reps per scenario do we have?
 table( s$scen )
@@ -28,7 +28,7 @@ library(tidyverse)
 # names of joint rejection variables for all methods
 ( jt.rej.names = c( names(s)[ grep( "jt.rej.", names(s) ) ],
                     names(s)[ grep( "rej.jt", names(s) ) ] ) )
-  
+
 # MAKE SURE TO LIST THEM IN SAME ORDER  AS JT.REJ.NAMES
 method.names = c("bonf.naive",
                  "holm",
@@ -42,11 +42,11 @@ method.names = c("bonf.naive",
 # reshape wide to long
 # https://stackoverflow.com/questions/12466493/reshaping-multiple-sets-of-measurement-columns-wide-format-into-single-columns
 lp = reshape( s[ , names(s) %in% c(jt.rej.names, "scen") ],
-               varying = jt.rej.names,
-               v.names="jt.rej",
-               times = method.names,
+              varying = jt.rej.names,
+              v.names="jt.rej",
+              times = method.names,
               timevar = "method",  # name of the above variable
-               direction="long" )
+              direction="long" )
 
 # sanity check
 # long data should have 1 row per method-scenario combination
@@ -67,16 +67,19 @@ pwr$group = paste( "X-Y correlation: ", pwr$rho.XY,
                    sep = "" )
 
 # for more plotting joy
+labels = c("NB", "H", "MP", "LP", "J1", "J5", "WS", "R")
 pwr$method.label = NA
-pwr$method.label[ pwr$method == "bonf.naive" ] = "NB"
-pwr$method.label[ pwr$method == "holm" ] = "H"
-pwr$method.label[ pwr$method == "minP" ] = "MP"
-pwr$method.label[ pwr$method == "meanP" ] = "LP"  # for "log-P"
-pwr$method.label[ pwr$method == "ours.0.01" ] = "J1"
-pwr$method.label[ pwr$method == "ours.0.05" ] = "J5"
-pwr$method.label[ pwr$method == "Wstep" ] = "WS"
-pwr$method.label[ pwr$method == "Romano" ] = "R"
+pwr$method.label[ pwr$method == "bonf.naive" ] = labels[1]
[…]

### 2018-05-18 R code/For Sherlock/doParallel.R

@@ -7,7 +7,7 @@ scen = args[2]  # this will be a letter
 boot.reps = as.numeric( args[3] )
 
 # get scen parameters
-setwd("/share/PI/manishad/multTest")
+setwd("/home/groups/manishad/multTest")
 scen.params = read.csv( "scen_params.csv" )
 p = scen.params[ scen.params$scen.name == scen, ]
 
@@ -22,17 +22,16 @@ alpha = c( crit.bonf, 0.01, 0.05 )
 sim.reps = 5
 
 # EDITED FOR C++ ISSUE WITH PACKAGE INSTALLATION
-library(doParallel, lib.loc = "/share/PI/manishad/Rpackages/")
-library(foreach, lib.loc = "/share/PI/manishad/Rpackages/")
-library(mvtnorm, lib.loc = "/share/PI/manishad/Rpackages/")
-library(StepwiseTest, lib.loc = "/share/PI/manishad/Rpackages/")  # Romano
-library(matrixcalc, lib.loc = "/share/PI/manishad/Rpackages/")
+library(doParallel, lib.loc = "/home/groups/manishad/Rpackages/")
+library(foreach, lib.loc = "/home/groups/manishad/Rpackages/")
+library(mvtnorm, lib.loc = "/home/groups/manishad/Rpackages/")
+library(StepwiseTest, lib.loc = "/home/groups/manishad/Rpackages/")  # Romano
+library(matrixcalc, lib.loc = "/home/groups/manishad/Rpackages/")
+
+# for use in ml load R
+# install.packages( c("doParallel", "foreach", "mvtnorm", "StepwiseTest", "matrixcalc"), lib = "/home/groups/manishad/Rpackages/" )
 
 source("functions.R")
-
-# names of methods that resample under H0 vs. HA
-h0.methods = c("Y", "resid", "h0.parametric")
-ha.methods = c("fcr", "ha.resid", "ha.resid.2")
 
 # set the number of cores
 registerDoParallel(cores=16)
@@ -43,17 +42,17 @@ registerDoParallel(cores=16)
 # # setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/Sandbox/2018-1-13")
 # # p = read.csv("scen_params.csv")  # should be a single row, I think
 # 
-# n = 100
+# n = 1000
 # nX = 1
 # nY = 40
 # rho.XX = 0
[…]

### 2018-05-18 R code/For Sherlock/functions.R

@@ -301,8 +301,13 @@ sim_data = function( .n, .cor ) {
 # fit regression model for a single outcome Y regressed on all Xs
 # return p-value for the exposure of interest (X1)
 
-fit_model = function( Y.name, .dat, .resid, .sigma, .tval, .intercept ) {
+# p: parameters matrix
 
+fit_model = function( Y.name,
+                      .dat,
+                      .center.stats,
+                      .bhat.orig ) {
+  
   # extract names
   X.names = names( .dat )[ grep( "X", names( .dat ) ) ]
   
@@ -314,29 +319,35 @@ fit_model = function( Y.name, .dat, .resid, .sigma, .tval, .intercept ) {
   RHS = paste(X.names, collapse=" + ")
   formula = paste( c( Y.name, RHS ), collapse = " ~ ")
   
+  
   # extract p-value for exposure of interest (X1)
   m = lm( eval( parse(text=formula) ), data=.dat )
-  pval = summary(m)$coefficients[ "X1", 4 ]
+
+  resid.return = residuals(m)
+  sigma.return = summary(m)$sigma
+  intercept.return = coef(m)[["(Intercept)"]]
+  se.return = summary(m)$coefficients["X1", "Std. Error"]
   
-  if ( ! .resid ) resid.return = NA
-  else resid.return = residuals(m)
-    
-  if ( ! .sigma ) sigma.return = NA
-  else sigma.return = summary(m)$sigma
+  # if resampling was done under Ha, center stats by the original-sample estimates
+  if( !.center.stats ) {
+    bhat.return = coef(m)[["X1"]]
+    tval.return = summary(m)$coefficients[ "X1", 3 ]
+  }
+  if( .center.stats ) {
+    # second term pulls out the correct beta-hat from the original vector by using the Yname that we're regressing on
+    bhat.return = coef(m)[["X1"]] - .bhat.orig[ as.numeric( substr( Y.name, 2, nchar(Y.name) ) ) ]
+    tval.return = bhat.return / se.return
+  }
   
-  if ( ! .intercept ) intercept.return = NA
-  else intercept.return = coef(m)[["(Intercept)"]]
+  df = nrow(.dat) - length(X.names) - 1
+  pval.return = 2 * ( 1 - pt( abs(tval.return), df = df ) )
   
-  # extract test stat for this test
-  # for use with Romano's code
-  if ( ! .tval ) tval.return = NA
-  else tval.return = summary(m)$coefficients[ "X1", 3 ]
-  
[…]

### 2018-05-18 R code/For Sherlock/gen_sbatch.R

@@ -3,7 +3,7 @@
 ########################### SET SIMULATION PARAMETERS MATRIX ###########################
 
 # FOR CLUSTER USE
-path = "/share/PI/manishad/multTest"
+path = "/home/groups/manishad/multTest"
 setwd(path)
 
 # FOR LOCAL USE
@@ -20,7 +20,7 @@ prop.corr = c(0.5, 1)  # exchangeable vs. half-correlated matrix
 
 # bootstrap iterates and type
 boot.reps = 2000
-bt.type = c( "resid", "ha.resid" )
+bt.type = c( "ha.resid" )
 
 
 # matrix of scenario parameters
@@ -33,8 +33,8 @@ names(scen.params) = c( "bt.type", "n", "nX", "nY", "rho.XX", "rho.YY", "rho.XY"
 ######## Remove Some Scenarios ########
 # for rhoXY = 0.10, only want info for confidence intervals
 # so cut out the ha.resid ones
-scen.params = scen.params[ scen.params$rho.XY != 0.10 |
-                             scen.params$rho.XY == 0.10 & scen.params$bt.type == "resid", ]
+# scen.params = scen.params[ scen.params$rho.XY != 0.10 |
+#                              scen.params$rho.XY == 0.10 & scen.params$bt.type == "resid", ]
 
 # for rho.XY = 0, don't need to manipulate prop.corr because all 0 anyway
 scen.params = scen.params[ scen.params$rho.XY != 0 |
@@ -43,7 +43,7 @@ scen.params = scen.params[ scen.params$rho.XY != 0 |
 
 ######## Name the Scenarios ########
 # if merging results with other simulations, set this to the last letter already used
-start.at = "a"
+#start.at = "a"
 
 # remove letters that are privileged variables in R
 # letter.names = c(letters, LETTERS, paste(letters, letters, sep="") )
@@ -71,7 +71,7 @@ n.reps.in.doParallel = 5
 n.files = ( n.reps.per.scen / n.reps.in.doParallel ) * n.scen
 
 
-path = "/share/PI/manishad/multTest"
+path = "/home/groups/manishad/multTest"
 
 scen.name = rep( scen.params$scen.name, each = ( n.files / n.scen ) )
 jobname = paste("job", 1:n.files, sep="_")
@@ -100,29 +100,30 @@ sbatch_params <- data.frame(jobname,
 
 #generateSbatch(sbatch_params, runfile_path)
 
[…]

### 2018-05-18 R code/For Sherlock/stitch_on_sherlock.R

@@ -20,7 +20,7 @@ write.csv( s, "stitched.csv" )
 table(s$scen)
 
 # move it to Desktop
-scp mmathur@sherlock:/share/PI/manishad/multTest/sim_results/stitched.csv ~/Desktop
-scp mmathur@sherlock:/share/PI/manishad/multTest/scen_params.csv ~/Desktop
+scp mmathur@login.sherlock.stanford.edu:/home/groups/manishad/multTest/sim_results/overall_stitched/stitched.csv ~/Desktop
+scp mmathur@login.sherlock.stanford.edu:/home/groups/manishad/multTest/scen_params.csv ~/Desktop
 
 

### 2018-05-18 R code/For Sherlock/analysis.R
Update analysis code for Freedman resampling

@@ -139,62 +139,6 @@ ggplot( data = pwr, aes( x = rho.YY, y = power,
   ggtitle("Power of bootstrapped hypothesis test of joint null")
 
 
-# PREVIOUS ATTEMPT TO HAVE LABELS IN GRAPH
-
-# legend.labs = c("Naive Bonferroni",
-#                 "Holm",
-#                 "Westfall minP",
-#                 "Mean log-P",
-#                 "Ours (alpha = 0.01)",
-#                 "Ours (alpha = 0.05)",
-#                 "Romano",
-#                 "Westfall step-down")
-
-
-# # hackily use acronyms in the legend
-# # https://stackoverflow.com/questions/49965758/change-geom-texts-default-a-legend-to-label-string-itself/49966057?noredirect=1#comment86951861_49966057
-# oldK = GeomText$draw_key # to save for later
-# 
-# # define new key
-# # if you manually add colours then add vector of colours 
-# # instead of `scales::hue_pal()(length(var))`
-# GeomText$draw_key <- function (data, params, size, 
-#                                var = unique(pwr$method.label), 
-#                                longvar = unique(pwr$method), 
-#                                cols=scales::hue_pal()(length(var))) {
-#   
-#   #browser()
-#   
-#   # sort as ggplot sorts these alphanumerically / or levels of factor
-#   txt <- if(is.factor(var)) levels(var) else sort(var)
-#   txt <- txt[match(data$colour, cols)]
-#   
-#   textGrob(txt, 0.5, 0.5,  
-#            just="center", 
-#            gp = gpar(col = alpha(data$colour, data$alpha), 
-#                      fontfamily = data$family, 
-#                      fontface = data$fontface, 
-#                      fontsize = data$size * .pt))
-# }
-# 
-# 
-# ggplot( data = pwr, aes( x = rho.YY, y = power,
-#                          color = method,
-#                          label = method ) ) +
-#   geom_text( aes( label = method.label) ) +
-#   theme_bw() +
-#   facet_wrap( ~group) +
-#   #facet_wrap(~ group, nrow = 2 ) +  # for changing rows/columns
-#   ylab("Power") +
-#   scale_x_continuous( limits = c( min(x.breaks), max(x.breaks) ), breaks = x.breaks ) +
-#   scale_y_continuous( limits = c( min(y.breaks), max(y.breaks) ), breaks = y.breaks ) +
[…]

### 2018-05-25 Project log/multiple_outcomes_project_log.Rmd
.

@@
 <!-- NEW SECTION -->
-\section{Proof sketch following Freedman paper}
 
-\begin{enumerate}
-\item{ Define multivariate Mallows metric, $\tilde{d}_2$. Prove that it's still a metric and that Eq. (2.3) still holds. Since it's a metric, triangle inequality (2.6) still holds:
-$$\frac{1}{2} \tilde{d}_2(F_n, F)^2 \le \tilde{d}_2(\widehat{F}_n, F_n) + \tilde{d}_2( F_n, F)$$}
+\section{Back to Romano vs. Westfall}
 
-\item{ $\bigstar$ Show that Theorem 2.1 still holds for $\tilde{d}_2$, assuming the same design matrix for all regressions:
-$$\tilde{d}_2^p( \Psi (F_n), \Psi (F)) \le n \cdot \text{trace} [X(n)'X(n)]^{-1} \cdot d_2(F,G)$$
-}
+Romano’s crit values: 95th percentile of biggest CENTERED test stat (N(0,1)) among the subset of hypotheses that ‘s being considered, where the original test stats are computed not under null
+(pg 1384)
 
-\item{Work on each piece of the triangle inequality:
-	\begin{enumerate}
-    	\item $\tilde{d}_2(F_n, F) \to 0$ because we can still apply SLLN to joint distribution (analog to B\&F Lemmas 8.3/8.4).
-        \item $\tilde{d}_2( \widehat{F}_n, F_n) \le \frac{p}{n} \sum_{w=1}^W \sigma^2_w \to 0$ (analog to Lemmas 2.1/2.2).
-    \end{enumerate}
-}
-
-\item{Lemma 2.4 still holds because each term $\to 0$ by B\&F Lemma 2.4:
-$$\frac{1}{n} \sum_{w=1}^W \| \widehat{\epsilon}_{w} (n) - \widehat{\epsilon}_{w} (n) \|^2 \to 0$$
-}
-
-\item{ Lemma 2.5 still holds using our Eq. (2.3), our Lemma 2.4, and the fact that $\widehat{F}_n = \tilde{F}_n$ for us:
-
-$$d_2( \widehat{F}_n, F_n ) \to 0$$}
-
-\item{$\bigstar$ Show that $\Psi (F_n)$ is close to $\Psi (\widehat{F}_n)$ by writing difference between vector of coefficient estimates in terms of residuals and resampled residuals (loose analog to Theorem 2.2).}
-\end{enumerate}
+Westfall's crit values: adjust p-values using the p-values under the strong null
 
@@
 \section{Resolved questions}
+
+\subsection{Why is it OK for us to resample using Romano's method when applying Westfall?}
+
+Because subset pivotality holds for OLS.
+Think about the example below regarding pairwise correlations where subset pivotality \emph{doesn't} hold.
+In that case, the resampling methods would differ because with Westfall, the correlation between two of the test stats in the resamples is reflective of the third test stat being zero.
+With Romano, the correlation between those test stats is reflective of the potentially \emph{non-zero} estimate of the third test stat.
+
+But with OLS, because of subset pivotality, the joint distribution of any group of test stats under the strong null has nothing to do with the truth or falsehood of the others, so whether we set all the other test stats to zero during data generation (Westfall) or after, by centering the test stats (Romano), doesn't matter.
+
+\subsection{What's up with subset pivotality and how does Romano manage to avoid it?}
+
+From the examples in Romano \& Wolf JASA paper (2003), subset pivotality fails when, e.g., the SE or correlation structure of some of the test stats depends on parameters used in \emph{other} tests.
+For example (Romano 2003 JASA, Example 4.1), say you're testing all pairs of correlations for three variables.
+X and Y are truly uncorrelated, and X and Z are uncorrelated (nulls are true).
+But Y and Z are correlated.
+Then the joint distribution of the X-Y and X-Z correlations depends on the Y-Z correlation, which makes sense intuitively.
+So subset pivotality fails.
+In this case, Westfall's approach would be to resample under the strong null, so all three correlations are null and the X-Y and X-Z correlations themselves are uncorrelated.
+But this does not reflect all possible distributions under which the X-Y and X-Z correlations are null since their distribution depends on the Y-Z correlation.
+Romano's approach would be to resample under the original distribution, then center the test stats.
+So then we would have correlations that are all centered at 0, but the X-Y and X-Z correlations are correlated if, in reality, the Y-Z correlation is non-null.
+
+But I think this also means that if you have test stats whose distribution depends on the parameter being tested (not on the other parameters, so not talking about subset pivotality), like $\widehat{p}$ for a Bernoulli RV (whose SE depends on $p$), then Westfall's method would have the right SEs (reflecting the null), but Romano's would not.
+However, I think this is actually \emph{not} a problem.
+As with a single hypothesis test, the "validity" of a hypothesis test depends only its behavior (false positive rate) for samples \emph{generated under the null} -- and when the null does hold, $\widehat{SE} \to SE$ anyway, so asymptotically, it does not matter if you estimate SEs from the data or plug in SEs under the null.
+Analogously, in the multiple testing setting, the test needs to control FWER only when the null holds for the subset of hypotheses we are considering.
+This is different from subset pivotality, which concerns dependence of the set of nulls under consideration on the \emph{other} hypotheses from outside the set.
 

### 2018-05-25 R code/For Sherlock/analysis.R

@@ -15,6 +15,9 @@ nrow(s) / (1000 * nrow(scen.params))
 # how many reps per scenario do we have?
 table( s$scen )
 
+# because of weird row with scen=FALSE
+library(gdata)
+s = drop.levels(s)
 
 
 ########################### DATA PREP ###########################
@@ -24,6 +27,9 @@ table( s$scen )
 
 library(reshape2)
 library(tidyverse)
+library(dplyr)
+library(magrittr)
+library(tidyr)
 
 # names of joint rejection variables for all methods
 ( jt.rej.names = c( names(s)[ grep( "jt.rej.", names(s) ) ],
@@ -50,13 +56,20 @@ lp = reshape( s[ , names(s) %in% c(jt.rej.names, "scen") ],
 
 # sanity check
 # long data should have 1 row per method-scenario combination
-# nrow( lp[ lp$scen == "a", ] ); nrow( s[ s$scen == "a", ]) * length(method.names)
+# nrow( lp[ lp$scen == 1, ] ); nrow( s[ s$scen == 1, ]) * length(method.names)
 
+
+# fix factor levels because of scenario that was TRUE/FALSE as strings
+fake = NA
+fake[lp$jt.rej == "FALSE"] = FALSE
+fake[lp$jt.rej == "TRUE"] = TRUE
+lp$jt.rej = fake
 
 # since each scenario still has multiple rows, take means 
 pwr = lp %>% group_by(scen, method) %>%
   summarise( power = mean(jt.rej) )
 
+# merge in scenario parameters
 pwr = merge(pwr, scen.params, by = "scen" )
 
 # for plotting joy
@@ -81,19 +94,6 @@ pwr$method.label[ pwr$method == "Romano" ] = labels[8]
 # sanity check
 # table(pwr$method, pwr$method.label)
 
-# # sanity check
-# # should have NAs for method-bt.type combinations that are impossible
-# pwr %>% filter( method %in% c("Romano") ) %>%
-#   filter( bt.type == "h0.resid" )
-# 
-# # should have NA power
-# pwr %>% filter( method %in% c("minP", 
-#                               "meanP",
-#                               "ours.0.01", 
-#                               "ours.0.05", 
-#                               "Wstep") ) %>%
-#   filter( bt.type == "ha.resid")
-
 
 ##### Save Results ##### 
 pwr2 = pwr
@@ -113,9 +113,12 @@ lp2 = lp
[…]

### 2018-05-25 R code/For Sherlock/functions.R

@@ -466,8 +466,8 @@ sbatch_not_run = function(.results.singles.path,
                           .results.write.path,
                         .name.prefix) {
   
-  # .results.singles.path = "/share/PI/manishad/multTest/sim_results/short"
-  # .results.write.path = "/share/PI/manishad/multTest/sim_results"
+  # .results.singles.path = "/home/groups/manishad/multTest/sim_results/short"
+  # .results.write.path = "/home/groups/manishad/multTest/sim_results"
   # .name.prefix = "short"
   
   # get list of all files in folder
@@ -488,7 +488,7 @@ sbatch_not_run = function(.results.singles.path,
   missed.nums = all.nums[ !all.nums %in% sbatch.nums ]
   
   # give info
-  print("The max job number is: ", max(sbatch.nums) )
+  print( paste("The max job number is: ", max(sbatch.nums) ) )
   print( paste( "Number of jobs that weren't run: ",
         ifelse( length(missed.nums) > 0, length(missed.nums), "none" ) ) )
   

### 2018-05-25 R code/For Sherlock/gen_sbatch.R

@@ -100,30 +100,30 @@ sbatch_params <- data.frame(jobname,
 
 #generateSbatch(sbatch_params, runfile_path)
 
-#5400 files
+#6600 files
 
-setwd( paste(path, "/sbatch_files", sep="") )
-for (i in 1:1000) {
-  system( paste("sbatch -p normal,owners /home/groups/manishad/multTest/sbatch_files/", i, ".sbatch", sep="") )
-}
-
-
-# ######## If Running Only Some Jobs To Fill Gaps ######## 
-# 
-# # run in Sherlock ml load R
-# # sbatch_not_run( "$PI_HOME/multTest/sim_results/short", 
-# #                 "$PI_HOME/multTest/sim_results",
-# #                 .name.prefix = "short" )
-# # scp mmathur@sherlock:$PI_HOME/multTest/sim_results/missed_job_nums.csv ~/Desktop
-# 
-# setwd("$PI_HOME/multTest/sim_results")
-# missed.nums = read.csv("missed_job_nums.csv")$x
-# 
-# 
 # setwd( paste(path, "/sbatch_files", sep="") )
-# for (i in missed.nums) {
+# for (i in 6583:6600) {
 #   system( paste("sbatch -p normal,owners /home/groups/manishad/multTest/sbatch_files/", i, ".sbatch", sep="") )
 # }
+
+
+######## If Running Only Some Jobs To Fill Gaps ######## 
+
+# run in Sherlock ml load R
+sbatch_not_run( "/home/groups/manishad/sim_results/short", 
+                "/home/groups/manishad/sim_results",
+                .name.prefix = "short" )
+scp mmathur@sherlock:$PI_HOME/multTest/sim_results/missed_job_nums.csv ~/Desktop
+
+setwd("$PI_HOME/multTest/sim_results")
+missed.nums = read.csv("missed_job_nums.csv")$x
+
+
+setwd( paste(path, "/sbatch_files", sep="") )
+for (i in missed.nums) {
+  system( paste("sbatch -p normal,owners /home/groups/manishad/multTest/sbatch_files/", i, ".sbatch", sep="") )
+}
 
 
 

### 2018-05-25 R code/For Sherlock/stitch_on_sherlock.R

@@ -1,26 +1,25 @@
 
 # run this interactively in ml load R
 
-path = "/share/PI/manishad/multTest"
+path = "/home/groups/manishad/multTest"
 setwd(path)
 source("functions.R")
 
 ######## FOR STITCHING LONG FILES ########
 
-s = stitch_files(.results.singles.path = "/share/PI/manishad/multTest/sim_results/short",
-                 .results.stitched.write.path = "/share/PI/manishad/multTest/sim_results",
+s = stitch_files(.results.singles.path = "/home/groups/manishad/multTest/sim_results/short",
+                 .results.stitched.write.path = "/home/groups/manishad/multTest/sim_results/overall_stitched",
                  .name.prefix = "short",
                  .stitch.file.name="stitched.csv")
 
-# write results
-write.csv( s, "stitched.csv" )
-
 # how close are we to being done?
-(nrow(s)/5) / 12000
+n.reps.per.scen = 1000
+nrow(s) / (n.reps.per.scen * nrow(scen.params))
 table(s$scen)
 
 # move it to Desktop
 scp mmathur@login.sherlock.stanford.edu:/home/groups/manishad/multTest/sim_results/overall_stitched/stitched.csv ~/Desktop
+Vegemite2017
 scp mmathur@login.sherlock.stanford.edu:/home/groups/manishad/multTest/scen_params.csv ~/Desktop
-
+Vegemite2017
 

### 2018-06-07 Project log/multiple_outcomes_project_log.Rmd
Fix functions.R bug in cell_corr that was making results wrong for prop.corr != 1

--- 2018-05-13 Project log/multiple_outcomes_project_log.Rmd
@@
 <!-- NEW SECTION -->
-\section{Proof sketch following Freedman paper}
-
-\begin{enumerate}
-\item{ Define multivariate Mallows metric, $\tilde{d}_2$. Prove that it's still a metric and that Eq. (2.3) still holds. Since it's a metric, triangle inequality (2.6) still holds:
-$$\frac{1}{2} \tilde{d}_2(F_n, F)^2 \le \tilde{d}_2(\widehat{F}_n, F_n) + \tilde{d}_2( F_n, F)$$}
-
-\item{ $\bigstar$ Show that Theorem 2.1 still holds for $\tilde{d}_2$, assuming the same design matrix for all regressions:
-$$\tilde{d}_2^p( \Psi (F_n), \Psi (F)) \le n \cdot \text{trace} [X(n)'X(n)]^{-1} \cdot d_2(F,G)$$
-}
-
-\item{Work on each piece of the triangle inequality:
-	\begin{enumerate}
-    	\item $\tilde{d}_2(F_n, F) \to 0$ because we can still apply SLLN to joint distribution (analog to B\&F Lemmas 8.3/8.4).
-        \item $\tilde{d}_2( \widehat{F}_n, F_n) \le \frac{p}{n} \sum_{w=1}^W \sigma^2_w \to 0$ (analog to Lemmas 2.1/2.2).
-    \end{enumerate}
-}
-
-\item{Lemma 2.4 still holds because each term $\to 0$ by B\&F Lemma 2.4:
-$$\frac{1}{n} \sum_{w=1}^W \| \widehat{\epsilon}_{w} (n) - \widehat{\epsilon}_{w} (n) \|^2 \to 0$$
-}
-
-\item{ Lemma 2.5 still holds using our Eq. (2.3), our Lemma 2.4, and the fact that $\widehat{F}_n = \tilde{F}_n$ for us:
-
-$$d_2( \widehat{F}_n, F_n ) \to 0$$}
-
-\item{$\bigstar$ Show that $\Psi (F_n)$ is close to $\Psi (\widehat{F}_n)$ by writing difference between vector of coefficient estimates in terms of residuals and resampled residuals (loose analog to Theorem 2.2).}
-\end{enumerate}
 
@@
 
-MICE thing doesn't work because we're regressing $Y_2$ on $Y_1$, and if data are generated under $H_A$, these are correlated because both outcomes are correlated with $X$.
-This remains true even if we leave $X$ out of the imputation model.
+Chatterjee generalized bootstrap paper: This seems to say that for logistic regression \emph{confidence interval} (not under $H_0$), we can sample with replacement from observations' contributions to the score equations, and then solve these new score equations.
+I don't think it's valid to simply center these test stats as for OLS because, as shown in the FCR-center simulations, the SEs will be wrong.
+But maybe we could plug in the null values of the parameters into the score equations?
 
-FCR-center works for OLS but not logistic regression because for the latter, SE depends on value of parameter. (Is this subset pivotality?)
+R algorithm:
+\begin{enumerate}
+\item Write function that calculates score contribution as function of parameters.
+\item Setting the parameter of interest to null value, calculate score contribution for each observation. 
+\item Find the parameter that solves the new score equations?
+\end{enumerate}
 
-Marginally vs. conditionally: See simulation writeups.
+Actually, I think we have a broader problem: How do we do bootstrap hypothesis tests with GLMs at all?
 
-Bootstrapping brainteaser: Observations really are independent in FCR resamples. I think the previous issues with this were actually not due to failed assumptions. 
-\url{https://stats.stackexchange.com/questions/339237/are-observations-independent-in-bootstrapped-resamples}
+Does the Chatterjee paper mean that you can always use FCR to get CIs for any GLM?
 
@@
 
+\subsection{Why doesn't MICE idea work?}
+
+MICE thing doesn't work because we're regressing $Y_2$ on $Y_1$, and if data are generated under $H_A$, these are correlated because both outcomes are correlated with $X$.
+This remains true even if we leave $X$ out of the imputation model.
+
 \subsection{What does Westfall do for logistic regression?}
@@
 \section{Simulation notes}
+
+\subsection{2018-6-6 *more scenarios}
+
+Trying more scenarios to have smaller proportion of effects but still with observed rejections way outside CI.
+
+\begin{itemize}
+\item $\rho_{XY} = 0.25$ with 20\% correlated: Has 100\% power, so need to go down. 
+\item $\rho_{XY} = 0.25$ with 5\% correlated: Same as above
+\item $\rho_{XY} = 0.15$ with 5\% correlated: Same as above
+\end{itemize}
+
+\subsection{2018-5-18 *all scenarios Freedman resampling}
+
+These are in the paper.
 

### 2018-06-07 R code/For Sherlock/analysis.R

@@ -97,11 +97,11 @@ pwr$method.label[ pwr$method == "Romano" ] = labels[8]
 
 ##### Save Results ##### 
 pwr2 = pwr
-setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/Simulation results/2019-5-18 all scenarios Freedman resampling")
-write.csv( pwr2, "results_pwr.csv")
+#setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/Simulation results/2019-5-18 all scenarios Freedman resampling")
+#write.csv( pwr2, "results_pwr.csv")
 
 lp2 = lp
-write.csv( lp2, "results_pwr_long.csv")
+#write.csv( lp2, "results_pwr_long.csv")
 
 lp2 = lp
 
@@ -132,7 +132,7 @@ p = ggplot( data = pwr, aes( x = rho.YY, y = power,
                          color = method,
                          label = method ) ) +
   geom_text( aes( label = method.label) ) +
-  theme_bw() +
+  theme_classic() +
   facet_wrap( ~group) +
   #facet_wrap(~ group, nrow = 2 ) +  # for changing rows/columns
   ylab("Power") +
@@ -144,8 +144,8 @@ p = ggplot( data = pwr, aes( x = rho.YY, y = power,
   theme(legend.position="none") # remove legend
 
 name = "joint_test_full.png"
-ggsave( filename = paste(name),
-        plot=p, path=NULL, width=10, height=8, units="in")
+#ggsave( filename = paste(name),
+#        plot=p, path=NULL, width=10, height=8, units="in")
 
 
 
@@ -216,7 +216,7 @@ legend.labs = c( "J1 (alpha = 0.01)", "J5 (alpha = 0.05)" )
 ci2.short = ci2[ ci2$rho.XY %in% c(0, 0.05, 0.1), ]
 
 # for Appendix version, run this with ci2 instead of ci2.short
-p = ggplot( data = ci2.short ) +
+p = ggplot( data = ci2 ) +
   # bootstrap results
   geom_point( aes( x = rho.YY, y = n.rej.bt.mn, color = method ), size=2.5 ) +
   #geom_line( aes( x = rho.YY, y = n.rej.bt.mn, color = method ), size=1.1 ) +
@@ -228,7 +228,7 @@ p = ggplot( data = ci2.short ) +
   scale_shape_manual( values = c('the shape' = 4),
[…]

### 2018-06-07 R code/For Sherlock/functions.R

@@ -35,23 +35,6 @@ adjust_minP = function( p, p.bt ) {
 # Westfall textbook, pages 66-67
 adj_Wstep = function( p, p.bt ) {
   
-  # # TEST ONLY
-  # p = structure(c(0.636112475277457, 0.116408752390887, 0.260603159221029,
-  #                 0.575086408582202, 0.625789849782125), .Dim = c(5L, 1L))
-  # p.bt = structure(c(0.728140113074014, 0.828557135138776, 0.728464927503583,
-  #                    0.376518353153738, 0.241363541131996, 0.415330227654616, 0.75686321503498,
-  #                    0.865494617532842, 0.839819273240726, 0.540707408457488, 0.149482739547785,
-  #                    0.506274953008223, 0.122605643925883, 0.747018941160177, 0.168076477020114,
-  #                    0.0582566955778811, 0.284459435324328, 0.561050485854955, 0.13839444320949,
-  #                    0.641335055023296, 0.195313599009888, 0.174692123827015, 0.982035588216878,
-  #                    0.713525352129964, 0.548770486028363), .Dim = c(5L, 5L), .Dimnames = list(
-  #                      NULL, c("result.1", "result.2", "result.3", "result.4", "result.5"
-  #                      )))
-  
-  # if using read-in or simulated dat
-  # p = pvals
-  
-  
   # attach indices to original p-values
   # to keep track of their original order
   p.dat = data.frame( ind = 1:length(p), p )
@@ -94,8 +77,25 @@ adj_Wstep = function( p, p.bt ) {
   #plot( p.dat$p, p.dat$p.adj )
 }
 
+
+# sanity check
+# # TEST ONLY
+# p = structure(c(0.636112475277457, 0.116408752390887, 0.260603159221029,
+#                 0.575086408582202, 0.625789849782125), .Dim = c(5L, 1L))
+# p.bt = structure(c(0.728140113074014, 0.828557135138776, 0.728464927503583,
+#                    0.376518353153738, 0.241363541131996, 0.415330227654616, 0.75686321503498,
+#                    0.865494617532842, 0.839819273240726, 0.540707408457488, 0.149482739547785,
+#                    0.506274953008223, 0.122605643925883, 0.747018941160177, 0.168076477020114,
[…]

### 2018-06-07 R code/For Sherlock/gen_sbatch.R

@@ -14,14 +14,14 @@ n = 1000
 nX = 1
 nY = 40
 rho.XX = 0
-rho.YY = c(0.25, 0.5, 0)
-rho.XY = c(0, 0.02, 0.03, 0.04, 0.05, 0.10) # null hypothesis: 0
-prop.corr = c(0.5, 1)  # exchangeable vs. half-correlated matrix
+rho.YY = c(0, 0.1, 0.3, 0.6)
+rho.XY = c(0, 0.03, 0.05, 0.10, 0.15 ) # null hypothesis: 0
+prop.corr = c(0.20, 0.5, 1)  # exchangeable vs. half-correlated matrix
+
 
 # bootstrap iterates and type
 boot.reps = 2000
 bt.type = c( "ha.resid" )
-
 
 # matrix of scenario parameters
 scen.params = expand.grid( bt.type, n, nX, nY, rho.XX, rho.YY, rho.XY, prop.corr )
@@ -43,7 +43,7 @@ scen.params = scen.params[ scen.params$rho.XY != 0 |
 
 ######## Name the Scenarios ########
 # if merging results with other simulations, set this to the last letter already used
-#start.at = "a"
+start.at = 1
 
 # remove letters that are privileged variables in R
 # letter.names = c(letters, LETTERS, paste(letters, letters, sep="") )
@@ -52,7 +52,7 @@ scen.params = scen.params[ scen.params$rho.XY != 0 |
 # 
 # scen.params$scen.name = letter.names[ 1:dim(scen.params)[1] ]
 
-scen.params$scen.name = 1:nrow(scen.params)
+scen.params$scen.name = start.at : ( start.at + nrow(scen.params) - 1 )
 n.scen = length(scen.params[,1])
 
 # write the csv file of params (to Sherlock)
@@ -100,30 +100,33 @@ sbatch_params <- data.frame(jobname,
 
 #generateSbatch(sbatch_params, runfile_path)
 
-#6600 files
+n.files
 
-# setwd( paste(path, "/sbatch_files", sep="") )
-# for (i in 6583:6600) {
-#   system( paste("sbatch -p normal,owners /home/groups/manishad/multTest/sbatch_files/", i, ".sbatch", sep="") )
-# }
+# sent first 1,000 at 1:30pm on Tues
 
-
-######## If Running Only Some Jobs To Fill Gaps ######## 
-
-# run in Sherlock ml load R
-sbatch_not_run( "/home/groups/manishad/sim_results/short", 
[…]

### 2018-06-07 R code/For Sherlock/stitch_on_sherlock.R

@@ -17,6 +17,9 @@ n.reps.per.scen = 1000
 nrow(s) / (n.reps.per.scen * nrow(scen.params))
 table(s$scen)
 
+# quick check
+mean(s$jt.rej.holm)
+
 # move it to Desktop
 scp mmathur@login.sherlock.stanford.edu:/home/groups/manishad/multTest/sim_results/overall_stitched/stitched.csv ~/Desktop
 Vegemite2017

### 2018-06-07 R code/For Sherlock/within_prep.R

@@ -1,113 +1,0 @@
-
-# when outputting all bootstrap iterates, this script aggregates each simulation rep's results
-
-# get scenario parameters
-path = "/share/PI/manishad/multTest"
-setwd(path)
-scen.params = read.csv("scen_params.csv")
-
-
-# get parameters (ASSUMES ONLY 1 SCENARIO)
-alpha = scen.params$alpha
-nY = scen.params$nY
-N0 = alpha * nY
-
-# get individual simulation results
-path = "/share/PI/manishad/multTest/sim_results"
-setwd(path)
-
-# get list of all files in folder
-all.files = list.files(path, full.names=TRUE)
-
-# we only want the ones whose name includes "results"
-keepers = all.files[ grep( "results_job", all.files ) ]
-n.files = length(keepers)
-
-print(n.files)
-print(head(keepers))
-
-library(doParallel)
-library(foreach)
-
-# set the number of cores
-registerDoParallel(cores=16)
-
-parallel.time = system.time( {
-  s = foreach( i = 1:n.files, .combine=rbind ) %dopar% {
-    
-    cat("\n\nStarting file number", i, ", called", keepers[i])
-    
-    d = read.csv( keepers[i] )
-  
-    # simulation rep
-    n.boot = max(d$bt.iterate)
-    n.sims.run = dim(d)[1] / n.boot
-    d$sim.rep = rep( seq(1, n.sims.run), each = n.boot )
-    
-    # convert time in seconds to time in hours
-    d$boot.hrs = d$boot.time / (60*60)
-    
-    # job file
-    d$job.file = keepers[i]
-    
-    ##### Statistical Performance ######
-    # compute CI coverage
-    library(data.table)
-    dt = data.table(d)
-#      dt[ , lo.bt := quantile( n.rej.bt, alpha / 2 ), by=sim.rep ]
-#      dt[ , hi.bt := quantile( n.rej.bt, 1 - (alpha / 2) ), by=sim.rep ]
-# 
-#     # coverage
-#     dt$covers = (dt$lo.bt <= N0 ) & (dt$hi.bt >= N0)
-     
-    # hypothesis test
-    # Hall & Wilson style
-    #dt[ , crit.val := quantile( abs( n.rej.bt - n.rej ), 1 - alpha ), by=sim.rep ]
-    #dt[ , bt.rej := abs( n.rej - N0 ) > crit.val, by = sim.rep ]
-    
-    # one-sided hypothesis test
-    # basic percentile method
[…]

### 2018-06-07 R code/For Sherlock/functions.R
Document & audit cell_corr

@@ -101,14 +101,7 @@ get_crit = function( p.dat, col.p ) {
   
   # sort bootstrapped p-values according to original ones
   col.p.sort = col.p[ p.dat$ind ]
-  
-  # # OLD VERSION (wrong)
-  # qstar = rep( NA, length(col.p.sort) )
-  # for ( i in 1:length(col.p.sort) ) {
-  #   if (i == 1) qstar[i] = col.p.sort[1]
-  #   else qstar[i] = min( qstar[i-1], col.p.sort[i] )
-  # }
-  
+
   qstar = rep( NA, length(col.p.sort) )
 
   k = length(col.p.sort)
@@ -125,16 +118,21 @@ get_crit = function( p.dat, col.p ) {
 
 ########################### FN: RETURN CORRELATION BETWEEN TWO ARBITRARY CELLS ###########################
 
-# eventually could make this an argument in the scen_params matrix somehow so that it can be varied
-# assume that X1 is the covariate of interest
-# assume that none of the other covariates are associated with any outcomes
+# Returns correlation between a pair of variables given
+#  desired correlation structure.
 
-# .half: Should only the last half the outcomes have the specified correlation?
-#  (if FALSE, correlation matrix is exchangeable; otherwise first half are 0 and 
-#  second half are rho.XY)
-# This is only included for backward compatibility because now we have the more flexible .prop.corr thing
+# Notes: 
+# Assumes X1 is the covariate of interest and none of the other covariates
+#  is associated with any outcomes.
 
-# .prop.corr: What proportion of Ys should be correlated with X?
+# Arguments:
+# vname.1: Quoted name of first variable
+# vname.2: Quoted name of second variable
+# .rho.XX: Correlation between pairs of Xs
+# .rho.XY: Correlation between pairs of X-Y (of non-null ones)
+# .nY: Number of Ys
+# .prop.corr: Proportion of X-Y pairs that are non-null
+#  (non-nulls will be first .prop.corr * .nY pairs)
 
 cell_corr = function( vname.1,
                       vname.2,
@@ -142,7 +140,6 @@ cell_corr = function( vname.1,
                       .rho.YY,
                       .rho.XY,
                       .nY,
-                      #.half = 0,
                       .prop.corr = 1) {
   
[…]

### 2018-06-07 R code/For Sherlock/functions.R
Document & audit make_corr_mat

@@ -118,6 +118,8 @@ get_crit = function( p.dat, col.p ) {
 
 ########################### FN: RETURN CORRELATION BETWEEN TWO ARBITRARY CELLS ###########################
 
+# AUDITED
+
 # Returns correlation between a pair of variables given
 #  desired correlation structure.
 
@@ -246,27 +248,25 @@ cell_corr = function( vname.1,
 
 ########################### FN: CREATE CORRELATION MATRIX ###########################
 
+# Makes correlation matrix
+
+# Notes:
+# If correlation matrix isn't positive definite, try reducing some of the correlations
+
+# Arguments:
 # .nX: number of covariates including the one of interest
 # .nY: number of outcomes
-
-# if correlation matrix isn't positive definite, try reducing some of the correlations
+# .rho.XX: Correlation between pairs of Xs
+# .rho.XY: Correlation between pairs of X-Y (of non-null ones)
+# .prop.corr: Proportion of X-Y pairs that are non-null
+#  (non-nulls will be first .prop.corr * .nY pairs)
 
 make_corr_mat = function( .nX,
                           .nY,
                           .rho.XX,
                           .rho.YY,
                           .rho.XY,
-                          #.half = FALSE,
                           .prop.corr = 1) {
-  
-  
-  # TEST ONLY
-  # .nX = 1
-  # .nY = 40
-  # .rho.XX = 0
-  # .rho.YY = 0
-  # .rho.XY = 0.15
-  # .prop.corr = 0.5
   
   nVar = .nX + .nY
   
@@ -275,6 +275,7 @@ make_corr_mat = function( .nX,
   Y.names = paste0( "Y", 1 : .nY )
   vnames = c( X.names, Y.names )
   
+  # initialize correlation matrix
   cor = as.data.frame( matrix( NA, nrow = nVar, ncol = nVar ) )
   names(cor) = vnames
   row.names(cor) = vnames
@@ -282,28 +283,34 @@ make_corr_mat = function( .nX,
   # populate each cell 
   for ( r in 1:dim(cor)[1] ) {
     for ( c in 1:dim(cor)[2] ) {
-      cor[ r, c ] = cell_corr( vnames[r], vnames[c], .rho.XX, .rho.YY, .rho.XY, .nY, .prop.corr )
+      cor[ r, c ] = cell_corr( vname.1 = vnames[r],
+                               vname.2 = vnames[c],
+                               .rho.XX = .rho.XX,
+                               .rho.YY = .rho.YY,
[…]

### 2018-06-07 R code/For Sherlock/functions.R
Documented & audited sim_data

@@ -248,6 +248,8 @@ cell_corr = function( vname.1,
 
 ########################### FN: CREATE CORRELATION MATRIX ###########################
 
+# AUDITED :) 
+
 # Makes correlation matrix
 
 # Notes:
@@ -315,18 +317,24 @@ make_corr_mat = function( .nX,
 
 ########################### FN: SIMULATE 1 DATASET ###########################
 
-# simulate 1 dataset with correlated covariates
-# then simulate 
+# AUDITED :) 
+
+# Simulates 1 dataset with MVN(0,1) correlated covariates and outcomes
+
+# Arguments: 
+# .n: sample size
+# .cor: correlation matrix from above function
 
 sim_data = function( .n, .cor ) {
 
+  # variable names
   vnames = names( .cor )
   
   # simulate the dataset
   # everything is a standard Normal
   d = as.data.frame( rmvnorm( n = .n,
-                              mean = rep( 0, dim( .cor )[1] ),
-                              sigma = as.matrix( .cor ) ) )
+                              mean = rep( 0, dim(.cor)[1] ),
+                              sigma = as.matrix(.cor) ) )
   names(d) = vnames
   
   # return the dataset
@@ -335,14 +343,15 @@ sim_data = function( .n, .cor ) {
 
 # # test drive
 # cor = make_corr_mat( .nX = 1,
-#                .nY = 6,
+#                .nY = 40,
 #                .rho.XX = 0,
 #                .rho.YY = 0.25,
 #                .rho.XY = 0.1,
-#                .prop.corr = 0.5 )
+#                .prop.corr = 0.2 )
 # 
-# d = sim_data( .n = 5000, .cor = cor )
-# round( cor(d), 2 ); cor
+# d = sim_data( .n = 1000, .cor = cor )
+# # plot empirical vs. real correlations
+# plot( as.numeric(cor(d)), as.numeric(as.matrix(cor)) ); abline( a = 0, b = 1, col="red")
 
 
 

### 2018-06-08 R code/For Sherlock/functions.R
Documented & audited fit_model

@@ -357,10 +357,23 @@ sim_data = function( .n, .cor ) {
 
 ########################### FN: FIT REGRESSION MODEL FOR A PARTICULAR OUTCOME ###########################
 
-# fit regression model for a single outcome Y regressed on all Xs
-# return p-value for the exposure of interest (X1)
+# AUDITED :) 
 
-# p: parameters matrix
+# Fits regression model for a single outcome Y regressed on all Xs
+# Returns p-value and various stats for just the exposure of interest (X1)
+
+# Arguments: 
+#
+# Y.name: Quoted name of a single outcome
+#
+# .dat: Dataframe
+#
+# .center.stats: If TRUE, subtract .bhat.orig (original data estimates) from 
+#   estimates (for use with bootstrapping under null)
+#
+# .bhat.orig: Original data estimates for use with .center.stats=TRUE; ignored
+#   otherwise. Even though fit_model is for only one outcome, .bhat.orig should
+#  have length W (one per outcome). 
 
 fit_model = function( Y.name,
                       .dat,
@@ -378,7 +391,6 @@ fit_model = function( Y.name,
   RHS = paste(X.names, collapse=" + ")
   formula = paste( c( Y.name, RHS ), collapse = " ~ ")
   
-  
   # extract p-value for exposure of interest (X1)
   m = lm( eval( parse(text=formula) ), data=.dat )
 
@@ -393,8 +405,10 @@ fit_model = function( Y.name,
     tval.return = summary(m)$coefficients[ "X1", 3 ]
   }
   if( .center.stats ) {
-    # second term pulls out the correct beta-hat from the original vector by using the Yname that we're regressing on
-    bhat.return = coef(m)[["X1"]] - .bhat.orig[ as.numeric( substr( Y.name, 2, nchar(Y.name) ) ) ]
+    # pull out the correct beta-hat from the original vector by using the
+    #  Yname that we're regressing on
+    bhat.orig.w = .bhat.orig[ as.numeric( substr( Y.name, 2, nchar(Y.name) ) ) ]
+    bhat.return = coef(m)[["X1"]] - bhat.orig.w
     tval.return = bhat.return / se.return
   }
   
@@ -410,10 +424,31 @@ fit_model = function( Y.name,
   
 }
 
-# test drive
-# cor = make_corr_mat(3, 40)
+# # test drive 1: not centered
+# cor = make_corr_mat( .nX = 1,
[…]

### 2018-06-08 R code/For Sherlock/functions.R
Documented & audited dataset_result

@@ -451,15 +451,28 @@ fit_model = function( Y.name,
 #                  .bhat.orig = c( NA, res$bhat ) )  # NA because we've only fit one model
 
 
+
 ########################### FN: GIVEN DATASET, RETURN STATS ###########################
-# .resid: should it return dataset with residuals (for regenerating residuals)?
-# . sigma: should it return fitted error SD (for regenerating residuals)?
-# .intercept: should it return fitted intercept (for regenerating residuals)?
-dataset_result = function(
-                          .dat,
+
+# AUDITED :) 
+
+# Arguments: 
+#
+# .alpha: Alpha-level for individual hypothesis tests
+#
+# .dat: Dataframe
+#
+# .center.stats: If TRUE, subtract .bhat.orig (original data estimates) from 
+#   estimates (for use with bootstrapping under null)
+#
+# .bhat.orig: Original data estimates for use with .center.stats=TRUE; ignored
+#   otherwise. Even though fit_model is for only one outcome, .bhat.orig should
+#  have length W (one per outcome). 
+
+dataset_result = function( .dat,
                            .alpha,
                            .center.stats = FALSE,
-                           .bhat.orig = NA ) {  # used for centering test stats
+                           .bhat.orig = NA ) {  
 
   # extract names of outcome variables
   X.names = names( .dat )[ grep( "X", names(.dat) ) ]
@@ -473,8 +486,7 @@ dataset_result = function(
   #  each entry is another list
   # with elements "pval" (scalar) and "resid" (length matches number of subjects)
   lists = lapply( X = Y.names,
-                  FUN = function(y) fit_model(
-                                                Y.name = y,
+                  FUN = function(y) fit_model( Y.name = y,
                                                .dat = .dat,
                                                .center.stats = .center.stats,
                                                .bhat.orig = .bhat.orig ) )
@@ -484,12 +496,12 @@ dataset_result = function(
   pvals = as.vector( u[ names(u) == "pval" ] )
 
   # save residuals
[…]

### 2018-06-08 R code/For Sherlock/functions.R
Documented & audited adj_minP

@@ -2,10 +2,15 @@
 
 ######################## FNS FOR WESTFALL's SINGLE-STEP ########################
 
-# returns minP-adjusted p-values (single-step)
+# AUDITED :) 
 
-# p = original p-values
-# p.bt = bootstrapped p-values (vector) - an m X B matrix
+# Returns minP-adjusted p-values (single-step)
+# See Westfall text, pg. 48.
+
+# Arguments: 
+# p: Original p-values (vector)
+# p.bt: Bootstrapped p-values (an m X B matrix)
+
 adjust_minP = function( p, p.bt ) {
   
   n.boot = ncol(p.bt)
@@ -20,14 +25,23 @@ adjust_minP = function( p, p.bt ) {
   return(p.adj)
 }
 
-# # test
+# # sanity check
 # B = 200
 # n.tests = 10
+# 
+# # generate fake p-values under strong null
 # p.bt = matrix( runif(B*n.tests, 0, 1), nrow = n.tests)
-# p = runif( n.tests, 0, .1)
+# 
+# # generate fake p-values from real dataset
+# p = runif( n.tests, 0, .1)  
 # 
 # p.adj = adjust_minP( p, p.bt )
-# plot(p,p.adj)
+# plot(p, p.adj)
+# 
+# # manually adjust second p-value
+# mins = apply( p.bt, MARGIN = 2, FUN = min )
+# prop.table( table( mins <= p[2] ) )[["TRUE"]]
+# p.adj[2]
 
 
 ######################## FNS FOR WESTFALL's STEP-DOWN ########################
@@ -95,7 +109,7 @@ adj_Wstep = function( p, p.bt ) {
 # plot( p, p.adj.Wstep )
 
 
-
+########################### FN: CALCULATE CRITICAL VALUES FOR WESTFALL ###########################
 
 get_crit = function( p.dat, col.p ) {
   

### 2018-06-15 R code/For Sherlock/doParallel.R
Work on read-me file

@@ -1,5 +1,52 @@
 
+# ######### FOR DEBUGGING ON CLUSTER #########
+# 
+# # THIS WORKS
+# 
+# # load command line arguments
+# args = commandArgs(trailingOnly = TRUE)
+# jobname = "fake"
+# scen = args[2]  
+# boot.reps = as.numeric(args[3])
+# 
+# # get scen parameters
+# setwd("/home/groups/manishad/multTest")
+# scen.params = read.csv( "scen_params.csv" )
+# p = scen.params[ scen.params$scen.name == scen, ]
+# 
+# print(p)
+# 
+# # no longer included in parameters because it's a vector
+# crit.bonf = 0.05 / p$nY
+# alpha = c( crit.bonf, 0.01, 0.05 )
+# 
+# # simulation reps to run within this job
+# # this need to match n.reps.in.doParallel in the genSbatch script
+# sim.reps = 200
+# 
+# # EDITED FOR C++ ISSUE WITH PACKAGE INSTALLATION
+# library(doParallel, lib.loc = "/home/groups/manishad/Rpackages/")
+# library(foreach, lib.loc = "/home/groups/manishad/Rpackages/")
+# library(mvtnorm, lib.loc = "/home/groups/manishad/Rpackages/")
+# library(StepwiseTest, lib.loc = "/home/groups/manishad/Rpackages/")  # Romano
+# library(matrixcalc, lib.loc = "/home/groups/manishad/Rpackages/")
+# 
+# # for use in ml load R
+# # install.packages( c("doParallel", "foreach", "mvtnorm", "StepwiseTest", "matrixcalc"), lib = "/home/groups/manishad/Rpackages/" )
+# 
+# source("functions.R")
+# 
+# # set the number of cores
+# registerDoParallel(cores=16)
+
+
+
+
 ######### FOR CLUSTER USE #########
+
+# because Sherlock 2.0 restores previous workspace
+rm( list = ls() )
+
 # load command line arguments
 args = commandArgs(trailingOnly = TRUE)
 jobname = args[1]
@@ -19,7 +66,7 @@ alpha = c( crit.bonf, 0.01, 0.05 )
 
 # simulation reps to run within this job
 # this need to match n.reps.in.doParallel in the genSbatch script
-sim.reps = 5
+sim.reps = 20
 
 # EDITED FOR C++ ISSUE WITH PACKAGE INSTALLATION
[…]

### 2018-06-15 R code/For Sherlock/gen_sbatch.R

@@ -1,9 +1,5 @@
 
 
-# NOTE:
-# Had sent 1-1000 using normal,owners
-# Just sent 1001-1010 using manishad
-
 ########################### SET SIMULATION PARAMETERS MATRIX ###########################
 
 # FOR CLUSTER USE
@@ -14,17 +10,29 @@ setwd(path)
 # path = "~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/Sandbox/2018-1-13"
 # setwd(path)
 
+
+# FOR DEBUGGING ONLY
 n = 1000
 nX = 1
 nY = 40
 rho.XX = 0
-rho.YY = c(0, 0.1, 0.3, 0.6)
-rho.XY = c(0, 0.03, 0.05, 0.10, 0.15 ) # null hypothesis: 0
-prop.corr = c(0.20, 0.5, 1)  # exchangeable vs. half-correlated matrix
+rho.YY = c(0.6)
+rho.XY = c(0) # null hypothesis: 0
+prop.corr = c(1)  # exchangeable vs. half-correlated matrix
+
+# n = 1000
+# nX = 1
+# nY = 40
+# rho.XX = 0
+# rho.YY = c(0, 0.1, 0.3, 0.6)
+# rho.XY = c(0, 0.03, 0.05, 0.10, 0.15 ) # null hypothesis: 0
+# prop.corr = c(0.20, 0.5, 1)  # exchangeable vs. half-correlated matrix
 
 
+# ~~~ CHANGED FOR DEBUGGING
+boot.reps = 500
 # bootstrap iterates and type
-boot.reps = 2000
+#boot.reps = 2000
 bt.type = c( "ha.resid" )
 
 # matrix of scenario parameters
@@ -44,10 +52,14 @@ names(scen.params) = c( "bt.type", "n", "nX", "nY", "rho.XX", "rho.YY", "rho.XY"
 scen.params = scen.params[ scen.params$rho.XY != 0 |
                              scen.params$rho.XY == 0 & scen.params$prop.corr == 1, ]
 
+# choose order in which you want them to run
+scen.params = scen.params[ order(scen.params$rho.XY,
+                                 scen.params$prop.corr,
+                                 decreasing = FALSE), ]
 
 ######## Name the Scenarios ########
 # if merging results with other simulations, set this to the last letter already used
-start.at = 1
+start.at = 4
 
 # remove letters that are privileged variables in R
 # letter.names = c(letters, LETTERS, paste(letters, letters, sep="") )
[…]

### 2018-06-15 R code/For Sherlock/stitch_on_sherlock.R
Work on read-me file

@@ -18,7 +18,10 @@ nrow(s) / (n.reps.per.scen * nrow(scen.params))
 table(s$scen)
 
 # quick check
-mean(s$jt.rej.holm)
+mean(s$jt.pval.0.05)
+mean(s$rej.jt.0.05)
+mean(s$jt.rej.Romano)
+mean(s$jt.rej.Wstep)
 
 # move it to Desktop
 scp mmathur@login.sherlock.stanford.edu:/home/groups/manishad/multTest/sim_results/overall_stitched/stitched.csv ~/Desktop

### 2018-06-15 R code/For Sherlock/doParallel.R
Clear previous R workspace to avoid weird Sherlock 2.0 bugs

@@ -66,7 +66,7 @@ alpha = c( crit.bonf, 0.01, 0.05 )
 
 # simulation reps to run within this job
 # this need to match n.reps.in.doParallel in the genSbatch script
-sim.reps = 20
+sim.reps = 10
 
 # EDITED FOR C++ ISSUE WITH PACKAGE INSTALLATION
 library(doParallel, lib.loc = "/home/groups/manishad/Rpackages/")
@@ -469,7 +469,7 @@ for ( j in 1:sim.reps ) {
 
 # results dataframe should now have boot.reps * sim.reps rows
 
-# ~~~ TEMP - THESE RESULTS ACTUALLY MAKE SENSE...
+# sanity check of results
 r = results
 print( mean(r$jt.pval.0.05) )
 print( mean(r$rej.jt.0.05) )

### 2018-06-15 R code/For Sherlock/gen_sbatch.R
.

@@ -11,28 +11,27 @@ setwd(path)
 # setwd(path)
 
 
-# FOR DEBUGGING ONLY
-n = 1000
-nX = 1
-nY = 40
-rho.XX = 0
-rho.YY = c(0.6)
-rho.XY = c(0) # null hypothesis: 0
-prop.corr = c(1)  # exchangeable vs. half-correlated matrix
-
+# # ~~~~ FOR DEBUGGING ONLY
 # n = 1000
 # nX = 1
 # nY = 40
 # rho.XX = 0
-# rho.YY = c(0, 0.1, 0.3, 0.6)
-# rho.XY = c(0, 0.03, 0.05, 0.10, 0.15 ) # null hypothesis: 0
-# prop.corr = c(0.20, 0.5, 1)  # exchangeable vs. half-correlated matrix
+# rho.YY = c(0.6)
+# rho.XY = c(0) # null hypothesis: 0
+# prop.corr = c(1)  # exchangeable vs. half-correlated matrix
+
+n = 1000
+nX = 1
+nY = 40
+rho.XX = 0
+rho.YY = c(0, 0.1, 0.3, 0.6)
+rho.XY = c(0, 0.03, 0.05, 0.10, 0.15) # null hypothesis: 0
+prop.corr = c(0.20, 0.5, 1)  # exchangeable vs. half-correlated matrix
 
 
-# ~~~ CHANGED FOR DEBUGGING
-boot.reps = 500
+# boot.reps = 500
 # bootstrap iterates and type
-#boot.reps = 2000
+boot.reps = 1000
 bt.type = c( "ha.resid" )
 
 # matrix of scenario parameters
@@ -59,7 +58,7 @@ scen.params = scen.params[ order(scen.params$rho.XY,
 
 ######## Name the Scenarios ########
 # if merging results with other simulations, set this to the last letter already used
-start.at = 4
+start.at = 1
 
 # remove letters that are privileged variables in R
 # letter.names = c(letters, LETTERS, paste(letters, letters, sep="") )
@@ -84,8 +83,8 @@ source("functions.R")
 
 # ~~CHANGED
 # number of sbatches to generate (i.e., iterations within each scenario)
-n.reps.per.scen = 200
-n.reps.in.doParallel = 20
+n.reps.per.scen = 500
+n.reps.in.doParallel = 10
 # n.reps.per.scen = 500
 # n.reps.in.doParallel = 5
[…]

### 2018-06-16 Project log/multiple_outcomes_project_log.Rmd
Log new sims

@@
 
+\subsection{2018-6-XXX re-running Freedman simulations }
+
+Re-running Freedman simulations after catching bug in prop.corr function that was causing too few of the X-Y pairs to be correlated. Also reduced $B$ from 2000 to 1000 and sim.reps from 1000 to 500.
+
+Results do look a bit less stable with these reduced values.
+When everything is done running, maybe try running an additional 1000 sim reps of the strong null scenarios (1-4) to see if that causes all the methods to hit 5% power exactly.
+If not, then maybe this suggests we really do need B=2000 (and should put that in paper).
+
 \subsection{2018-4-7 unsaved local simulations using \texttt{mice\_validation.R} }

### 2018-06-16 R code/For Sherlock/analysis.R
.

@@ -12,7 +12,8 @@ s = read.csv("stitched.csv", header=TRUE)
 s = merge(s, scen.params, by = "scen" )
 
 # what percent done are we?
-nrow(s) / (1000 * nrow(scen.params))
+n.reps.per.scen = 500
+nrow(s) / (500 * nrow(scen.params))
 
 # how many reps per scenario do we have?
 table( s$scen )
@@ -139,7 +140,7 @@ pwr = pwr[ pwr$method != "meanP", ]
 pwr.short = pwr[ !pwr$rho.XY %in% c(0.02, .1), ]
 
 library(ggplot2)
-x.breaks = seq(0, 0.5, 0.25)
+x.breaks = c(0, 0.1, 0.3, 0.6)
 y.breaks = seq(0, 1, 0.1)
 
 colors = c( "#999999", "orange", "#009E73", "black", "#E69F00", "#D55E00", "black", "darkgreen" )
@@ -165,7 +166,7 @@ name = "joint_test_full.png"
 #ggsave( filename = paste(name),
 #        plot=p, path=NULL, width=10, height=8, units="in")
 
-
+p
 
 ########################### NULL CI PLOTS ###########################
 
@@ -261,9 +262,9 @@ p = ggplot( data = ci2 ) +
 # ggsave( filename = paste("null_ci_full.png"),
 #         plot=p, path=NULL, width=16, height=8, units="in")
 
-ggsave( filename = paste("null_ci_short.png"),
-        plot=p, path=NULL, width=10, height=8, units="in")
-
+# ggsave( filename = paste("null_ci_short.png"),
+#         plot=p, path=NULL, width=10, height=8, units="in")
+p
 
 
 ########################### OTHER STATS FOR PAPER ###########################

### 2018-06-16 R code/For Sherlock/gen_sbatch.R

@@ -122,10 +122,13 @@ n.files
 
 # 2600 files
 setwd( paste(path, "/sbatch_files", sep="") )
-for (i in 1:100) {
-  system( paste("sbatch -p manishad /home/groups/manishad/multTest/sbatch_files/", i, ".sbatch", sep="") )
+for (i in 401:500) {
+  system( paste("sbatch -p owners /home/groups/manishad/multTest/sbatch_files/", i, ".sbatch", sep="") )
 }
 
+# experiment to see if it ever runs
+sbatch -p normal /home/groups/manishad/multTest/sbatch_files/2599.sbatch
+sbatch -p owners /home/groups/manishad/multTest/sbatch_files/2600.sbatch
 
 # ######## If Running Only Some Jobs To Fill Gaps ######## 
 # 

### 2018-06-16 R code/sim_study_local.R

@@ -1,59 +1,0 @@
-
-# remove all variables
-#rm(list=ls(all=TRUE))
-
-#################### SET SCENARIO PARAMETERS #################### 
-n = 50
-alpha = c( 0.05 )
-nX = 1
-nY = 5000
-rho.XX = 0.1
-rho.YY = c( 0.5 )
-rho.XY = 0  # null
-
-reps = 1
-boot.reps = 1
-
-# matrix of scenario parameters
-scen.params = expand.grid( n, alpha, nX, nY, rho.XX, rho.YY, rho.XY )
-names(scen.params) = c( "n", "alpha", "nX", "nY", "rho.XX", "rho.YY", "rho.XY" )
-
-# name the scenarios
- scen.params$scen.name = c(letters, LETTERS)[ 1:dim(scen.params)[1] ]
-#scen.params$scen.name = c(letters, LETTERS)[ 7: ( dim(scen.params)[1] + 6) ]
-n.scen = length(scen.params[,1])
-
-
-
-#################### RUN SIMULATION #################### 
-setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2/R code/For Sherlock")
-source("functions.R")
-
-path = "~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2/Simulation results"
-setwd(path)
-
-# loop through the scenarios and run each one
-system.time(
-for (s in scen.params$scen.name[1]) {  # REMOVE THE 1
-    rs = sim_scenario( .scen=s, .reps=reps, .boot.reps = boot.reps)
-    rs$scen.name = s
-    rs2 = merge(scen.params, rs)
-    write.csv( rs2, paste(Sys.Date(), "_scen_", s, ".csv", sep="") )
-}
-)
-
-# stitch results files
-s = stitch_files(.results.singles.path = path,
-                      .name.prefix="_scen", .stitch.file.name="stitched_results.csv")
-
-# remove annoying rows
-
-# organize it and remove annoying rows
-s = s[ order(s$alpha, s$rho.YY), c( names(scen.params), "n.rejected", "lo.bt", "hi.bt" ) ]
-
-# higher correlation between Ys => wider CI
-# lower alpha => narrower CI
-
-#################### SUMMARIZE RESULTS ####################
-
-

### 2018-06-17 R code/For Sherlock/analysis.R
Clean up analysis code

@@ -1,4 +1,53 @@
 
+
+########################### HELPER FNS ###########################
+
+# Note that these access global variables for, e.g., method.label
+
+pwr_plot = function(dat) {
+  library(ggplot2)
+  ggplot( data = dat, aes( x = rho.YY, y = power,
+                           color = method,
+                           label = method ) ) +
+    geom_text( aes( label = method.label) ) +
+    theme_bw() +
+    facet_wrap( ~group) +
+    #facet_wrap(~ group, nrow = 2 ) +  # for changing rows/columns
+    ylab("Power") +
+    scale_color_manual( values = colors) +
+    scale_x_continuous( limits = c( min(x.breaks), max(x.breaks) ), breaks = x.breaks ) +
+    scale_y_continuous( limits = c( min(y.breaks), max(y.breaks) ), breaks = y.breaks ) +
+    xlab( "Correlation between each pair of outcomes" ) +
+    #ggtitle("Power of bootstrapped hypothesis test of joint null") +
+    theme(legend.position="none") # remove legend
+}
+
+ci_plot = function(dat) {
+  library(ggplot2)
+  ggplot( data = dat ) +
+    # bootstrap results
+    geom_point( aes( x = rho.YY, y = n.rej.bt.mn, color = method ), size=2.5 ) +
+    geom_errorbar( aes( x=rho.YY, ymin = bt.lo.mn,
+                        ymax = bt.hi.mn, color = method ), width=0.02, size=1.05 ) +
+    
+    # original dataset results
+    geom_point( aes( x = rho.YY, y = n.rej.mn, shape="the shape" ), color = "black", size=3.5 ) +
+    scale_shape_manual( values = c('the shape' = 4),
+                        name = "Original dataset", guide = 'legend', labels = c("Mean rejections")) +
+    
+    theme_bw() +
+    facet_wrap(~ group, ncol = 3 ) +
+    ylab("Average null CIs") +
+    scale_color_manual( name="Joint test", values = colors, labels = legend.labs ) +
+    
+    xlab( "Correlation between each pair of outcomes" ) +
+    #ggtitle("Average CI limits and rejections in resamples and originals") +
+    theme(legend.position="none") # remove legend
+}
+
+
+
+
[…]

### 2018-06-17 R code/For Sherlock/gen_sbatch.R
.

@@ -122,13 +122,12 @@ n.files
 
 # 2600 files
 setwd( paste(path, "/sbatch_files", sep="") )
-for (i in 401:500) {
+for (i in 2401:2600) {
   system( paste("sbatch -p owners /home/groups/manishad/multTest/sbatch_files/", i, ".sbatch", sep="") )
 }
 
-# experiment to see if it ever runs
-sbatch -p normal /home/groups/manishad/multTest/sbatch_files/2599.sbatch
-sbatch -p owners /home/groups/manishad/multTest/sbatch_files/2600.sbatch
+
+
 
 # ######## If Running Only Some Jobs To Fill Gaps ######## 
 # 

### 2018-06-17 R code/For Sherlock/stitch_on_sherlock.R

@@ -13,7 +13,7 @@ s = stitch_files(.results.singles.path = "/home/groups/manishad/multTest/sim_res
                  .stitch.file.name="stitched.csv")
 
 # how close are we to being done?
-n.reps.per.scen = 1000
+n.reps.per.scen = 500
 nrow(s) / (n.reps.per.scen * nrow(scen.params))
 table(s$scen)
 

### 2018-06-22 R package/NRejections/R/NRejections.R
Removed mean-centering from package checks

@@ -5,7 +5,7 @@
 # library(NRejections)
 # test()
 
-
+# PROBLEM: DO WE REALLY NEED MEAN-CENTERING? WHAT ABOUT CATEGORICAL COVARIATES?
 
 ########################### CHECK FOR BAD USER INPUT ###########################
 
@@ -51,21 +51,22 @@ fix_input = function( X,
     d = d[ , analysis.vars ]
   }
   
+  # ~~~~ REMOVED THIS
   ##### Mean-Centered Covariates #####
-  tolerance = .Machine$double.eps ^ 0.5
-  zero = rep( 0, length(covars) )
-  # as.matrix handles case where there is only 1 covariate
-  diffs = abs( apply( as.matrix(d[ , covars ]), 2, mean ) - zero )
-  
-  # indices of non-centered covariates
-  not.cent = which( diffs > tolerance )
-  
-  if ( length(not.cent) > 0 ) {
-    warning( paste( "The following covariates have been mean-centered, with implications for model interpretation: ", 
-                    paste( covars[not.cent], collapse = ", " ), 
-                    sep = "" ) )
-    d = as.data.frame( apply( d, 2, function(col) col - mean(col) ) )
-  }
+  # tolerance = .Machine$double.eps ^ 0.5
+  # zero = rep( 0, length(covars) )
+  # # as.matrix handles case where there is only 1 covariate
+  # diffs = abs( apply( as.matrix(d[ , covars ]), 2, mean ) - zero )
+  # 
+  # # indices of non-centered covariates
+  # not.cent = which( diffs > tolerance )
+  # 
+  # if ( length(not.cent) > 0 ) {
+  #   warning( paste( "The following covariates have been mean-centered, with implications for model interpretation: ", 
+  #                   paste( covars[not.cent], collapse = ", " ), 
+  #                   sep = "" ) )
+  #   d = as.data.frame( apply( d, 2, function(col) col - mean(col) ) )
+  # }
 }
 
 
@@ -122,6 +123,8 @@ corr_tests = function( d,
                        alpha = 0.05,
                        alpha.fam = 0.05,
                        method = "nreject" ) {
+  
+  #browser()
   
   # check for and fix bad user input
   d = fix_input( X = X,
@@ -284,7 +287,6 @@ fit_model = function( X,
                       bhat.orig = NA,  # bhat.orig is a single value now for just the correct Y
                       alpha = 0.05 ) {
 
-  #browser()
[…]

### 2018-06-22 R code/Auxiliary code/Check theory/check_theory.R
About to change S2 vs. S1 issue in check_theory

@@ -1,0 +1,119 @@
+
+# Goal: See if the convergence result seems to hold for OLS. 
+
+N = 10000
+
+##### Scenario 1: mean-centered ##### 
+# variables are mean-centered
+set.seed(451)
+varX = 3
+X = as.matrix( data.frame( int = rep(1, N), 
+                X1 = rnorm(N, mean=0, sd=sqrt(varX)) ) )
+              #  X2 = rnorm(N, mean=0, sd = 10) ) )
+
+##### Scenario 2: not mean-centered ##### 
+
+set.seed(451)
+mean = 5
+varX = mean^2
+X = as.matrix( data.frame( int = rep(1, N), 
+                           X1 = rnorm(N, mean=mean, sd=sqrt(varX)) ) )
+#  X2 = rnorm(N, mean=0, sd = 10) ) )
+
+
+t(X) %*% X
+S1
+S2
+
+A = solve( t(X) %*% X ) %*% t(X)
+dim(A)  # should be p X N
+
+# we're interested in the second regression coefficient (X1)
+Arow = A[2,]
+B = t(Arow)
+
+# sum of squared elements of B
+SSB = sum( B^2 )
+SSB * N
+# doesn't seem to blow up :) 
+
+
+
+##### Check Math - Version 2 ##### 
+
+### do I have the right expression for (X'X)?
+X1 = X[,2]
+
+S1 = sum(X1)
+S2 = sum(X1^2)
+
+# elements of matrix
+S1
+S2
+t(X) %*% X
+# yes :) 
+
+
+### do I have the right expression for (X'X)^{-1}?
+X1 = X[,2]
+
+c1 = 1 / ( N * S2 - S1^2 )
+
+# elements of matrix
+c1 * S2
+c1 * -S1
+c1 * N
+solve( t(X) %*% X )
+# yes :) 
+
+
+### do I have the right expression for A = (X'X)^{-1}?
+
+A = solve( t(X) %*% X ) %*% t(X) 
+
+# reproduce first column of A
+A[1,1]; c1 * ( S2 - X1[1]*S1 )
+A[2,1]; c1 * ( -S1 + X1[1]*N )
+# yes :) 
+
+
+### do I have the right expression for thing that's supposed to go to 0?
+# RHS of first line after "the regularity condition concerns the term"
+# vs. the next line
[…]

### 2018-06-22 R code/Auxiliary code/Compare joint tests/Compare Westfall minP vs step-down/compare_two_westfalls.R

@@ -1,0 +1,159 @@
+
+# Goal: 
+
+# load Westfall fns
+setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/git_multiple_outcomes/R code/For Sherlock")
+source("functions.R")
+
+
+######################## GENERATE DATA ########################
+
+# parameters chosen to have a lot of small p-values so that we can
+#  see how they're affected
+
+# try with my own example
+# source functions
+source("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/git_multiple_outcomes/R code/For Sherlock/functions.R")
+
+# generate data
+set.seed(452)
+n = 1000
+nY = 100
+nX = 1
+rho.XX = 0
+rho.YY = .25
+rho.XY = 0.10
+alpha = 0.05
+
+# make initial dataset from which to bootstrap
+cor = make_corr_mat( .nX = nX, .nY = nY, .rho.XX = rho.XX,
+                     .rho.YY = rho.YY, .rho.XY = rho.XY)
+d = sim_data( .n = n, .cor = cor )
+X.names = "X1"
+Y.names = names(d)[2:length(names(d))]
+
+samp.res = dataset_result( .dat = d, .alpha = 0.05, .resid = TRUE, .sigma = TRUE, .tval = TRUE )
+tvals = as.matrix( samp.res$tvals )
+pvals = as.matrix( samp.res$pvals )
+
+n.rej = samp.res$rej
+names(n.rej) = paste( "n.rej.", as.character(alpha), sep="" )
+
+
+######################## RUN BOOTSTRAPPING ########################
+
+source("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/git_multiple_outcomes/R code/For Sherlock/functions.R")
+
+boot.reps = 500
+
+library(doParallel)
+library(foreach)
+registerDoParallel(16)
+
+# run all bootstrap iterates
+# CHANGED TO CBIND
+sec = system.time( {
+  r = foreach( i = 1:boot.reps, .combine=cbind ) %dopar% {
+    # draw bootstrap sample 
+    ids = sample( 1:nrow(d), replace=TRUE )
+    
+    ##### Bootstrap Under Null #2 (Resample Residuals) #####
+    # see Westfall, pg. 133
+    if ( TRUE ) {
+      # fix the existing covariates
+      Xs = as.data.frame( d[ , 1 : length(X.names) ] )
[…]

### 2018-06-22 R code/Auxiliary code/Compare joint tests/compare_joint_tests.R

@@ -1,0 +1,217 @@
+
+# Goal: Compare 3 joint tests
+# 1.) Ours: Reject strong null when number of rejections is extreme
+# 2.) Westfall: Reject when smallest adjusted p-value is <0.05
+# 3.) Romano: Same
+
+
+# load Westfall fns
+setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/git_multiple_outcomes/R code/For Sherlock")
+source("functions.R")
+  
+
+######################## GENERATE DATA ########################
+
+# try with my own example
+# source functions
+source("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/git_multiple_outcomes/R code/For Sherlock/functions.R")
+
+# generate data
+set.seed(452)
+n = 500
+nY = 100
+nX = 1
+rho.XX = 0
+rho.YY = .25
+rho.XY = 0.20
+alpha = 0.05
+
+# make initial dataset from which to bootstrap
+cor = make_corr_mat( .nX = nX, .nY = nY, .rho.XX = rho.XX,
+                     .rho.YY = rho.YY, .rho.XY = rho.XY)
+d = sim_data( .n = n, .cor = cor )
+X.names = "X1"
+Y.names = names(d)[2:length(names(d))]
+
+samp.res = dataset_result( .dat = d, .alpha = 0.05, .resid = TRUE, .sigma = TRUE, .tval = TRUE )
+tvals = as.matrix( samp.res$tvals )
+pvals = as.matrix( samp.res$pvals )
+
+n.rej = samp.res$rej
+names(n.rej) = paste( "n.rej.", as.character(alpha), sep="" )
+
+
+
+
+######################## FN FOR ROMANO-WOLF ########################
+
+boot.reps = 200
+
+#nrej_romano = function( d, samp.res, boot.reps ) {
+  
+  Y.names = names(d)[2:length(names(d))]
+  
+  library(StepwiseTest)
+  # docs: https://cran.r-project.org/web/packages/StepwiseTest/StepwiseTest.pdf
+  
+  rm(.Random.seed, envir=globalenv())  # reset seed
+
+  bt.stat = foreach( i=1:boot.reps, .combine=cbind ) %do% {
+    
+    ids = sample( 1:nrow(d), replace=TRUE )
+    b = d[ids,]
+    
+    # get regression tvals for bootstrap sample
+    lists = lapply( X = Y.names,
+                    FUN = function(y) fit_model( Y.name = y,
[…]

### 2018-06-22 R code/Auxiliary code/Compare joint tests/plot_adjusted_pvalues.R

@@ -1,0 +1,174 @@
+
+
+# load Westfall fns
+setwd("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/git_multiple_outcomes/R code/For Sherlock")
+source("functions.R")
+
+root.path = "/Users/mmathur/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/git_multiple_outcomes/Manuscript/Analyses"
+setwd(root.path)
+
+
+######################## GENERATE DATA ########################
+
+# parameters chosen to have a lot of small p-values so that we can
+#  see how they're affected
+
+# try with my own example
+# source functions
+source("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/git_multiple_outcomes/R code/For Sherlock/functions.R")
+library(matrixcalc)
+library(mvtnorm)
+
+# generate data
+set.seed(452)
+n = 1000
+nY = 100
+nX = 1
+rho.XX = 0
+rho.YY = .25
+rho.XY = 0.08
+alpha = 0.05
+
+# make initial dataset from which to bootstrap
+cor = make_corr_mat( .nX = nX, .nY = nY, .rho.XX = rho.XX,
+                     .rho.YY = rho.YY, .rho.XY = rho.XY)
+d = sim_data( .n = n, .cor = cor )
+X.names = "X1"
+Y.names = names(d)[2:length(names(d))]
+
+samp.res = dataset_result( .dat = d, .alpha = 0.05, .center.stats = FALSE )
+
+bhats = samp.res$bhats
+tvals = as.matrix( samp.res$tvals )
+pvals = as.matrix( samp.res$pvals )
+resids = samp.res$resid
+
+n.rej = samp.res$rej
+names(n.rej) = paste( "n.rej.", as.character(alpha), sep="" )
+
+# name the variables
+X.names = paste0( "X", 1 : nX )
+Y.names = paste0( "Y", 1 : nY )
+
+
+######################## RUN BOOTSTRAPPING ########################
+
+source("~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/git_multiple_outcomes/R code/For Sherlock/functions.R")
+
+boot.reps = 500
+
+library(doParallel)
+library(foreach)
+registerDoParallel(16)
+
+# compute Y-hat using residuals
+Ys = d[ , Y.names ]  
+Yhat = Ys - resids
+
+# fix the existing covariates
[…]

### 2018-06-22 R code/Auxiliary code/Check theory/check_theory.R
Rmk 4.2 works without mean-centering!

@@ -81,7 +81,7 @@ A[2,1]; c1 * ( -S1 + X1[1]*N )
 # RHS of first line after "the regularity condition concerns the term"
 # vs. the next line
 
-c2 = 1 / ( N * S2^2 - 2 * S2 * S1^2 + (S2 / N) )
+c2 = 1 / ( N * S2^2 - 2 * S2 * S1^2 + (S1^4 / N) )
 #c2 = 1 / ( N * S2^2 - 2 * S2 * S1^2 + (S2 / N) )
 term = sum( ( (N * X1) - S1 )^2 )
 
@@ -89,23 +89,30 @@ N * sum( A[2,]^2 ); c2 * term
 # yes :) 
 # ~~~~~ DOESN'T WORK IF NOT MEAN-CENTERED
 
-### keep simplifying
+
+### keep simplifying the sum term
 # pull sum through
 ( N^2 * S2 - 2 * N * S1^2 + N*S1^2 ) * c2
 
-# final result
-(N^2 * S2 - N*S1^2 ) / (N * S2^2 - 2 * S2*S1^2 + S2/N )
+# penultimate
+(N^2 * S2 - N*S1^2 ) / (N * S2^2 - 2 * S2*S1^2 + (S1^4)/N )
 # still matches!! :D
+
+# last line
+num = (S2/N - S1^2/(N^2) )
+denom =  ( S2^2/(N^2) - 2 * (S2/N) * S1^2/(N^2) + (S1^4)/(N^4) )
+num/denom
+# still matches!! :D
+
 
 # if mean-centered, should approach this as n -> infty
 1/varX
 
-
+# if not mean-centered
+varX; num
+varX^2; denom
 
 ### understand what happens with mean-centering
-# last line
-(S2/N - S1^2/(N^2) ) / ( S2^2/(N^2) - 2 * (S2/N) * S1^2/(N^2) + S2/(N^4) )
-# numerator goes to Var(X)
 
 # if Var(X) = (E(X))^2, then denom goes to 0 becasue:
 ( E.X2 = varX + mean^2 )

### 2018-06-23 Applied example/analyses_applied_example.R
Clean up applied example folder

@@ -1,0 +1,398 @@
+
+
+# Note: To avoid rerunning analyses (e.g., bootstraps) and only have access to resulting objects, run this.
+# This also allows exact reproduction of bootstrap analyses. 
+
+root.path = "~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/git_multiple_outcomes/Applied example"
+setwd(root.path)
+setwd("Analysis objects for manuscript")
+load("analysis_objects.rds")
+
+
+############################## READ IN DATA ############################## 
+
+root.path = "~/Dropbox/Personal computer/HARVARD/THESIS/Thesis paper #2 (MO)/git_multiple_outcomes/Applied example"
+setwd(root.path)
+
+# read in data
+setwd("Prepped data")
+d = read.csv("flourish_prepped.csv", header=TRUE); nrow(d)
+# should be 2697
+
+# source helper functions
+setwd(root.path)
+source("helper_applied_example.R")
+
+# names of adjusted covariates
+# see SAS file 9
+covars = c( "A1SEPA_z",
+            "A1PAGE_M2",
+            "A1PRSEX",
+            "raceA",
+            "A1SE2",
+            "A1SE3",
+            "A1SE4",
+            "A1PC1",
+            "sibling",
+            "CEDUC4cat",
+            "A1PC14",
+            "A1SE9",
+            "A1SE7",
+            "A1SE8c",
+            "mom_smk",
+            "dad_smk",
+            "B1PA58",
+            "A1SE6" )
+
+# names of outcomes
+outcomes = c("flourish_z",
+             "emotion_z",
+             "social_z",
+             "psych_z",
+             "B1SPOSAFz",
+             "B1SQ1z",
+             "B1SSWBMSz",
+             "B1SSWBSIz",
+             "B1SSWBAOz",
+             "B1SSWBSCz",
+             "B1SSWBSAz",
+             "B1SPWBA1z",
+             "B1SPWBE1z",
+             "B1SPWBG1z",
+             "B1SPWBR1z",
+             "B1SPWBU1z",
+             "B1SPWBS1z")
+
+# # make codebook saying variable values
+# library(dataMaid)
+# makeCodebook(d)
+
+############################## APPENDIX TABLE E.3: DEMOGRAPHICS ##############################
+
+# merge table-friendly variable names into dataset
[…]

### 2018-06-23 Applied example/data_prep_applied_example.R

@@ -1,0 +1,124 @@
+
+############################## READ IN SAS DATA ############################## 
+
+library(rprojroot)
+root.path = rprojroot::find_rstudio_root_file()
+setwd(root.path)
+
+# # this part takes FOREVER, so is commented out
+# # read in SAS data from Ying
+# # N = 2773
+# setwd("Prepped data")
+# library(sas7bdat)
+# d = read.sas7bdat("flourish_ying.sas7bdat")
+# nrow(d)
+# 
+# # sanity check for inclusion criteria in beginning of Ying's SAS file 9
+# # should always be 2
+# table(d$A1STATUS)
+# table(d$B1STATUS)
+# # yes
+# 
+# # write as csv to avoid lengthy read-in process
+# write.csv(d, "flourish_ying.csv")
+
+# read in Ying's data as csv to avoid SAS conversion process
+setwd("Prepped data")
+read.csv("flourish_ying.csv", header = TRUE); nrow(d)
+
+
+
+
+############################## RECODE COVARIATES PER CHEN (2017) ############################## 
+
+library(car)
+
+setwd(root.path)
+source("helper_applied_example.R")
+
+# initialize new dataset with recoded variables
+d2 = d
+
+# recode some categoricals
+d2$A1PRSEX = recode( d2$A1PRSEX, " 1='a.Male'; 2='b.Female' ")
+table(d$A1PRSEX, d2$A1PRSEX)
+
+d2$raceA = recode( d2$race, " 'White'='a.White'; 'Black'='b.Black'; 'others'='c.Others' ")
+table(d$race, d2$raceA)
+
+d2$CEDUC4cat = recode( d2$CEDUC4cat, " 1='a.Less than high school'; 2='b.High school'; 3='c.Some college'; 4='d.College degree or more' ")
+table(d$CEDUC4cat, d2$CEDUC4cat)
+
+d2$A1SE7 = recode( d2$A1SE7, " 1='a.Rural'; 2='b.Small town'; 3='c.Medium town'; 4='d.Suburbs'; 5='e.City'; 6='f.Moved around'; 7='g.Unsure';")
+table(d$A1SE7, d2$A1SE7)
+
+d2$A1SE6 = recode( d2$A1SE6, " 1='a.Very important'; 2='b.Somewhat important'; 3='c.Not very important'; 4='d.Not at all important'; 7='e.Unsure' ")
+table(d$A1SE6, d2$A1SE6)
+
+# recode binaries from 1/2 to 0/1 scheme
[…]

### 2018-06-23 Applied example/analyses_applied_example.R
Tidy up applied example code

@@ -63,9 +63,6 @@ outcomes = c("flourish_z",
              "B1SPWBU1z",
              "B1SPWBS1z")
 
-# # make codebook saying variable values
-# library(dataMaid)
-# makeCodebook(d)
 
 ############################## APPENDIX TABLE E.3: DEMOGRAPHICS ##############################
 
@@ -76,6 +73,7 @@ cb = read.xlsx(file="Analysis dataset codebook.xlsx",
                sheetName="codebook",
                header=TRUE)
 
+# for now, only keep variables needed for table
 temp = d[ , covars ]
 names(temp) = cb$Long.name[ cb$Variable. %in% names(temp) ]
 
@@ -123,15 +121,19 @@ for( i in 1:length(outcomes) ) {
   resids[,i] = raw.res$resids
 }
 
-# sanity check: fit one model manually
+# housekeeping
+names(stats) = names(raw.res$stats)
+stats$outcome = outcomes  # gets turned into "1" for some reason
+ 
+
+# # sanity check: fit one model manually
 # m = lm( flourish_z ~ A1SEPA_z + A1PAGE_M2 + A1PRSEX + raceA + A1SE2 + A1SE3 + A1SE4 + A1PC1 + sibling + CEDUC4cat + A1PC14 + A1SE9 + A1SE7 + A1SE8c + mom_smk + dad_smk + B1PA58 + A1SE6,
 #     data = d)
 # summary(m)
 # confint(m)
+# # compare: 
+# stats[1,]
 
-# housekeeping
-names(stats) = names(raw.res$stats)
-stats$outcome = outcomes  # gets turned into "1" for some reason
 
 # observed rejections
 ( th.0.05 = sum(stats$reject.0.05) )
@@ -152,6 +154,21 @@ stats.pretty = data.frame( outcome = stats$outcome,
 stats.pretty$outcome = cb$Long.name[ cb$Variable. %in% stats.pretty$outcome ]
 
 print( xtable(stats.pretty), include.rownames = FALSE )
+
+
+# # sanity check - check the ones with exactly duplicated results
+# # positive affect
[…]

### 2018-06-23 R code/Auxiliary code/Check theory/check_theory.R

@@ -13,12 +13,12 @@ X = as.matrix( data.frame( int = rep(1, N),
 
 ##### Scenario 2: not mean-centered ##### 
 
-set.seed(451)
+#set.seed(451)
 mean = 5
-varX = mean^2
+varX = mean*3
 X = as.matrix( data.frame( int = rep(1, N), 
-                           X1 = rnorm(N, mean=mean, sd=sqrt(varX)) ) )
-#  X2 = rnorm(N, mean=0, sd = 10) ) )
+                           X1 = rnorm( N, mean=mean, sd=sqrt(varX) ),
+                           X2 = rnorm(N, mean=mean, sd = 20) ) )
 
 
 t(X) %*% X
@@ -105,12 +105,9 @@ num/denom
 # still matches!! :D
 
 
-# if mean-centered, should approach this as n -> infty
+# should approach this as n -> infty
 1/varX
 
-# if not mean-centered
-varX; num
-varX^2; denom
 
 ### understand what happens with mean-centering
 

### 2018-06-23 Applied example/analyses_applied_example.R
Tidy up applied example code

@@ -63,9 +63,6 @@ outcomes = c("flourish_z",
              "B1SPWBU1z",
              "B1SPWBS1z")
 
-# # make codebook saying variable values
-# library(dataMaid)
-# makeCodebook(d)
 
 ############################## APPENDIX TABLE E.3: DEMOGRAPHICS ##############################
 
@@ -76,6 +73,7 @@ cb = read.xlsx(file="Analysis dataset codebook.xlsx",
                sheetName="codebook",
                header=TRUE)
 
+# for now, only keep variables needed for table
 temp = d[ , covars ]
 names(temp) = cb$Long.name[ cb$Variable. %in% names(temp) ]
 
@@ -123,15 +121,19 @@ for( i in 1:length(outcomes) ) {
   resids[,i] = raw.res$resids
 }
 
-# sanity check: fit one model manually
+# housekeeping
+names(stats) = names(raw.res$stats)
+stats$outcome = outcomes  # gets turned into "1" for some reason
+ 
+
+# # sanity check: fit one model manually
 # m = lm( flourish_z ~ A1SEPA_z + A1PAGE_M2 + A1PRSEX + raceA + A1SE2 + A1SE3 + A1SE4 + A1PC1 + sibling + CEDUC4cat + A1PC14 + A1SE9 + A1SE7 + A1SE8c + mom_smk + dad_smk + B1PA58 + A1SE6,
 #     data = d)
 # summary(m)
 # confint(m)
+# # compare: 
+# stats[1,]
 
-# housekeeping
-names(stats) = names(raw.res$stats)
-stats$outcome = outcomes  # gets turned into "1" for some reason
 
 # observed rejections
 ( th.0.05 = sum(stats$reject.0.05) )
@@ -153,6 +155,8 @@ stats.pretty$outcome = cb$Long.name[ cb$Variable. %in% stats.pretty$outcome ]
 
 print( xtable(stats.pretty), include.rownames = FALSE )
 
+
+# BOOKMARK: STOPPED HERE
 
 # ############################## BOOTSTRAPPING ##############################
 # 

### 2018-06-23 Applied example/data_prep_applied_example.R

@@ -5,13 +5,12 @@ library(rprojroot)
 root.path = rprojroot::find_rstudio_root_file()
 setwd(root.path)
 
-# # this part takes FOREVER, so is commented out
+# ##### This part takes ~5-10 min to run, so is commented out #####
 # # read in SAS data from Ying
-# # N = 2773
 # setwd("Prepped data")
 # library(sas7bdat)
 # d = read.sas7bdat("flourish_ying.sas7bdat")
-# nrow(d)
+# nrow(d) # should be 2,773
 # 
 # # sanity check for inclusion criteria in beginning of Ying's SAS file 9
 # # should always be 2
@@ -24,8 +23,8 @@ setwd(root.path)
 
 # read in Ying's data as csv to avoid SAS conversion process
 setwd("Prepped data")
-read.csv("flourish_ying.csv", header = TRUE); nrow(d)
-
+d = read.csv("flourish_ying.csv", header = TRUE); nrow(d)
+# should be 2,773 still
 
 
 
@@ -40,11 +39,13 @@ source("helper_applied_example.R")
 d2 = d
 
 # recode some categoricals
+# see the file "00. Parental warmth and flourishing_data step.sas"
+# also see MIDUS codebooks or "Analysis dataset codebook" for the original source
 d2$A1PRSEX = recode( d2$A1PRSEX, " 1='a.Male'; 2='b.Female' ")
 table(d$A1PRSEX, d2$A1PRSEX)
 
-d2$raceA = recode( d2$race, " 'White'='a.White'; 'Black'='b.Black'; 'others'='c.Others' ")
-table(d$race, d2$raceA)
+d2$raceA = recode( d2$raceA, " 1='a.White'; 2='b.Black'; 3='c.Others' ")
+table(d$raceA, d2$raceA)
 
 d2$CEDUC4cat = recode( d2$CEDUC4cat, " 1='a.Less than high school'; 2='b.High school'; 3='c.Some college'; 4='d.College degree or more' ")
 table(d$CEDUC4cat, d2$CEDUC4cat)
@@ -55,8 +56,10 @@ table(d$A1SE7, d2$A1SE7)
 d2$A1SE6 = recode( d2$A1SE6, " 1='a.Very important'; 2='b.Somewhat important'; 3='c.Not very important'; 4='d.Not at all important'; 7='e.Unsure' ")
 table(d$A1SE6, d2$A1SE6)
 
[…]

### 2018-06-24 Applied example/analyses_applied_example.R
Tweak codebook

@@ -307,7 +307,7 @@ table( stats$pval <= alpha.bonf )
 # in order of reporting in manuscript
 
 # sample size
-N = nrow(d)
+( N = nrow(d) )
 
 # average correlation magnitude
 d.outs = d[, outcomes]
@@ -327,8 +327,12 @@ stats.pretty[ stats.pretty$outcome == "Overall flourishing (continuous)", ]
 # mean effect size
 mean(stats$b)
 
+# null intervals
+null.int.0.05
+null.int.0.01
+
 # exact binomial distribution inference (alpha = 0.05)
-qbinom( .025, size=1, prob=0.05)
+qbinom( .025, size=17, prob=0.05)
 qbinom( .975, size=17, prob=0.05)
 # prob of rejecting all 18
 ( pval.binom = .05^17 )
